### 2022-07-06 01_studies/01_Laborstudie ProVisioNET/Fitbit/fitbit_proj_final/acc.Rmd

---
title: "R Notebook"
output: html_notebook
---
# Packages

```{r}
library(needs)
needs(tidyverse,
      lubridate,
      janitor)
```

# Daten

```{r}
df_acc <- read_tsv(file = "data/ProVisioNET.tsv") %>% 
  clean_names()
```
```{r}
df_acc <- transmute(df_acc,
                    ID = str_replace(str_sub(participant_name,
                                             start = 19,
                                             end = 21
                                             ),
                                     pattern = "_",
                                     replacement = ""
                                     ),
                    gyro_x = as.numeric(str_replace(gyro_x,
                                                    pattern = ",",
                                                    replacement = "."
                                                    )
                                        ),
                    gyro_y = as.numeric(str_replace(gyro_y,
                                                    pattern = ",",
                                                    replacement = "."
                                                    )
                                        ),
                    acc_x = as.numeric(str_replace(accelerometer_x,
                                                    pattern = ",",
                                                    replacement = "."
                                                    )
                                        ),
                    acc_z = as.numeric(str_replace(accelerometer_z,
                                                    pattern = ",",
                                                    replacement = "."
                                                    )
                                        ),
                    time = as.numeric(milliseconds(recording_timestamp
                                                   )
                                      ),
                    sensor = sensor,
                    meter_x = rep(0),
                    meter_z = rep(0)
                    ) %>% 
  filter(as.numeric(ID) > 100) %>%
  select(!contains("gyro")) %>%
  na.omit() %>%
  select(!"sensor")

# df_acc <- df_acc %>% 
#   filter(as.numeric(ID) > 100)
# 
# df_acc_acc <- df_acc %>%
#   filter(sensor == "Accelerometer") %>%
#   select(!contains("gyro")) %>%
#   pivot_longer(cols = c(acc_x, acc_z),
#                names_to = "acc")
# 
# df_acc_gyro <- df_acc %>%
#   filter(sensor == "Gyroscope") %>%
#   select(!contains("acc"))
```

```{r}
plot_acc <- 
ggplot(data = df_acc_acc,
       mapping = aes(x = time,
                     y = value
                     )
       ) +
  geom_path() +
  geom_path() +
  facet_grid(rows = vars(ID),
             cols = vars(acc)
             )

ggsave(plot = plot_acc,
       filename = "Plots/plot_acc.pdf",
       height = 10,
       width = 5)

# df_sum <- 
# df_acc_acc %>% 
#   group_by(acc, ID) %>%
#   summarize(sum = sum(abs(value)),
#             time = max(time)
#             )
```

```{r}
# First step: create tibbles for each individual.
for (i in unique(df_acc$ID)) {

  df <- filter(df_acc,
               ID == i
               )

  assign(value = df,
         x = paste0("df_", i)
         )

}

# For each of the tibbles, create two vectors (x and z) 
  # for meters for each time period.

for (i in unique(df_acc$ID)) {

  df <- get(paste0("df_", i))

# Vectors x and z with length equal to time periods 
  # present in each tibble; initial with 0 meters for each   # time period.

  x <- rep(0, length(get(paste0("df_", i))$time))
  z <- rep(0, length(get(paste0("df_", i))$time))

# For each time period, calculate the meters walked in 
  # two dimensions (x and z), given the initial velocity 
  # for each period is 0 m/s. First value stays 0 for
  # convenience.
  # See formula 1.

  for (j in 2:length(get(paste0("df_", i))$time)) {

    x[[j]] <- 0.5 * get(paste0("df_", i))$acc_x[j] * (get(paste0("df_", i))$time[j] - get(paste0("df_", i))$time[j-1])^2

    z[[j]] <- 0.5 * get(paste0("df_", i))$acc_z[j] * (get(paste0("df_", i))$time[j] - get(paste0("df_", i))$time[j-1])^2

  }

# Save both vectors in new tibble.

  df <- tibble(meter_x = x,
               meter_z = z,
               ID = paste0(i))
  assign(x = paste0("df_meter_", i),
           value = df)

}

# Combine all tibbles containing distance walked per 
  # time period.
  # Calculate sum of absolute distance walked for each 
  # participant and dimension.
  # See formula 2.

df_meter_all <- 
  mget(ls(pattern = "meter")) %>%
  bind_rows() %>% 
  group_by(ID) %>%
  summarise(meter_x = sum(abs(meter_x)),
            meter_z = sum(abs(meter_z)))

# Calculate walked distance in two-dimensional space.
  # See formula 3.

df_meter_all$ges <- sqrt(df_meter_all$meter_x^2 + df_meter_all$meter_z^2)
```

$$\Delta x_{d,i} = 0.5 (t_{i-1}-t_{i})^2, \text{ with }i=2,...,t_n; d = z,x$$
$$x_d=\sum_{i=1}^n |\Delta x_{d,i}|, \text{ with }i=2,...,t_n; d = z,x$$
$$\sum x=\sqrt{x_z^2+x_x^2}$$

- 4 Szenarien (still, still + kopf, viel gehen, wenig gehen, fahrstuhl, treppe) + whatever

```{r}
saveRDS(object = df_meter_all,
        file = "data/df_meter_all.rds")

rm(list = ls())
```

# Plot

```{r}
ggplot(data = df_acc_hr,
       mapping = aes(x = mean,
                      y = ges
                      )
        ) +
   geom_point(mapping = aes(color = status)) +
   geom_smooth(method = "lm",
               se = F) +
   labs(x = "Average Heart Rate\n(in Beats per Minute)",
        y = "Meters Walked\n(in Meters)")
```