### 2022-07-07 01_studies/01_Laborstudie ProVisioNET/Fitbit/fitbit_proj_final/Test Data.Rmd

---
title: "Test Data"
output: html_notebook
---

# Packages

```{r}
library(needs)
needs(tidyverse,
      janitor,
      lubridate)
```

# Data

```{r message=FALSE}
df_test <- read_tsv(file = "data/Acc_Test_Data.tsv") %>% 
  clean_names()
```

# Let the wrangle begin

```{r}
df_test <- 
  df_test %>%
  transmute(time = as.numeric(milliseconds(recording_timestamp_ms
                                           )
                              ),
            name = participant_name,
            acc_x = as.numeric(str_replace(accelerometer_x_m_s2,
                                           pattern = ",",
                                           replacement = "."
                                           )
                               ),
            acc_z = as.numeric(str_replace(accelerometer_z_m_s2,
                                           pattern = ",",
                                           replacement = "."
                                           )
                               )
            )
```

# Loops

```{r}
# First step: create tibbles for each individual.
for (i in unique(df_test$name)) {

  df <- filter(df_test,
               name == i
               )

  assign(value = df,
         x = paste0("df_", i)
         )

}

# For each of the tibbles, create two vectors (x and z) 
  # for meters for each time period.

for (i in unique(df_test$name)) {

  df <- get(paste0("df_", i))

# Vectors x and z with length equal to time periods 
  # present in each tibble; initial with 0 meters for each   # time period.

  x <- rep(0, length(get(paste0("df_", i))$time))
  z <- rep(0, length(get(paste0("df_", i))$time))

# For each time period, calculate the meters walked in 
  # two dimensions (x and z), given the initial velocity 
  # for each period is 0 m/s. First value stays 0 for
  # convenience.
  # See formula 1.

  for (j in 2:length(get(paste0("df_", i))$time)) {

    x[[j]] <- 0.5 * get(paste0("df_", i))$acc_x[j] * (get(paste0("df_", i))$time[j] - get(paste0("df_", i))$time[j-1])^2

    z[[j]] <- 0.5 * get(paste0("df_", i))$acc_z[j] * (get(paste0("df_", i))$time[j] - get(paste0("df_", i))$time[j-1])^2

  }

# Save both vectors in new tibble.

  df <- tibble(meter_x = x,
               meter_z = z,
               ID = paste0(i))
  assign(x = paste0("df_meter_", i),
           value = df)

}

# Combine all tibbles containing distance walked per 
  # time period.
  # Calculate sum of absolute distance walked for each 
  # participant and dimension.
  # See formula 2.

df_meter_all <- 
  mget(ls(pattern = "meter")) %>%
  bind_rows() %>% 
  group_by(ID) %>%
  summarise(meter_x = sum(abs(meter_x)),
            meter_z = sum(abs(meter_z))
            )

# Calculate walked distance in two-dimensional space.
  # See formula 3.

df_meter_all$ges <- sqrt(df_meter_all$meter_x^2 + df_meter_all$meter_z^2)
```

# Ergebnisse

```{r}
df_meter_all
```

```{r}
df_test %>%
  pivot_longer(cols = contains("acc"),
               names_to = "acc") %>%
ggplot(mapping = aes(x = time,
                     y = value
                     )
       ) +
  geom_path() +
  facet_grid(rows = vars(name),
             cols = vars(acc) 
             ) +
  theme(strip.text.y = element_text(angle = 0))
```