### 2022-12-20 02_paper/02_study/02_fitbit paper/Presentation_Analysis_update.Rmd
create updated presentation

---
title: "FitBit Analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
```

```{r data wrangling, include = FALSE}
library(needs)
needs(ltm,
      broom,
      ppcor,
      jtools,
      lm.beta,
      tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr,
      lme4,
      viridis)

# disruption & confidence rating
df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  select(ID, # select relevant columns 
         disruption_factor, # -99 = subject did not notice event; -100 = experimenter did not asked for event
         confident_factor,
         event
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confident_factor = ifelse(confident_factor < 0,
                                   yes = NA,
                                   no = confident_factor),
         event = as_factor(event)
         )

# demographic data
df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            gender = factor(LI02_01_1,
                            levels = 1:2,
                            labels = c("male","female")
                            ),
            age = LI03_01, # 1 = male; 2 = female
            teaching_experience = LI04_01)

# merge two data frames by ID 
df_merge <- merge(df_demo,
                  df_rating,
                  by = "ID") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132")
         )

# rm(list = c("df_demo","df_rating"))

data_path <- "./data/heart_rate_data"

main <-
map(.x = dir(path = data_path,
             pattern = ".csv"),
    ~ read_csv(file.path(data_path, .), 
               id = "id",
               col_types = c("t","n")
               ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "after",
                               str_detect(string = id, pattern = "_m") ~ "main",
                               str_detect(string = id, pattern = "_p") ~ "pre",
                               str_detect(string = id, pattern = "_s") ~ "subsequent",
                               str_detect(string = id, pattern = "_i") ~ "interview",
                               TRUE ~ "overall"
                               ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
      # filter(time <= 600) %>% # filter for 10min intervals
    filter(time <= 9000) %>% # filter for 2 hours 30min (maximal duration of study)
    select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

# main <-
#   left_join(main, df_merge,
#             by = "ID")
```

# Plot Gesamtverlauf

```{r}
main %>% 
  filter(time_span == "overall") %>% 
  group_by(ID) %>% 
  summarise(mean_hr = mean(heart_rate),
            sd_hr = sd(heart_rate)) %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  transmute(ID = ID,
            heart_rate_std = (heart_rate - mean_hr) / sd_hr,
            )

```