### 2023-02-15 02_paper/02_study/02_fitbit paper/FitBit_Plots_MK.Rmd
fitbit

---
title: "Fitbit & ET Glasses - Methods"
output: html_notebook
---
```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Packages
# install.packages("needs")
# Choose "No"
library(needs)
needs(tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr)
```

# Initial Preparations

Run following Code only if new data is added to the initial data.

```{r message=FALSE, warning=FALSE,include=FALSE}
# fitbit data
source(file = "data.wrangling.R")

# demographic data
# prepare data (selected from questionnaire data)
demo.data <- readxl::read_xlsx("./Data/data_empschul_labor_lehrperson.xlsx")

# select relevant columns
demo.data <- demo.data %>% select(LI02_01_1,
                                  LI03_01,
                                  LI04_01,
                                  LI06_05)

# rename columns
demo.data <- rename(demo.data, c("ID" = "LI06_05",
                                  "Gender" = "LI02_01_1",
                                  "Age" = "LI03_01",
                                  "Teaching Experience" = "LI04_01"))

# exclude cases with no fitbit data
demo.data <- demo.data %>% filter(!ID %in% c("126",
                                             "132",
                                             )
                                  )

```

# FitBit

All participants were given a FitBit Smart Watch to wear during the experiment.

## Heart Rate

```{r}
# aggregate data

df_heart_all <-
  read_rds(file = "data/heart_all.rds") |> 
  mutate(status = factor(status,
                         levels = c("Expert","Novice"),
                         labels = c("Expert","Novice")
                         )
         )

df_heart_min <-
  read_rds(file = "data/heart_min.rds")

df_heart_aggr <- 
  read_rds(file = "data/heart_aggr.rds") |> 
  mutate(status = factor(status,
                         levels = c("Novice","Expert"),
                         labels = c("Novice","Expert")
                         )
         )

# Mean, SD and Cohen D for experts & novices
# novices
mean_nov <- 
  df_heart_aggr %>%
  filter(status == "Novice" ) %>% 
  pull(mean) %>% 
  mean() %>% 
  round(., digits = 0)

sd_nov <- 
  df_heart_aggr %>%
  filter(status == "Novice") %>% 
  pull(mean) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_exp <-
  df_heart_aggr %>%
  filter(status == "Expert") %>% 
  pull(mean) %>% 
  mean() %>% 
  round(., digits = 0)

sd_exp <- 
  df_heart_aggr %>%
  filter(status == "Expert") %>% 
  pull(mean) %>% 
  sd() %>% 
  round(., digits = 0)

# effect size for expertise differences
d_heartrate <- CohenD(x = df_heart_aggr$mean[df_heart_aggr$status == "Expert"],
               y = df_heart_aggr$mean[df_heart_aggr$status == "Novice"],
               na.rm = TRUE)

```

The heart rate of each participant were measured during the experiment.

```{r}
# Plots

############## FLOW PLOT #############
plot_flow <-
ggplot(data = df_heart_all,
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`,
                     group = ID)
       ) +
  geom_line(mapping = aes(color = status),
            size = 0.3,
            alpha = 1
            ) +
  xlim(0, 600) +
  scale_color_viridis_d(option = "A",
                       end = 0.8)  +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.text = element_text(size = 16)) +
  facet_grid(cols = vars(Type),
             scales = "free_x")

############### FLOW PLOT EXPERT NOVICE #############

plot_flow_leg <-
ggplot(data = df_heart_all,
       mapping = aes(x = Time,
                     y = `Heart Rate`
                     )
       ) +
  xlim(0, 600) +
  geom_smooth(mapping = aes(color = status),
              method = "loess") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  ggtitle("Experts' and novices' heart rate per phase") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        strip.text = element_text(size = 16),
        plot.title = element_text(size = 20, 
                                  face = "bold")) +
  facet_grid(cols = vars(Type),
             scales = "free_x")

legend <- cowplot::get_legend(plot_flow_leg)

# saving plot 
ggsave(plot = plot_flow_leg,
       filename = "Plots/plot_flow_leg.svg",
       height = 5,
       width = 8,
       units = "in")

############### LOESS PLOT #############

plot_loess <-
ggplot(data = filter(df_heart_all),
       mapping = aes(x = as.numeric(Time),
                     y = `Heart Rate`
                     )
       ) +
  stat_smooth(geom = "line",
              mapping = aes(color = status),
              method = "loess",
              se = FALSE) +
  xlim(0,600) +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  labs(x = "Time (in Seconds)",
       y = "Heart Rate\n(in Beats per Minute)") +
  theme_minimal() +
  facet_grid(cols = vars(Type),
             scales = "free_x",
             ) +
  theme(text = element_text(family = "serif"),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

############### BOX PLOT #############

##### Box 1 Plot ##### 
plot_box1 <- 
ggplot(data = df_heart_aggr,
       mapping = aes(x = Type,
                     y = mean,
                     color = status
                     )
       ) +
  geom_boxplot(color = "darkgrey",
               fill = "grey") +
  geom_line(data = filter(df_heart_aggr,
                          status == "Novice"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 1,
            alpha = 0.5
            ) +
  geom_line(data = filter(df_heart_aggr,
                          status == "Expert"),
            mapping = aes(x = Type,
                          y = mean,
                          group = as_factor(ID)
                          ),
            linetype = 6,
            alpha = 0.5
            ) +
  geom_point(size = 2,
             alpha = 0.5) +
  labs(x = "",
       y = "Average Heart Rate\n(in Beats per Minute)") +
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        strip.text = element_blank(),
        legend.position = "none",
        axis.text.x = element_blank()
        ) +
  aes(fill = Type)

plot_box1

##### Box 2 Plot ##### 
plot_box <-
  ggplot(data = df_heart_aggr,
         mapping = aes(x = Type,
                       y = mean)) + 
  geom_boxplot(mapping = aes(fill = status),
               outlier.shape = NA) +
  geom_point(mapping = aes(shape = status),
             alpha = 0.5,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  labs(x ="",
       y = "Average Heart Rate\n(in Beats per Minute)") + 
  ggtitle("Experts' and novices' heart rate per phase") +
  scale_fill_brewer(palette  = "RdBu") +  
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.text = element_text(size = 15),
        legend.title = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        strip.text = element_text(size = 16),
        plot.title = element_text(size = 20, 
                                  face = "bold"))

plot_box

# saving plot 
ggsave(plot = plot_box,
       filename = "Plots/plot_box.svg",
       height = 5,
       width = 8,
       units = "in")

##############

# lo_mat <- rbind(c(1,1,1,1,1,1,NA),
#                 c(2,2,2,2,2,2,4),
#                 c(3,3,3,3,3,3,NA))
# 
# plot_heart <- 
# grid.arrange(grobs = list(plot_flow, plot_loess, plot_box1, legend),
#              layout_matrix = lo_mat)
# 
# plot_heart

```

# Repeated Measures ANOVA

```{r}

# disruption & confidence rating
df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  select(ID, # select relevant columns 
         disruption_factor, # -99 = subject did not notice event; -100 = experimenter did not asked for event
         confidence_factor,
         event
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confidence_factor = ifelse(confidence_factor < 0,
                                   yes = NA,
                                   no = confidence_factor),
         event = as_factor(event)
         ) %>% 
  na.omit() 
  # group_by(ID) %>% 
  # summarise("mean_disrup" = round(mean(disruption_factor), 2),
  #           "mean_confi" = round(mean(confidence_factor), 2))

# demographic data
df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            gender = factor(LI02_01_1,
                            levels = 1:2,
                            labels = c("male","female")
                            ),
            age = LI03_01, # 1 = male; 2 = female
            teaching_experience = LI04_01)

# merge two data frames by ID 
df_merge <- merge(df_demo,
                  df_rating,
                  by = "ID") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132")
         )

# rm(list = c("df_demo","df_rating"))

data_path <- "./data/heart_rate_data"

main <-
map(.x = dir(path = data_path,
             pattern = ".csv"),
    ~ read_csv(file.path(data_path, .), 
               id = "id",
               col_types = c("t","n")
               ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "after",
                               str_detect(string = id, pattern = "_m") ~ "main",
                               str_detect(string = id, pattern = "_p") ~ "pre",
                               str_detect(string = id, pattern = "_s") ~ "subsequent",
                               str_detect(string = id, pattern = "_i") ~ "interview",
                               TRUE ~ "overall"
                               ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
      # filter(time <= 600) %>% # filter for 10min intervals
    filter(time <= 9000) %>% # filter for 2 hours 30min (maximal duration of study)
    select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

main <-
  left_join(main, df_merge,
            by = "ID")

df_anova <-
  main %>% 
  filter(time_span != "overall") %>%   
  select("ID",
         "time_span",
         "heart_rate",
         "teaching_experience",
         "disruption_factor",
         "confidence_factor") %>% 
  mutate(time_span = fct_recode(time_span,
                                "Pre-Lehrphase" = "pre",
                                "Lehrphase" = "main",
                                "Post-Lehrphase" = "subsequent",
                                "Interviewphase" = "interview",
                                "Endphase" = "after"
                                ),
         time_span = factor(time_span,
                            levels = c("Pre-Lehrphase",
                                       "Lehrphase",
                                       "Post-Lehrphase",
                                       "Interviewphase",
                                       "Endphase"
                                       )
                            )
         ) %>%
  distinct() %>% 
  group_by(ID, time_span) %>%
  summarise(mean_hr = mean(heart_rate, na.rm = TRUE),
            mean_confi = mean(disruption_factor, na.rm = TRUE),
            mean_disrup = mean(disruption_factor, na.rm = TRUE),
            mean_te = mean(teaching_experience, na.rm = TRUE)
         )

# # merge demo data and fitbit data
# df_anova <- merge(df_anova,
#                   df_heart_aggr,
#                   by="ID")

# clean variable names
# df_anova <- df_anova %>% clean_names()

# Visualization
bxp <- ggboxplot(df_anova, x = "time_span", y = "mean_hr", add = "point")
bxp

# check for outliers
outliers <-
  df_anova %>%
  group_by(time_span) %>%
  identify_outliers(mean_hr)

# check for normality assumption --> can be checked by computing Shapiro-Wilk test for each time point. If the data is normally distributed, the p-value should be greater than 0.05
norm_assump <-
  df_anova %>%
  group_by(time_span) %>%
  shapiro_test(mean_hr)

# QQ plot draws the correlation between a given data and the normal distribution. 
plot <- 
  ggqqplot(df_anova, "mean_hr", facet.by = "time_span")

df_anova_subset <- subset(df_anova, select = c(time_span, mean_hr))

# # anova
# res.aov <- rstatix::anova_test(data = df_anova_subset,
#                                dv = mean_hr,
#                                wid = ,
#                                within = time_span)
# get_anova_table(res.aov)

# anova
df_anova_phase <- 
  aov(df_anova_subset$mean_hr~df_anova_subset$time_span)
summary(df_anova_phase)

# pairwise comparisons
pwc <- 
  pairwise.t.test(df_anova_subset$mean_hr, 
                  df_anova_subset$time_span, 
                  p.adjust="holm")

# effect size of ANOVA
install.packages("DescTools")
library(DescTools)
EtaSq(df_anova_phase)

effect_size_anova <-
  sqrt(0.3399757/(1-0.3399757))

# effect size of pairwise comparisons
library(dplyr)
library(rstatix)
df_anova_subset %>%
cohens_d(mean_hr~time_span) %>% 
as.data.frame()

# # pairwise comparisons
# pwc <- df_anova_subset %>%
#   pairwise_t_test(
#     mean_hr ~ time_span, paired = TRUE,
#     p.adjust.method = "bonferroni"
#     )
# pwc

# # Visualization: box plots with p-values
# pwc <- 
#   pwc %>% 
#   add_xy_position(x = "time_span")
# 
# # boxplot
# bxp_anova <- ggboxplot(data = df_anova, 
#                  x = "type", 
#                  y = "mean", 
#                  add = "point")
# 
# bxp_anova + 
#   stat_pvalue_manual(pwc) +
#   labs(
#     subtitle = get_test_label(res.aov, detailed = TRUE),
#     caption = get_pwc_label(pwc)
#     )

# # saving plot 
# ggsave(plot = bxp_anova,
#        filename = "Plots/bxp_anova.svg",
#        height = 5,
#        width = 8,
#        units = "in")

```

## Steps

Using the FitBit we could also measure the steps a person walked. Unfortunately, the measurement was not as reliable as hoped. To confirm, we should simulate situations with a fixed number of steps.

```{r}
df_steps <- read_rds(file = "final_data/steps.rds")
set.seed(5678)
ggplot(data = df_steps,
       mapping = aes(x = type,
                     y = Steps
                     )
       ) +
  geom_boxplot(mapping = aes(color = type),
               fill = "grey",
               alpha = 0.6) +
  geom_jitter(mapping = aes(color = type),
              alpha = 0.9,
              width = 0.3) + 
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  labs(x = "",
       y = "Steps During\nMain Phase") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
        )

### Validation of steps 
# selecting some rows at random with sample_n()
sample_n(df_steps, 6)

# Output:
# Experts: 219, 218, 214
# Novices: 108, 117, 110

```

# Combined Analysis

## Steps and Heart Rate

```{r}
ggplot(data = df_steps,
       mapping = aes(x = Steps,
                     y = mean)
       ) +
  geom_point(mapping = aes(shape = type,
                           color = type),
             alpha = 0.7,
             size = 2.5) +
    geom_smooth(mapping = aes(color = type),
              method = "lm",
              se = F) +
  scale_color_viridis_d(option = "A",
                       end = 0.8
                       ) +
  labs(x = "Steps",
       y = "Average Heart Rate\n(in Beats per Minute)"
       ) +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        legend.title = element_blank(),
        )
```

## Heart Rate and Pupil Dilation

```{r}
df_dil_aggr <- read_rds(file = "final_data/heart_dilation.rds")

df_dil_aggr %>%
  pivot_longer(cols = contains("dil"),
               names_to = "left_right",
               values_to = "Dilation") %>%
  mutate(type = factor(type,
                       levels = c("Expert","Novice"),
                       labels = c("Expert","Novice")
                       )
         ) |> 
  ggplot(mapping = aes(x = heart_rate_z,
                       y = Dilation)) +
  geom_point(mapping = aes(color = type),
             alpha = 0.7) +
  scale_color_viridis_d(option = "A",
                       end = 0.8) +
  geom_smooth(method = "lm",
              linetype = 2,
              color = "darkviolet") +
  labs(x = "Heart Rate\n(z-standardized)") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        text = element_text(family = "serif")) +
  facet_grid(cols = vars(factor(left_right,
                                levels = c("dil_left_z","dil_right_z"),
                                labels = c("Left Eye","Right Eye")
                                )
                         )
             ) 
```

# Measuring

```{r, warning=FALSE, message=FALSE}
df_heart_all %>% 
  clean_names() %>% 
  pivot_wider(names_from = time,
              values_from = heart_rate,
              names_prefix = "x")  %>%
  select(!contains("x")) ->
data_cache

df_heart_all %>% 
  clean_names() %>% 
  pivot_wider(names_from = time,
              values_from = heart_rate,
              names_prefix = "x")  %>%
  filter(id == "101") %>% 
  rowwise() %>%
  na_mean() %>%
  pivot_longer(cols = contains("x"),
               names_to = "time",
               values_to = "heart_rate") %>% 
  mutate(time = str_remove_all(time, "x")) %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate
                       )
         ) +
  geom_line()

 mutate(id = data_cache$id,
        ) %>% 
 right_join(x = ., 
            y = data_cache,
            by = "id") %>% 
 pivot_longer(cols = contains("x"),
               names_to = "time",
               values_to = "heart_rate") %>%
  mutate(time = str_remove_all(string = time,
                               pattern = "x"),
         time = as.numeric(time),
         heart_rate = as.numeric(heart_rate),
         id = as_factor(id)) %>%
  filter(id == "101") %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate,
                       group = id)
       ) + 
  geom_line(mapping = aes(color = status),
            size = 0.3,
            alpha = 1
            ) +
  xlim(0, min(df_heart_min$min)) +
  facet_grid(cols = vars(type))

  df_heart_all_wide

for (i in colnames(df_heart_all_wide)[-1:-4]) {

  df_heart_all_wide %>% 
    select(i)

}
df_heart_all_wide %>% 
  select(contains("x")) %>%
  as.matrix() %>% 
  rowwise() %>% 
  na_kalman() %>% 
  as_tibble() ->
  imp

imp$id <- data_cache$id

left_join(data_cache, imp,
          by = "id") %>% 
   pivot_longer(cols = contains("x"),
               names_to = "time",
               values_to = "heart_rate") %>%
  filter(id == "101") %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate,
                       group = id)
       ) + 
  geom_line(mapping = aes(color = status),
            size = 0.3,
            alpha = 1
            )

df_heart_all %>% 
  clean_names() %>% 
  pivot_wider(names_from = time,
              values_from = heart_rate,
              names_prefix = "x")  %>%
  filter(id == "101", type == "Post") %>% 
  select(!c("id","type","status")) %>% 
  pivot_longer()
  na_kalman()

imp <- na_kalman(tsAirgap)
ggplot_na_imputations(tsAirgap, imp)

df_heart_all %>% 
  clean_names() %>% 
  pivot_wider(names_from = time,
              values_from = heart_rate,
              names_prefix = "x") %>% 
  pivot_longer(cols = contains("x"),
               names_to = "time",
               values_to = "heart_rate") %>% 
  group_by(id, type) %>% 
  mutate(heart_rate = na_kalman(heart_rate)) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = time, 
                       y = heart_rate,
                       group = id
                       )
         ) +
  geom_line()

```