### 2023-03-07 02_paper/02_study/02_fitbit paper/presentation_analysis_std.Rmd
update paper analysis

---
title: "Presentation_Analysis_std"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
```

# Preps

```{r}
library(needs)
needs(ltm,
      broom,
      ppcor,
      jtools,
      lm.beta,
      tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr,
      lme4,
      viridis,
      gridExtra,
      gridtext)

# disruption & confidence rating
df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  dplyr::select(ID, # select relevant columns 
         disruption_factor, # -99 = subject did not notice event; -100 = experimenter did not asked for event
         confidence_factor,
         event
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confidence_factor = ifelse(confidence_factor < 0,
                                   yes = NA,
                                   no = confidence_factor),
         event = as_factor(event)
         )

# demographic data
df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            gender = factor(LI02_01_1,
                            levels = 1:2,
                            labels = c("male","female")
                            ),
            age = LI03_01, # 1 = male; 2 = female
            teaching_experience = LI04_01)

# merge two data frames by ID 
df_merge <- merge(df_demo,
                  df_rating,
                  by = "ID") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132")
         )

# rm(list = c("df_demo","df_rating"))

data_path <- "./data/heart_rate_data"

main <-
map(.x = dir(path = data_path,
             pattern = ".csv"),
    ~ read_csv(file.path(data_path, .), 
               id = "id",
               col_types = c("t","n")
               ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "after",
                               str_detect(string = id, pattern = "_m") ~ "main",
                               str_detect(string = id, pattern = "_p") ~ "pre",
                               str_detect(string = id, pattern = "_s") ~ "subsequent",
                               str_detect(string = id, pattern = "_i") ~ "interview",
                               TRUE ~ "overall"
                               ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
      # filter(time <= 600) %>% # filter for 10min intervals
    filter(time <= 9000) %>% # filter for 2 hours 30min (maximal duration of study)
    dplyr::select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

main <-
  left_join(main, df_merge,
            by = "ID")

main %>% 
  dplyr::select("time", "heart_rate", "ID", "time_span") %>% 
  distinct() %>% 
  # filter(time_span == "overall") %>% 
  group_by(ID) %>% 
  summarize(mean_heart = mean(heart_rate),
            sd_heart = sd(heart_rate)) %>%
  ungroup() %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  mutate(heart_rate_std = (heart_rate - mean_heart)/
           sd_heart) -> main
```

# Plots

## Phases

```{r}
loess_plot_overall <-
  main %>% 
  filter(time_span == "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  distinct() %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate_std
                       )
         ) +
  # geom_point(alpha = 0.05) + 
  geom_smooth(se = FALSE,
              size = 0.5,
              colour = "black") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  scale_x_continuous(expand = c(0, NA),
                     breaks = seq(0, 9000, 600)) +
  scale_y_continuous(breaks = seq(-1, 1, 0.5)) + 
  labs(x = "Zeit (in Sekunden)",
       y = "") +
  ggtitle("Gesamtverlauf der durchschnittlichen Herzraten") +
  theme_apa() + 
  theme(axis.title.y = element_blank())

lm_plot_phases <-
  main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("pre","main","subsequent","interview","after"),
                            labels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase"))) %>% 
  distinct() %>% 
  ggplot(mapping = aes(x = time,
                       y = heart_rate_std
                       )
         ) +
  # geom_point(alpha = 0.05) + 
  geom_smooth(method = "lm",
              se = FALSE,
              size = 0.5,
              colour = "black") +
  scale_color_viridis_d(option = "A",
                        end = 0.8) +
  scale_linetype_manual(values = c(1, 6)) +
  # scale_x_continuous(expand = c(0, NA),
  #                    breaks = seq(0, 9000, 600)) +
  #scale_y_continuous(breaks = seq(0, 160, 5)) + 
  labs(x = "Zeit (in Sekunden)",
       y = "") +
  ggtitle("Verlauf der durchschnittlichen Herzraten in unterschiedlichen Phasen\n(Lineare Schätzung)") +
  theme_apa() +
  theme(axis.title.y = element_blank()) +
  facet_grid(cols = vars(time_span),
             scales = "free_x")

ki_plot_phases <-
  main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("pre","main","subsequent","interview","after"),
                            labels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase"))) %>% 
  distinct() %>%
  group_by(time_span) %>% 
  summarise(z_mean = mean(heart_rate_std),
            z_sd = sd(heart_rate_std),
            z_t = qt(p = 0.01 / 2,      # 0.01 KI
               df = n() - 1,
               lower.tail = F
               ),
            z_error = z_sd / sqrt(n()),
            lower = z_mean - z_t * z_error,
            upper = z_mean + z_t * z_error,
            ) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = factor(0),
                       y = z_mean)) + 
  geom_point() +
  geom_errorbar(mapping = aes(ymin = lower,
                              ymax = upper),
                width = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed") +
  labs(title = "1% Konfidenzintervall für unterschiedliche Phasen",
       x = "") +
  facet_grid(cols = vars(time_span)) +
  theme_apa() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank())

text_y <- text_grob("Durchschnittliche Herzrate (Standardisiert)",
                   rot = 90)

plot_overall <-
  grid.arrange(loess_plot_overall, lm_plot_phases, ki_plot_phases,
               ncol = 1,
               left = text_y
               )

# saving plot 
ggsave(plot = plot_overall,
       filename = "plots/verlauf_plot.svg",
       height = 10,
       width = 8,
       units = "in")
```

# t-test

```{r}
heart_rate_split <- 
  main %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("pre","main","subsequent","interview","after"),
                            labels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase"))) %>% 
  distinct() %>% 
  group_by(time_span, ID) %>% 
  summarise(z.mean = mean(heart_rate_std))

heart_rate_test_t <- 
  matrix(data = NA,
         nrow = 5,
         ncol = 5)

colnames(heart_rate_test_t) <- unique(heart_rate_split$time_span)
rownames(heart_rate_test_t) <- unique(heart_rate_split$time_span)

for (i in colnames(heart_rate_test_t)) {

  for (j in colnames(heart_rate_test_t)) {

    col <- heart_rate_split %>% 
      filter(time_span == i)

    row <- heart_rate_split %>% 
      filter(time_span == j)

    t <- t.test(x = row$z.mean,
                y = col$z.mean,
                pairwise = TRUE
                )

    heart_rate_test_t[i,j] <- round(t$statistic,
                                    digits = 3)

  }

}

heart_rate_test_p <- 
  matrix(data = NA,
         nrow = 5,
         ncol = 5)

colnames(heart_rate_test_p) <- unique(heart_rate_split$time_span)
rownames(heart_rate_test_p) <- unique(heart_rate_split$time_span)

for (i in colnames(heart_rate_test_p)) {

  for (j in colnames(heart_rate_test_p)) {

    col <- heart_rate_split %>% 
      filter(time_span == i)

    row <- heart_rate_split %>% 
      filter(time_span == j)

    t <- t.test(x = row$z.mean,
                y = col$z.mean,
                pairwise = TRUE
                )

    heart_rate_test_p[i,j] <- round(t$p.value,
                                    digits = 3)

  }

}

```

# Regression

## Prep

```{r}
heart_rate_regression_data <-   
  main %>% 
  group_by(ID) %>% 
  mutate(disruption_factor = mean(disruption_factor, na.rm = TRUE),
         confidence_factor = mean(confidence_factor, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select("disruption_factor","confidence_factor","ID","teaching_experience") %>% 
  distinct()

heart_rate_regression <-   
  main %>% 
  group_by(ID) %>% 
  mutate(disruption_factor = mean(disruption_factor, na.rm = TRUE),
         confidence_factor = mean(confidence_factor, na.rm = TRUE)) %>% 
  ungroup() %>% 
  filter(time_span != "overall") %>%
  dplyr::select("ID","time_span", "time","heart_rate_std","disruption_factor","confidence_factor") %>% 
  mutate(time_span = factor(time_span,
                            levels = c("pre","main","subsequent","interview","after"),
                            labels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase")),
         ID = factor(ID)) %>% 
  distinct() %>% 
  mutate(time = time / 60) %>% 
  split(.$time_span) %>% 
  map(~ lm(formula = heart_rate_std ~ time * ID,
           data = .)) %>% 
  map(~ .$coefficients %>% 
        tibble(ID = names(.),
               value = .
               ) %>% 
        filter(str_detect(string = ID,
                          pattern = "time"
                          )
               ) %>% 
        mutate(ID = if_else(ID == "time",
                              true = "101",
                              false = ID),
               ID = str_extract_all(string = ID,
                                      pattern = "[0-9]+"),
               ID = unlist(ID),
               ID = as.numeric(ID)
               )
      ) %>% 
  bind_rows(., .id = "time_span") %>% 
   left_join(x = .,
             y = heart_rate_regression_data,
             by = "ID")
```

## Plots

```{r}
heart_rate_regression %>% 
  mutate(time_span = factor(time_span,
                            levels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase")
                            )
         ) %>% 
  ggplot(mapping = aes(x = disruption_factor,
                       y = value
                       )
         ) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Lineare Regression",
       subtitle = "Disruption Faktor",
       x = "Disruption Faktor",
       y = "Anstieg") +
  theme_apa() +
  facet_grid(cols = vars(time_span))

heart_rate_regression %>% 
  mutate(time_span = factor(time_span,
                            levels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase")
                            )
         ) %>% 
  ggplot(mapping = aes(x = confidence_factor,
                       y = value
                       )
         ) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Lineare Regression",
       subtitle = "Konfidenz Faktor",
       x = "Konfidenz Faktor",
       y = "Anstieg") +
  theme_apa() +
  facet_grid(cols = vars(time_span))
```

## Regression

```{r}
heart_rate_regression %>% 
  mutate(time_span = factor(time_span,
                            levels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase")
                            )
         ) %>% 
  split(.$time_span) %>% 
  map(~ lm(data = .,
      formula = value ~ disruption_factor + confidence_factor + teaching_experience)) %>% 
  stargazer::stargazer(type = "latex",
                       column.labels = c("Pre-Phase","Unterrichtsphase","Subsequentphase","Interviewphase","Post-Phase"),
                       dep.var.labels = "Slope",
                       covariate.labels = c("Disruption Faktor","Konfidenz Faktor","Lehrerfahrung"),
                       intercept.bottom = F,
                       intercept.top = T,
                       )
```