### 2023-03-28 01_studies/01_Laborstudie ProVisioNET/R script/et aoi/action_alternatives_confidence_factor.Rmd
action alternatives

---
title: "Action alternatives & confidence factor"
output: html_document
date: "2023-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages and read in data
```{r}

library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl, 
      gridExtra,
      janitor,
      RColorBrewer,
      cowplot)

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

# SELF-REPORT DATA
df_sri <- 
  excel_sheets("data/coding_sri.xlsx") %>% 
  map_df(~read_xlsx("data/coding_sri.xlsx",.)) %>% 
  dplyr::select(ID, Event, `Confidence Factor`) %>% 
  filter(
    ID %in% c(101:126, 201:228),
    !ID %in% c(204, 226),
    Event %in% c("chatting")) %>%
  janitor::clean_names()

# ACTION ALTERNATIVES count
df_alternatives_count <- read_excel("data/coding_action_alternatives.xlsx") %>% 
  janitor::clean_names() %>% 
  group_by(id) %>% 
  # mutate(mean_code = mean(code)) %>% 
  distinct(count)

# ACTION ALTERNATIVES non-disruptive 
df_alternatives_nd <- read_excel("data/coding_action_alternatives.xlsx") %>% 
  janitor::clean_names() %>% 
  select(id, code) %>% 
  filter(code %in% c(1:3)) %>%
  group_by(id) %>%
  count() %>% 
  transmute(nd = n)

# merge data frames
# put all data frames into list
df_list <- list(df_sri, df_alternatives_count, df_alternatives_nd)

# merge all data frames in list
main <-
  df_list %>% reduce(full_join, by='id') %>%
  janitor::clean_names() %>%
  select(!event)

# replace NA with 0
main[is.na(main)] <- 0

main$id = ifelse(main$id < 200, "Novice","Expert")

```

```{r}
# confidence factor
describeBy(main$confidence_factor, main$id)
plot_confi <- boxplot(confidence_factor ~ id, data = main)

# action alternatives - count
describeBy(main$count, main$id)
plot_action_count <- boxplot(count ~ id, data = main)

# action alternatives - non disruptive alternatives  
describeBy(main$nd, main$id)
plot_action_quality <- boxplot(nd ~ id, data = main)

```

# confidence factor x count of action alternatives
```{r}
main %>% 
  filter(!is.na(confidence_factor) | !is.na(count)) %>% 
  # filter(id %in% c("Novice")) %>% 
  ggplot(mapping = aes(x = confidence_factor,
                       y = count,
                       colour = id
                       )
         ) +
  geom_point() + 
  geom_smooth(method = "lm",
              se = TRUE) +
  scale_color_viridis_d() +
  labs(x = "Confidence Factor",
       y = "Count of action alternatives") +
  theme_cowplot()

```

# confidence factor x non-disruptive action alternatives
```{r}
main %>% 
  filter(!is.na(confidence_factor) | !is.na(nd)) %>% 
  # filter(id %in% c("Novice")) %>% 
  ggplot(mapping = aes(x = confidence_factor,
                       y = nd,
                       colour = id
                       )
         ) +
  geom_point() + 
  geom_smooth(method = "lm",
              se = TRUE) +
  scale_color_viridis_d() +
  labs(x = "Confidence Factor",
       y = "Non-disruptive action alternatives") + 
  theme_cowplot()

```

# pearson & spearman correlation

```{r}
# novices: correlation between confidence factor and count
main %>% 
  filter(id %in% c("Novice")) -> corr.nov

nov_confi_count_corr <- cor.test(corr.nov$confidence_factor, corr.nov$count, method=c("pearson"))
print(nov_confi_count_corr, short = F)

# experts: correlation between confidence factor and count
main %>% 
  filter(id %in% c("Expert")) -> corr.exp

exp_confi_count_corr <- cor.test(corr.exp$confidence_factor, corr.exp$count, method=c("pearson"))
print(exp_confi_count_corr, short = F)

# full sample: correlation between confidence factor and count
full_confi_count_corr <- corr.test(main$confidence_factor, main$count, method=c("pearson"), adjust="holm")
print(full_confi_count_corr, short = F)

# novices: correlation between confidence factor and non-disruptive alternatives
nov_confi_nd_corr <- cor.test(corr.nov$confidence_factor, corr.nov$nd, method=c("pearson"))
print(nov_confi_nd_corr, short = F)

# experts: correlation between confidence factor and non-disruptive alternatives
exp_confi_nd_corr <- cor.test(corr.exp$confidence_factor, corr.exp$nd, method=c("pearson"))
print(exp_confi_nd_corr, short = F)

# full sample: correlation between confidence factor and non-disruptive alternatives
full_confi_nd_corr <- corr.test(main$confidence_factor, main$nd, method=c("pearson"), adjust="holm")
print(full_confi_nd_corr, short = F)

```