### 2023-05-30 01_studies/01_Laborstudie ProVisioNET/R script/action alternatives/reaction_action_alternatives.Rmd
Update reaction_action_alternatives.Rmd

---
title: "Reaction & action alternatives"
output: html_document
date: "2023-05-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages and read in data
```{r}

library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl, 
      gridExtra,
      janitor,
      RColorBrewer,
      cowplot,
      zoo)

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

# REACTION 
df_reaction <- read_excel("data/coding_reactions_MK.xlsx") %>% 
  janitor::clean_names() %>%
  mutate(reaction = as.numeric(case_when(code > 3 ~ '0',
                                         code <= 3 ~ '1',
                                         TRUE ~ 'F')
                               )
         ) %>% 
  select(id, reaction)

# ACTION ALTERNATIVES count
df_alternatives_count <- read_excel("data/coding_action_alternatives.xlsx") %>% 
  na.locf() %>% # replacing NA with most forward non-NA to it
  janitor::clean_names() %>% 
  group_by(id) %>% 
  # mutate(mean_code = mean(code)) %>% 
  mutate(count = max(count)) %>% 
  distinct(count)

# ACTION ALTERNATIVES non-disruptive 
df_alternatives_nd <- read_excel("data/coding_action_alternatives.xlsx") %>% 
  na.locf() %>% 
  janitor::clean_names() %>% 
  select(id, code) %>% 
  mutate(nd = as.numeric(case_when(code > 3 ~ '0',
                                 code <= 3 ~ '1',
                                 TRUE ~ 'F')
                         )
         ) %>% 
  select(id, nd) %>%
  group_by(id) %>% 
  summarise(nd = sum(nd))

# merge data frames
# put all data frames into list
df_list <- list(df_reaction,
                df_alternatives_count,
                df_alternatives_nd)

# merge all data frames in list
main <-
  df_list %>% reduce(full_join, by='id') %>%
  janitor::clean_names()

# drop NA 
main <- 
  na.omit(main)

# replace IDs with expertise factor
main$id = ifelse(main$id < 200, "Novice","Expert")

# creating new column with total reactions
main_novice <- main %>%
  filter(id %in% c("Novice")) %>% 
  mutate(reaction_sum = sum(reaction),
         nd_sum = sum(nd),
         sum_participants = sum(n()),
         percent_reaction = round((reaction_sum/sum_participants),
                                  digits = 2),
         percent_nd = round((nd/count),
                            digits = 2)
         )

# creating new column with total reactions
main_expert <- main %>%
  filter(id %in% c("Expert")) %>% 
  mutate(reaction_sum = sum(reaction),
         nd_sum = sum(nd),
         sum_participants = sum(n()),
         percent_reaction = round((reaction_sum/sum_participants),
                                  digits = 2),

         percent_nd = round((nd/count),
                            digits = 2)
         )

# merge novice & expert data frame
main <- rbind(main_expert, main_novice)

```

# descriptives & plots
```{r}
# reaction 
describeBy(main$percent_reaction, main$id)
plot_action_count <- boxplot(percent_reaction ~ id, data = main)

# action alternatives - count
describeBy(main$count, main$id)
plot_action_count <- boxplot(count ~ id, data = main)

# action alternatives - non disruptive alternatives (sum score)
describeBy(main$nd, main$id)
plot_action_quality <- boxplot(nd ~ id, data = main)

# action alternative - non disruptive alternatives (relative score)
describeBy(main$percent_nd, main$id)
plot_reaction <- boxplot(percent_nd ~ id, data = main)

```

# t-test and effect size
```{r}

#### reaction
# t-test for expertise differences
t.test(x = main$reaction[main$id == "Expert"],
       y = main$reaction[main$id == "Novice"])

# effect size for expertise differences
CohenD(x = main$reaction[main$id == "Expert"],
       y = main$reaction[main$id == "Novice"],
       na.rm = TRUE)

### action alternatives count
# t-test for expertise differences
t.test(x = main$count[main$id == "Expert"],
       y = main$count[main$id == "Novice"])

# effect size for expertise differences
CohenD(x = main$count[main$id == "Expert"],
       y = main$count[main$id == "Novice"],
       na.rm = TRUE)

### action alternatives non-disruptive (sum score)
# t-test for expertise differences
t.test(x = main$nd[main$id == "Expert"],
       y = main$nd[main$id == "Novice"])

# effect size for expertise differences
CohenD(x = main$nd[main$id == "Expert"],
       y = main$nd[main$id == "Novice"],
       na.rm = TRUE)

### action alternatives non-disruptive (relative score)
# t-test for expertise differences
t.test(x = main$percent_nd[main$id == "Expert"],
       y = main$percent_nd[main$id == "Novice"])

# effect size for expertise differences
CohenD(x = main$percent_nd[main$id == "Expert"],
       y = main$percent_nd[main$id == "Novice"],
       na.rm = TRUE)

```

# correlation: count x efficacy