### 2023-07-15 02_paper/02_study/02_fitbit paper/test.Rmd
fitbit data

---
title: "Getting familiar with Rmarkdown PDF"
author: |
  | Arimitra Maiti
  | Author-2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    fig_caption: yes
    highlight: haddock
    number_sections: yes
    toc: yes
    toc_depth: 2
    citation_package: natbib
abstract: This is the section for abstract with a limit of 150 words.\par \textbf{Keywords:}
  Keyword-1, Keyword-2, Keyword-3, Keyword-4
subtitle: |
  | Examples for Towards Data Science
  | Focussing on Metadata
  | Reviewed by: Towards Data Science Team
geometry: left=2.5cm,right=2.5cm,top=2cm,bottom=2cm
fontfamily: mathpazo
thanks: |
  This is the section for acknoledgement with a limit of 150 words
header-includes:
- \usepackage{graphicx}
- \usepackage{float}
- \usepackage{amsmath}
- \nocite{*}
bibliography: example.bib
biblio-style: apalike
natbiboptions: round
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r echo=FALSE, results='asis'}
library("rmarkdown")
library("tinytex")
library("stargazer")
library("sandwich")

m1 = lm(y~ lag.quarterly.revenue, data = freeny)
m2 = lm(y~ lag.quarterly.revenue + price.index, data = freeny)
m3 = lm(y~ lag.quarterly.revenue + price.index + income.level, data = freeny)
m4 = lm(y~ lag.quarterly.revenue + price.index + income.level + market.potential, data = freeny)

model.lst = list(m1, m2, m3, m4)

stargazer(m1,
          m2,
          m3,
          m4,
          title="How quarterly revenue is affected by revenue, price index, inccome and market potential?",
          type = "latex",
          float = TRUE,
          report = "vcs*",
          se=lapply(model.lst, function(x) sqrt(diag(vcovHC(x, type = "HC1")))),
          no.space = TRUE,
          header=FALSE,
          single.row = TRUE,
          font.size = "small",
          intercept.bottom = F,
          column.labels = c("Iteration-1", "Iteration-2", "Iteration-3", "Iteration-4"),
          column.separate = c(1,1, 1, 1),
          digits = 2,
          t.auto = F,
          p.auto = F,
          notes.align = "l",
          notes = c("datasets::freeny", "lm() function", "vcovHC(type = 'HC1')-Robust SE"),
          notes.append = TRUE
          )

```