### 2024-03-12 01_studies/01_Laborstudie ProVisioNET/R script/et aoi/et_aoi.Rmd
et

---
title: "AOI TOI - number of fixations, duration of fixation & time to first fixation"
output: html_notebook
---

# Load packages and read in data
```{r}
library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl, 
      gridExtra,
      janitor,
      RColorBrewer,
      cowplot,
      svglite)

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

# return a character vector with names of .tsv data in data folder
file_names <- list.files(path = "data",
                         pattern = "interval_complete.tsv")

# read every object names in file_names and save it as a tibble
for (i in file_names) {

  work_data <- 
    read_tsv(file = paste0("data/", i),
             locale = locale(decimal_mark = ","))

  assign(value = work_data,
         x = str_remove(paste0("tib_", i),
                        pattern = ".tsv"
         )
  )

}

# Bind every tibble that contains "interval_complete" to a new tible
df_aoi <- 
  mget(ls(pattern = "interval_complete")) %>%
  bind_rows()

# Remove temporary data for a cleaner workspace
rm(list = ls(pattern = "^tib_ProVisio"))
rm(work_data)
rm(file_names)
rm(i)

```

# Number of fixation on all AOI

```{r}
##################### NUMBER OF FIXATION ON ALL AOIS ########################

# filter group & total duration & number fixation on all AOIs

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          Participant,
          starts_with("Total_duration_of_fixations"),
          starts_with("Number_of_fixations"),
          !"Total_duration_of_fixations.Disruptive_Person",
          !"Number_of_fixations.Disruptive_Person") %>% 
  rowwise() %>% 
  transmute(Group = Group,
            Participant = Participant,
            Sum_duration_fixation = sum(c_across(starts_with("Total_duration")
                                                    ),
                                           na.rm = TRUE
                                           ),
            Sum_number_fixation = sum(c_across(starts_with("Number_of")
                                                  ),
                                         na.rm = TRUE
                                         )
            ) %>% 
  mutate(Average_duration = round(Sum_duration_fixation/Sum_number_fixation,
                                         digits = 0),
         Group = as_factor(Group)) %>% 
  drop_na() -> df_aoi_sum

# mean & SD number fixation 
# novices
mean_number_nov <- 
  df_aoi_sum %>%
  filter(Group == "Novice") %>% 
  pull(Sum_number_fixation) %>% 
  mean() %>% 
  round(., digits = 0)

sd_number_nov <- 
  df_aoi_sum %>%
  filter(Group == "Novice") %>% 
  pull(Sum_number_fixation) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_number_exp <-
  df_aoi_sum %>%
  filter(Group == "Expert") %>% 
  pull(Sum_number_fixation) %>% 
  mean() %>% 
  round(., digits = 0)

sd_number_exp <- 
  df_aoi_sum %>%
  filter(Group == "Expert") %>% 
  pull(Sum_number_fixation) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting number of fixations on ALL AOIS
df_aoi_sum %>% 
  ggplot(mapping = aes(x = Group,
                       y = Sum_number_fixation)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste(
  #   c("Novice\n", "Expert\n"),
  #   glue("$M$ = "),
  #   c(mean_number_nov, mean_number_exp),
  #   glue("; $SD$ = "),
  #   c(sd_number_nov, sd_number_exp)
  # )) +

  # scale_x_discrete(labels = c(
  #   paste(
  #     "Novice\n",
  #     glue("<i>M</i>"),
  #     " = ",
  #     c(mean_number_nov, mean_number_exp)
  #   ),
  #   c(
  #     "Expert\n",
  #     ";",
  #     glue("<i>SD</i>"),
  #     " = ",
  #     c(sd_number_nov, sd_number_exp)
  #   ) 
  # )) +
  labs(x ="",
       y = "Anzahl Fixationen") +
  scale_fill_brewer(palette  = "RdBu") +  
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Anzahl der Fixationen in allen AOIs") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
        ) -> plot_number_group
plot_number_group

# saving plot number plot
ggsave(plot = plot_number_group,
       filename = "plots/plot_number_group.svg",
       height = 8,
       width = 14,
       units = "in")

# NUMBER OF FIXATIONS on all AOIs 
# t-test for expertise differences
t.test(x = df_aoi_sum$Sum_number_fixation[df_aoi_sum$Group == "Expert"],
       y = df_aoi_sum$Sum_number_fixation[df_aoi_sum$Group == "Novice"],
       var.equal = TRUE)

# NUMBER OF FIXATIONS on all AOIs
# effect size for expertise differences
d_number_all <- CohenD(x = df_aoi_sum$Sum_number_fixation[df_aoi_sum$Group == "Expert"],
                       y = df_aoi_sum$Sum_number_fixation[df_aoi_sum$Group == "Novice"],
                       na.rm = TRUE)

```

# Number of fixations on AOI "Disruptive person"

```{r}
##################### NUMBER OF FIXATION ON AOI DISRUPTIVE PERSON ########################

# filter group & total duration & number fixation on AOI Disruptive person

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          "Number_of_fixations.Disruptive_Person") %>%
  mutate(Group = as_factor(Group)) -> df_aoi_disrup

# mean & SD average duration AOI Disruptive person
# novices
mean_number_disrup_nov <- 
  df_aoi_disrup %>%
  filter(Group == "Novice") %>% 
  pull(Number_of_fixations.Disruptive_Person) %>% 
  mean() %>% 
  round(., digits = 0)

sd_number_disrup_nov <- 
  df_aoi_disrup %>%
  filter(Group == "Novice") %>% 
  pull(Number_of_fixations.Disruptive_Person) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_number_disrup_exp <-
  df_aoi_disrup %>%
  filter(Group == "Expert") %>% 
  pull(Number_of_fixations.Disruptive_Person) %>% 
  mean() %>% 
  round(., digits = 0)

sd_number_disrup_exp <- 
  df_aoi_disrup %>%
  filter(Group == "Expert") %>% 
  pull(Number_of_fixations.Disruptive_Person) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting average duration on AOI Disruptive person
df_aoi_disrup %>% 
  ggplot(mapping = aes(x = Group,
                       y = Number_of_fixations.Disruptive_Person)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  labs(x ="",
       y = "Anzahl Fixationen") + 
  scale_fill_brewer(palette  = "RdBu") +  
  ggtitle("Anzahl Fixationen AOI\n`Störperson`") +
    scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
  ) -> plot_number_disrup

# saving plot average plot
ggsave(plot = plot_number_disrup,
       filename = "plots/plot_number_disrup.svg",
       height = 8,
       width = 8,
       units = "in")

# AVERAGE DURATION OF FIXATIONS on AOI disruptive person 
# t-test for expertise differences
t.test(x = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Expert"],
       y = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Novice"],
       var.equal = TRUE)

# AVERAGE DURATION OF FIXATIONS 
# effect size for expertise differences
d_number_disrup <- CohenD(x = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Expert"],
                        y = df_aoi_disrup$Number_of_fixations.Disruptive_Person[df_aoi_disrup$Group == "Novice"],
                        na.rm = TRUE)

```

# Number of fixations on AOI "Students"

```{r}
##################### NUMBER OF FIXATIONS ON AOI STUDENTS ########################
# filter group & total duration & number fixation on AOI Students
df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          "Total_duration_of_fixations.Anna",
          "Total_duration_of_fixations.Bianca",
          "Total_duration_of_fixations.Carl(a)",
          "Number_of_fixations.Anna",
          "Number_of_fixations.Bianca",
          "Number_of_fixations.Carl(a)") %>% 
  rowwise() %>% 
  transmute(Group = Group,
            Stud_duration_fixation = sum(c_across(starts_with("Total_duration")
                                                 ),
                                         na.rm = TRUE
                                         ),
            Stud_number_fixation = sum(c_across(starts_with("Number_of")
                                                ),
                                       na.rm = TRUE
                                       )
            ) %>% 
  mutate(Average_duration_stud = round(Stud_duration_fixation/Stud_number_fixation, 
                                  digits = 0),
         Group = as_factor(Group)) -> df_aoi_stud

# mean & SD average duration on AOI Students
# novices
mean_number_stud_nov <- 
  df_aoi_stud %>%
  filter(Group == "Novice" ) %>% 
  pull(Stud_number_fixation) %>% 
  mean() %>% 
  round(., digits = 0)

sd_number_stud_nov <- 
  df_aoi_stud %>%
  filter(Group == "Novice") %>% 
  pull(Stud_number_fixation) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_number_stud_exp <-
  df_aoi_stud %>%
  filter(Group == "Expert") %>% 
  pull(Stud_number_fixation) %>% 
  mean() %>% 
  round(., digits = 0)

sd_number_stud_exp <- 
  df_aoi_stud %>%
  filter(Group == "Expert") %>% 
  pull(Stud_number_fixation) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting average duration on AOI Students
df_aoi_stud %>% 
  ggplot(mapping = aes(x = Group,
                       y = Stud_number_fixation)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  labs(x ="",
       y = "Anzahl Fixationen") + 
  scale_fill_brewer(palette  = "RdBu") +  
  ggtitle("Anzahl Fixationen AOI\n`Schüler:innen`") +
  theme_cowplot() +
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
  ) -> plot_number_stud

# saving plot average plot
ggsave(plot = plot_number_stud,
       filename = "plots/plot_number_stud.svg",
       height = 8,
       width = 8,
       units = "in")

# AVERAGE DURATION OF FIXATIONS on AOI Students 
# t-test for expertise differences
t.test(x = df_aoi_stud$Stud_number_fixation[df_aoi_stud$Group == "Expert"],
       y = df_aoi_stud$Stud_number_fixation[df_aoi_stud$Group == "Novice"],
       var.equal = TRUE)

# AVERAGE DURATION OF FIXATIONS on AOI Students
# effect size for expertise differences
d_number_stud <- CohenD(x = df_aoi_stud$Stud_number_fixation[df_aoi_stud$Group == "Expert"],
                      y = df_aoi_stud$Stud_number_fixation[df_aoi_stud$Group == "Novice"],
                      na.rm = TRUE)

```

# Average duration of fixations on all AOIs

```{r}
##################### AVERAGE DURATION ON ALL AOIS ########################

# filter group & total duration & number fixation on all AOIs

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          starts_with("Total_duration_of_fixations"),
          starts_with("Number_of_fixations"),
          !"Total_duration_of_fixations.Disruptive_Person",
          !"Number_of_fixations.Disruptive_Person") %>% 
  rowwise() %>% 
  transmute(Group = Group,
            Sum_duration_fixation = sum(c_across(starts_with("Total_duration")
                                                    ),
                                           na.rm = TRUE
                                           ),
            Sum_number_fixation = sum(c_across(starts_with("Number_of")
                                                  ),
                                         na.rm = TRUE
                                         )
            ) %>% 
  mutate(Average_duration = round(Sum_duration_fixation/Sum_number_fixation,
                                         digits = 0),
         Group = as_factor(Group)) -> df_aoi_sum

# mean & SD average duration 
# novices
mean_dur_nov <- 
  df_aoi_sum %>%
  filter(Group == "Novice") %>% 
  pull(Average_duration) %>% 
  mean() %>% 
  round(., digits = 0)

sd_dur_nov <- 
  df_aoi_sum %>%
  filter(Group == "Novice") %>% 
  pull(Average_duration) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_dur_exp <-
  df_aoi_sum %>%
  filter(Group == "Expert") %>% 
  pull(Average_duration) %>% 
  mean() %>% 
  round(., digits = 0)

sd_dur_exp <- 
  df_aoi_sum %>%
  filter(Group == "Expert") %>% 
  pull(Average_duration) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting average duration on ALL AOIS
df_aoi_sum %>% 
  ggplot(mapping = aes(x = Group,
                       y = Average_duration)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  labs(x ="",
       y = "Millisekunden") + 
  scale_fill_brewer(palette  = "RdBu") + 
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Durchschnittliche Fixationsdauer in allen AOIs") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
        ) -> plot_average_duration_group

# saving plot average plot
ggsave(plot = plot_average_duration_group,
       filename = "plots/plot_average_duration_group.svg",
       height = 8,
       width = 14,
       units = "in")

# AVERAGE DURATION OF FIXATIONS on all AOIs 
# t-test for expertise differences
t.test(x = df_aoi_sum$Average_duration[df_aoi_sum$Group == "Expert"],
       y = df_aoi_sum$Average_duration[df_aoi_sum$Group == "Novice"],
       var.equal = TRUE)

# AVERAGE DURATION OF FIXATIONS on all AOIs
# effect size for expertise differences
d_aver_all <- CohenD(x = df_aoi_sum$Average_duration[df_aoi_sum$Group == "Expert"],
                    y = df_aoi_sum$Average_duration[df_aoi_sum$Group == "Novice"],
                    na.rm = TRUE)

```

# Average duration of fixations on AOI "Disruptive Person"

```{r}
##################### AVERAGE DURATION ON AOI DISRUPTIVE PERSON ########################

# filter group & total duration & number fixation on AOI Disruptive person

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          "Total_duration_of_fixations.Disruptive_Person",
          "Number_of_fixations.Disruptive_Person") %>% 
  rowwise() %>% 
  mutate(Average_duration_disrup = round(Total_duration_of_fixations.Disruptive_Person/Number_of_fixations.Disruptive_Person, 
                                  digits = 0),
         Group = as_factor(Group)) -> df_aoi_disrup

# mean & SD average duration AOI Disruptive person
# novices
mean_dur_disrup_nov <- 
  df_aoi_disrup %>%
  filter(Group == "Novice") %>% 
  pull(Average_duration_disrup) %>% 
  mean() %>% 
  round(., digits = 0)

sd_dur_disrup_nov <- 
  df_aoi_disrup %>%
  filter(Group == "Novice") %>% 
  pull(Average_duration_disrup) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_dur_disrup_exp <-
  df_aoi_disrup %>%
  filter(Group == "Expert") %>% 
  pull(Average_duration_disrup) %>% 
  mean() %>% 
  round(., digits = 0)

sd_dur_disrup_exp <- 
  df_aoi_disrup %>%
  filter(Group == "Expert") %>% 
  pull(Average_duration_disrup) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting average duration on AOI Disruptive person
df_aoi_disrup %>% 
  ggplot(mapping = aes(x = Group,
                       y = Average_duration_disrup)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  labs(x ="",
       y = "Millisekunden") + 
  scale_fill_brewer(palette  = "RdBu") + 
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Ø Fixationsdauer AOI\n`Störperson`") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
  ) -> plot_average_duration_disrup

# saving plot average plot
ggsave(plot = plot_average_duration_disrup,
       filename = "plots/plot_average_duration_disrup.svg",
       height = 8,
       width = 8,
       units = "in")

# AVERAGE DURATION OF FIXATIONS on AOI disruptive person 
# t-test for expertise differences
t.test(x = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Expert"],
       y = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Novice"],
       var.equal = TRUE)

# AVERAGE DURATION OF FIXATIONS 
# effect size for expertise differences
d_aver_disrup <- CohenD(x = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Expert"],
                        y = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Novice"],
                        na.rm = TRUE)

```

# Average duration of fixations on AOI "Students"

```{r}
##################### AVERAGE DURATION ON AOI STUDENTS ########################

# filter group & total duration & number fixation on AOI Students
df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          "Total_duration_of_fixations.Anna",
          "Total_duration_of_fixations.Bianca",
          "Total_duration_of_fixations.Carl(a)",
          "Number_of_fixations.Anna",
          "Number_of_fixations.Bianca",
          "Number_of_fixations.Carl(a)") %>% 
  rowwise() %>% 
  transmute(Group = Group,
            Stud_duration_fixation = sum(c_across(starts_with("Total_duration")
                                                 ),
                                         na.rm = TRUE
                                         ),
            Stud_number_fixation = sum(c_across(starts_with("Number_of")
                                                ),
                                       na.rm = TRUE
                                       )
            ) %>% 
  mutate(Average_duration_stud = round(Stud_duration_fixation/Stud_number_fixation, 
                                  digits = 0),
         Group = as_factor(Group)) -> df_aoi_stud

# mean & SD average duration on AOI Students
# novices
mean_dur_stud_nov <- 
  df_aoi_stud %>%
  filter(Group == "Novice") %>% 
  pull(Average_duration_stud) %>% 
  mean() %>% 
  round(., digits = 0)

sd_dur_stud_nov <- 
  df_aoi_stud %>%
  filter(Group == "Novice") %>% 
  pull(Average_duration_stud) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_dur_stud_exp <-
  df_aoi_stud %>%
  filter(Group == "Expert") %>% 
  pull(Average_duration_stud) %>% 
  mean() %>% 
  round(., digits = 0)

sd_dur_stud_exp <- 
  df_aoi_stud %>%
  filter(Group == "Expert") %>% 
  pull(Average_duration_stud) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting average duration on AOI Students
df_aoi_stud %>% 
  ggplot(mapping = aes(x = Group,
                       y = Average_duration_stud)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  labs(x ="",
       y = "Millisekunden") + 
  scale_fill_brewer(palette  = "RdBu") +  
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Ø Fixationsdauer AOI\n `Schüler:innen`") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
  ) -> plot_average_duration_stud

# saving plot average plot
ggsave(plot = plot_average_duration_stud,
       filename = "plots/plot_average_duration_stud.svg",
       height = 8,
       width = 8,
       units = "in")

# AVERAGE DURATION OF FIXATIONS on AOI Students 
# t-test for expertise differences
t.test(x = df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Expert"],
       y = df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Novice"],
       var.equal = TRUE)

# AVERAGE DURATION OF FIXATIONS on AOI Students
# effect size for expertise differences
d_aver_stud <- CohenD(x = df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Expert"],
                      y = df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Novice"],
                      na.rm = TRUE)

```

# Time to first fixation on all AOIs

```{r}

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          starts_with("Time_to_first_fixation"),
          !"Time_to_first_fixation.Disruptive_Person") %>%
  rowwise() %>% 
  transmute(Group = Group,
            Sum_time_fixation = sum(c_across(starts_with("Time_to_first_fixation")
                                                    ),
                                           na.rm = TRUE
                                           ) 
            ) %>% 
  mutate(Sum_time_fixation_sec = round(Sum_time_fixation/1000,
                                         digits = 0),
         Group = as_factor(Group)
         ) -> df_ttff_sum

# mean & SD time to first fixation 
# novices
mean_time_nov <- 
  df_ttff_sum %>%
  filter(Group == "Novice") %>% 
  pull(Sum_time_fixation_sec) %>% 
  mean() %>% 
  round(., digits = 0)

sd_time_nov <- 
  df_ttff_sum %>%
  filter(Group == "Novice") %>% 
  pull(Sum_time_fixation_sec) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_time_exp <-
  df_ttff_sum %>%
  filter(Group == "Expert") %>% 
  pull(Sum_time_fixation_sec) %>% 
  mean() %>% 
  round(., digits = 0)

sd_time_exp <- 
  df_ttff_sum %>%
  filter(Group == "Expert") %>% 
  pull(Sum_time_fixation_sec) %>% 
  sd() %>% 
  round(., digits = 0)

# plotting average duration on ALL AOIS
df_ttff_sum %>% 
  ggplot(mapping = aes(x = Group,
                       y = Sum_time_fixation_sec)) +
  geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  labs(x ="",
       y = "Sekunden") + 
  scale_fill_brewer(palette  = "RdBu") +
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Zeit bis zur ersten Fixation in allen AOIs") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
        ) -> plot_time_group
plot_time_group

# saving plot average plot
ggsave(plot = plot_time_group,
       filename = "plots/plot_time_group.svg",
       height = 8,
       width = 14,
       units = "in")

# TIME TO FIRST FIXATIONS on all AOIs 
# t-test for expertise differences
t.test(x = df_ttff_sum$Sum_time_fixation_sec[df_ttff_sum$Group == "Expert"],
       y = df_ttff_sum$Sum_time_fixation_sec[df_ttff_sum$Group == "Novice"],
       var.equal = TRUE)

# TIME TO FIRST FIXATIONS on all AOIs 
# effect size for expertise differences
d_time_all <- CohenD(x = df_ttff_sum$Sum_time_fixation_sec[df_ttff_sum$Group == "Expert"],
                    y = df_ttff_sum$Sum_time_fixation_sec[df_ttff_sum$Group == "Novice"],
                    na.rm = TRUE)

```

# Time to first fixation on AOI "Disruptive Person"

```{r}
############### TIME TO FIRST FIXATION ####################

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          "Time_to_first_fixation.Disruptive_Person") %>% 
  rowwise() %>% 
  mutate(Group = as_factor(Group)) -> df_ttff_disrup

# changing milliseconds into seconds
df_ttff_disrup$Disrup_time_fixation_sec <-
  round(df_ttff_disrup$Time_to_first_fixation.Disruptive_Person / 1000,
        digits = 2)

# mean & SD time to first fixation "Disruptive Person"
# novices
mean_time_disrup_nov <- 
  df_ttff_disrup %>%
  filter(Group == "Novice") %>% 
  pull(Disrup_time_fixation_sec) %>% 
  mean() %>% 
  round(., digits = 0)

sd_time_disrup_nov <- 
  df_ttff_disrup %>%
  filter(Group == "Novice") %>% 
  pull(Disrup_time_fixation_sec) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_time_disrup_exp <-
  df_ttff_disrup %>%
  filter(Group == "Expert") %>% 
  pull(Disrup_time_fixation_sec) %>% 
  mean() %>% 
  round(., digits = 0)

sd_time_disrup_exp <- 
  df_ttff_disrup %>%
  filter(Group == "Expert") %>% 
  pull(Disrup_time_fixation_sec) %>% 
  sd() %>% 
  round(., digits = 0)

##################### plotting time to first fixation for groups ##################### 
plot_time_disrup <- 
  ggplot(data = df_ttff_disrup,
         mapping = aes(x = Group,
                       y = Disrup_time_fixation_sec)) +
 geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  # ylim(0,250) +
  labs(x ="",
       y = "Sekunden") + 
  scale_fill_brewer(palette  = "RdBu") +
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Zeit bis zur ersten Fixation\nAOI `Störperson`") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
  )

plot_time_disrup

# saving plot
ggsave(plot = plot_time_disrup,
       filename = "plots/plot_time_disrup.svg",
       height = 8,
       width = 8,
       units = "in")

# TIME TO FIRST FIXATIONS on disruptive person 
# t-test for expertise differences
t.test(x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"],
       y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"],
       var.equal = TRUE)

# TIME TO FIRST FIXATIONS on disruptive person 
# effect size for expertise differences
d_time_disrup <- CohenD(x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"],
                    y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"],
                    na.rm = TRUE)

```

# Time to first fixations on AOI "Students"

```{r}
# filter relevant rows and columns

df_aoi %>% 
  filter(TOI == "Lesson") %>% 
  select (Group,
          "Time_to_first_fixation.Anna",
          "Time_to_first_fixation.Bianca",
          "Time_to_first_fixation.Carl(a)") %>% 
  rowwise() %>% 
  transmute(Group = Group,
            Stud_time_fixation = sum(c_across(starts_with("Time_to_first_fixation")
                                                 ),
                                         na.rm = TRUE
                                         )
            ) %>% 
  mutate(Stud_time_fixation_sec = round(Stud_time_fixation/1000, 
                                  digits = 0),
         Group = as_factor(Group)) -> df_ttff_stud

# mean & SD time to first fixation "Disruptive Person"
# novices
mean_time_stud_nov <- 
  df_ttff_stud %>%
  filter(Group == "Novice") %>% 
  pull(Stud_time_fixation_sec) %>% 
  mean() %>% 
  round(., digits = 0)

sd_time_stud_nov <- 
  df_ttff_stud %>%
  filter(Group == "Novice") %>% 
  pull(Stud_time_fixation_sec) %>% 
  sd() %>% 
  round(., digits = 0)

# experts
mean_time_stud_exp <-
  df_ttff_stud %>%
  filter(Group == "Expert") %>% 
  pull(Stud_time_fixation_sec) %>% 
  mean() %>% 
  round(., digits = 0)

sd_time_stud_exp <- 
  df_ttff_stud %>%
  filter(Group == "Expert") %>% 
  pull(Stud_time_fixation_sec) %>% 
  sd() %>% 
  round(., digits = 0)

##################### plotting time to first fixation for AOI students ##################### 
plot_time_stud <- 
  ggplot(data = df_ttff_stud,
         mapping = aes(x = Group,
                       y = Stud_time_fixation_sec)) +
 geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  # scale_x_discrete(labels = paste0(c("Novice\n","Expert\n"),
  #                                  "M = ", c(mean_dur_nov, mean_dur_exp),
  #                                  "; SD = ", c(sd_dur_nov, sd_dur_exp)
  #                                  )
  #                  ) +
  # ylim(0,250) +
  labs(x ="",
       y = "Sekunden") + 
  scale_fill_brewer(palette  = "RdBu") +  
  scale_x_discrete(labels=c("Novice" = "Unerfahren",
                            "Expert" = "Erfahren")) +
  ggtitle("Zeit bis zur ersten Fixation\nAOI `Schüler:innen`") +
  theme_cowplot() +
  theme(legend.position="none",
        axis.text.x = element_text(size = 34),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 34),
        plot.title = element_text(size = 35, 
                                  face = "bold")
  )

plot_time_stud

# saving plot
ggsave(plot = plot_time_stud,
       filename = "plots/plot_time_stud.svg",
       height = 8,
       width = 8,
       units = "in")

# TIME TO FIRST FIXATIONS on AOI "Students"
# t-test for expertise differences
t.test(x = df_ttff_stud$Stud_time_fixation_sec[df_ttff_stud$Group == "Expert"],
       y = df_ttff_stud$Stud_time_fixation_sec[df_ttff_stud$Group == "Novice"],
       var.equal = TRUE)

# TIME TO FIRST FIXATIONS on AOI "Students" 
# effect size for expertise differences
d_time_stud <- CohenD(x = df_ttff_stud$Stud_time_fixation_sec[df_ttff_stud$Group == "Expert"],
                    y = df_ttff_stud$Stud_time_fixation_sec[df_ttff_stud$Group == "Novice"],
                    na.rm = TRUE)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.