### 2024-06-24 02_paper/02_study/02_fitbit paper/fitbit_paper_elsevier/FitBit_elsevier/FitBit_elsevier.Rmd
tabellen

---
title: "New approaches to teachers' experience of stress: Do heart rate measurements with fitness trackers provide an efficient, inexpensive, and robust measurement method?"
shorttitle: "Using fitbit to measure teachers' heart rate"
authornote: |
 We received funding from QualiFond of University Leipzig. We have no conflicts of interest to disclose. This article is based on data used at conference presentations (DACH-Nachwuchsakademie, 2022; EARLI SIG 11, 2022; EARLI SIG 27, 2022). 
author:
  - name: Mandy Klatt
    email: mandy.klatt@uni-leipzig.de
    affiliation: Leipzig University
    correspondingauthor: true
    footnote: 1
  - name: Peer Keßler
    email: Peer@example.com
    affiliation: Leipzig University
    footnote: 2
  - name: Gregor Kachel
    email: gregor@example.com
    affiliation: Leipzig University, Max Planck Institute for Evolutionary Anthropology
    footnote: 1
  - name: Christin Lotz
    email: christin@example.com
    affiliation: Leipzig University
    footnote: 1
  - name: Anne Deiglmayr
    email: anne@example.com
    affiliation: Leipzig University
    footnote: 3
address:
  - code: Leipzig University
    organization: Empirische Schul- und Unterrichtsforschung
    addressline: Marschnerstr. 29
    city: Leipzig
    state: Sachsen
    postcode: 04109
    country: Germany
  - code: Max Planck Institute for Evolutionary Anthropology
    organization: Department
    addressline: Deutscher Pl. 6
    postcode: 04103
    city: Leipzig
    state: Sachsen
    country: Germany
footnote:
  - code: 1
    text: 
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - code: 2
    text: 
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - code: 3
    text: "Supervision"
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  XXX In this proof-of-concept study, we aimed to advance the field of teacher stress by collecting heart rate data with wrist-worn devices and testing a methodology that has the potential to provide more insights on the non-invasive assessment of teacher stress. XXX

  <!-- https://tinyurl.com/ybremelq -->
keywords: 
  - heart rate
  - photoplethysmography
  - wearable electronic device
  - teaching
journal: "Computers & Education"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: r-references.bib
linenumbers: false
numbersections: true
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
header-includes:
  - \usepackage{tabularx}
---

```{r analysis-preferences, include=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r installing packages, include=FALSE}

library(needs)
needs(ltm,
      broom,
      ppcor,
      jtools,
      lm.beta,
      tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr,
      lme4,
      viridis, 
      stargazer)
```

```{r data wrangling, include=FALSE}
# disruption & confidence rating

df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~ read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  dplyr::select(ID, # select relevant columns 
         "disruption_factor", # -99 = subject did not notice event; -100 = experimenter did not asked for event
         "confidence_factor",
         "event"
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confidence_factor = ifelse(confidence_factor < 0,
                                   yes = NA,
                                   no = confidence_factor),
         event = as_factor(event)
         )

# demographic data

df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            gender = factor(LI02_01_1,
                            levels = 1:2,
                            labels = c("male","female")
                            ),
            age = LI03_01, # 1 = male; 2 = female
            teaching_experience = LI04_01)

# merge two data frames by ID

df_merge <- merge(df_demo,
                  df_rating,
                  by = "ID") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132")
         )

# rm(list = c("df_demo","df_rating"))

data_path <- "data/heart_rate_data"

main <-
  map(.x = dir(path = data_path,
               pattern = ".csv"),
      ~ read_csv(file.path(data_path, .), 
                 id = "id",
                 col_types = c("t","n")
                 ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "end",
                               str_detect(string = id, pattern = "_m") ~ "teaching",
                               str_detect(string = id, pattern = "_p") ~ "preparation",
                               str_detect(string = id, pattern = "_s") ~ "post",
                               str_detect(string = id, pattern = "_i") ~ "interview",
                               TRUE ~ "overall"
                               ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
    # filter(time <= 600) %>% # filter for 10min intervals
    filter(time <= 7200) %>% # filter for 2 hours (maximal duration of study)
    dplyr::select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

main <-
  left_join(main, df_merge,
            by = "ID")

main <- 
  main %>% 
  dplyr::select("time", "heart_rate", "ID", "time_span") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  dplyr::summarise(mean_heart = mean(heart_rate),
            sd_heart = sd(heart_rate)) %>%
  ungroup() %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  filter(!(time_span == "teaching" & time > 600)) %>%
  mutate(heart_rate_std = (heart_rate - mean_heart)/
           sd_heart)
```

```{r descriptives, include=FALSE}
# Mean, SD and range for overall (unstandardized)

descrip_hr_overall <- main %>%
  filter(time_span %in% c("overall")) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate),
                                      digits = 2),
            "Min heart rate in bpm" = min(heart_rate),
            "Max heart rate in bpm" = max(heart_rate)
  )

# Mean, SD and range for overall (standardized)

descrip_hr_overall_std <- main %>%
  filter(time_span %in% c("overall")) %>% 
  summarise(N = n_distinct(ID),
            "M std heart rate in bpm" = round(mean(heart_rate_std),
                                              digits = 2),
            "SD std heart rate in bpm" = round(sd(heart_rate_std),
                                               digits = 2),
            "Min std heart rate in bpm" = round(min(heart_rate_std),
                                                digits = 2),
            "Max heart rate in bpm" = round(max(heart_rate_std),
                                            digits = 2)
            )

# Mean, SD and range for individual phases (unstandardized)

descrip_hr_phases <- main %>% 
  filter(!time_span %in% c("overall")) %>% 
  group_by(time_span) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate),
                                      digits = 2),
            "Min heart rate in bpm" = min(heart_rate),
            "Max heart rate in bpm" = max(heart_rate)
  )

# Mean, SD and range for individual phases (standardized)

descrip_hr_phases_std <- main %>% 
  filter(!time_span %in% c("overall")) %>% 
  group_by(time_span) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate_std),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate_std),
                                      digits = 2),
            "Min heart rate in bpm" = round(min(heart_rate_std),
                                            digits = 2),
            "Max heart rate in bpm" = round(max(heart_rate_std),
                                            digits = 2)
  )

# # format and insert table in manuscript
# descrip_hr_phases %>%
# papaja::apa_table(
#   caption = "Mean, SD and range for teachers' heart rate in individual phases",
#   # note = "Write Note here",
#   escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
#   placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
#   font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
#   )

# Mean, SD and range for disruption and confidence factor

rating_table <- 
  df_rating %>% 
  filter(!is.na(confidence_factor),
         !is.na(disruption_factor)) %>% 
   summarise(N = n_distinct(ID), 
            "M disruption factor" = round(mean(disruption_factor),
                                     digits = 2),
            "SD disruption factor" = round(sd(disruption_factor),
                                      digits = 2),
            "Min disruption factor" = min(disruption_factor),
            "Max disruption factor" = max(disruption_factor),
            "M confidence factor" = round(mean(confidence_factor),
                                     digits = 2),
            "SD confidence factor" = round(sd(confidence_factor),
                                      digits = 2),
            "Min confidence factor" = min(confidence_factor),
            "Max confidence factor" = max(confidence_factor)
            )
```

```{r, include=FALSE}

desc_hr_phases <- bind_rows(x = descrip_hr_overall,
                            y = descrip_hr_phases) |> 
  select(-N) |> 
  rename(`M HR` = `M heart rate in bpm`,
         `SD HR` = `SD heart rate in bpm`,
         `Min` = `Min heart rate in bpm`,
         `Max` = `Max heart rate in bpm`,
         Interval = time_span) |> 
  select(c(Interval,
           `M HR`,
           `SD HR`,
           `Min`,
           `Max`)
         )

descrip_hr_overall_std <- descrip_hr_overall_std |> 
  rename(`M HR` = `M std heart rate in bpm`,
         `SD HR` = `SD std heart rate in bpm`,
         `Min` = `Min std heart rate in bpm`,
         `Max` = `Max heart rate in bpm`,
          ) |> 
  mutate(Interval = 'overall')

descrip_hr_phases_std <-  descrip_hr_phases_std |> 
  rename(`M HR` = `M heart rate in bpm`,
         `SD HR` = `SD heart rate in bpm`,
         `Min` = `Min heart rate in bpm`,
         `Max` = `Max heart rate in bpm`,
         Interval = time_span)

desc_hr_phases_std <- bind_rows(x = descrip_hr_overall_std,
                                y = descrip_hr_phases_std) |> 
  select(-N) |> 
    select(c(Interval,
           `M HR`,
           `SD HR`,
           `Min`,
           `Max`)
         )

desc_hr_phases[1, 1] <- 'overall'

df_desc_hr <- desc_hr_phases |> 
  mutate(across(everything(),
                ~ paste(.,
                        desc_hr_phases_std[[cur_column()]],
                        sep = "/")
                )
         ) #|> 
  # # rename(`\textit{M} HR` = `M HR`,
  #        # `\textit{SD} HR` = `SD HR`,
  #        # `\textit{Min}` = `Min`,
  #        `\textit{Max}` = `Max`,
  #       `\textit{Interval}` = Interval)

df_desc_hr$Interval <- c('Overall Course of 2h',
                         'Pre-teaching interval',
                         'Teaching interval',
                         'Post-teaching interval',
                         'Interview interval',
                         'End Interval')

names(df_desc_hr) <- paste0("\\textit{", names(df_desc_hr),"}")

stargazer(df_desc_hr,
          type = 'latex',
          out = 'tab_1.tex',
          summary = FALSE,
          rownames = FALSE,
          table.placement = 'h',
          align = FALSE
          ) -> tab_1

pattern <- "textit\\\\\\{([^}]*)\\\\\\}"
replacement <- "\textit{\\1}"

# Anwenden des Regex
tab_1_new <- str_replace_all(tab_1, pattern, replacement)

paste(tab_1_new, collapse = '')

writeLines(tab_1_new, 'tab_1.tex')

```

```{=latex}
\input{tab_1.tex}
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

# Present investigation 
The aim was to investigate whether HR measures assessed by a Fitbit Charge 4 are a suitable and effective method to **(1)** map teachers course of arousal over the course of a five phase lab study, including a micro teaching unit,  and **(2)**  examine whether HR measures can be predicted by self-reported data on cognitive appraisal.

Within the time frame of approximately two hours, we investigated five intervals with a duration of 10-minutes each: In the (1) pre-teaching phase, the subjects were prepared for the following micro teaching unit  teaching lesson and familiarized with the setting. During the (2) teaching phase, the participants taught a 15-minute self-prepared lesson to a "class" of three actors that simulated nine classroom disruptions . In the following (3) post-teaching phase, the subjects answered a questionnaires, followed by  and in the (4) interview phase, in which they subjects watched thea pre-recorded video of their 15-minute lesson to assess the self-report data of how disrupted subjects felt and how confident they felt in dealing with disruptions. In the (5) end phase, the subject answered another  second questionnaire.

According to previous findings that fitness trackers can be used as a low-cost, non-invasive method of measuring HR [@hajj2022wrist; @fuller2020reliability] and that different HRs of teachers can be measured in different teaching phases [@donker2020associations; @junker2021potential], we formulate the hypotheses as follows:

**Hypothesis 1**. In a first step, we wanted to display exploratively  the course of HR during the entire study.  Additionally, we presumed the highest mean HR in the (2) teaching phase and lower mean values in all other phases (**Hypothesis 1a**). Moreover, we expected an increase in HR in the (1) pre-teaching phase as the first phase and a decrease in the following phases (**Hypothesis 1b**).
**Hypothesis 2**. We statistically predicted  the subjects’ standardized mean HR for the (2) teaching, the (3) post-teaching, the (4) interview and the (5) end phase with teaching experience and self-repoted data on ….  With respect to teaching experience, Wwe expected a lower HR in teachers with more teaching experience for the four phases (**Hypothesis 2a**). According to the relationship between physiological arousal and cognitive appraisal, we controlled for shared variance with the self-reported data. Concerning the …, Wwe expected higher HR values for for the four phases that teachers who reported that they felt more disrupted by disruptions would have higher standardized mean HR  (**Hypotheses 2b**). In contrast, individuals with high confidence in dealing with disruptions would have a lower standardized mean HR in the four phases (**Hypothesis 2c**). When considering the three predictors in concert and controlling for their common variance, we expected teaching experience and self-reported data to remain substantial predictors (**Hypothesis 2d**).

# References {-}