### 2024-10-18 02_paper/02_study/01_fitbit_paper/Submission/tab_1.tex

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:table_1}Mean HR (M), standard deviations HR (SD), and range of teachers’ HR over the course of the entire study and the five intervals (unstandardized in BPM/z-standardized)}

\begin{tabular}{lllll}
\toprule
Interval & \multicolumn{1}{c}{M HR} & \multicolumn{1}{c}{SD HR} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max}\\
\midrule
Overall Course of 2h & 90.09/-0.04 & 15.76/0.99 & 51/-4.03 & 164/4.56\\
Pre-teaching interval & 77.95/-1.07 & 11.14/0.57 & 50/-2.68 & 120/2.96\\
Teaching interval & 82.32/-0.72 & 11.85/0.74 & 51/-2.51 & 132/4.39\\
Post-teaching interval & 93.61/0.27 & 14.02/0.76 & 60/-2.17 & 150/3.06\\
Interview interval & 96.28/0.48 & 14.11/0.88 & 56/-3.56 & 139/3.24\\
End Interval & 100.8/0.85 & 16.24/0.77 & 63/-2.18 & 164/4.37\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Please note that M and SD of the overall course vs. the individual intervals were not exactly 0 and 1 due to rounding differences.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}