### 2024-10-18 02_paper/02_study/01_fitbit_paper/Submission/tab_2.tex

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-2}Descriptive statistics (n, M, SD) for the mean intercepts and the mean slopes for the five intervals}

\begin{tabular}{llllllll}
\toprule
Phase & \multicolumn{1}{c}{n} & \multicolumn{1}{c}{Mean(Intercept)} & \multicolumn{1}{c}{SD(Intercept)} & \multicolumn{1}{c}{p-Value (Intercept)} & \multicolumn{1}{c}{Mean(Slope)} & \multicolumn{1}{c}{SD(Slope)} & \multicolumn{1}{c}{p-Value (Slope)}\\
\midrule
Pre Teaching Phase & 6896 & 0.05 & 0.82 & 0.57 & 0.08 & 0.13 & 0.00\\
Teaching Phase & 7150 & 1.02 & 0.69 & 0.00 & -0.04 & 0.11 & 0.00\\
Post Teaching Phase & 6664 & 0.55 & 0.55 & 0.00 & -0.06 & 0.10 & 0.00\\
Interview Phase & 6287 & -0.62 & 0.61 & 0.00 & -0.02 & 0.07 & 0.01\\
End Phase & 5990 & -1.00 & 0.50 & 0.00 & -0.01 & 0.07 & 0.14\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Please note that M and SD of the overall course vs. the individual intervals were not exactly 0 and 1 due to rounding differences.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}