### 2024-11-14 02_paper/02_study/02_expertise_paper/expertise_2024_09_26_no_outlierdetection_MK.Rmd
x paper update

---
title: "Through the Eyes of the Teacher - Multimodal Exploration of Expertise Differences in the Perception of Classroom Disruptions in a Laboratory Study"
output: 
  html_document:
    toc: true           # Add a table of contents
    number_sections: true
---

```{r include=FALSE}
# install.packages("needs")

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

library(needs)
needs(tidyverse,
      lubridate,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable,
      DescTools,
      formattable,
      psych,
      apaTables,
      rstatix,
      effsize,
      knitr
      )

```

## Participants
```{r demographicstable, include = TRUE, echo=FALSE}

# Read and process the data
demo_data <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  filter(!LI06_05 %in% c(201, 223)) %>%
  transmute(
    Group = ifelse(LI06_05 < 200, "Novice", "Expert"),
    Gender = as.numeric(ifelse(LI02_01_1 == 2, "1", "0")),
    Age = LI03_01,
    `Teaching Experience` = LI04_01,
    School_type = factor(
      replace(LI18, LI18 == 2, 1),
      levels = c(1, 3, 4, 5, -1),
      labels = c("Primary School", "Secondary School", "Special Education", "Vocational School", "Other School Type")
    ),
    Semester = as.numeric(LI11_01),
    Internship = LI13_01,
    `Extracurricular Teaching Experience` = as.numeric(ifelse(LI14 == 1, "1", "0")),
    `Secondary Teaching Activities` = as.numeric(ifelse(LI17 == 1, "1", "0"))
  )

# Calculate summaries used across both tables, handling cases where all values are NA
summary_data <- demo_data %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    `Women in %` = ifelse(all(is.na(Gender)), NA, round(mean(Gender, na.rm = TRUE) * 100, 2)),
    `M Age` = ifelse(all(is.na(Age)), NA, round(mean(Age, na.rm = TRUE), 2)),
    `SD Age` = ifelse(all(is.na(Age)), NA, round(sd(Age, na.rm = TRUE), 2)),
    `Min Age` = ifelse(all(is.na(Age)), NA, min(Age, na.rm = TRUE)),
    `Max Age` = ifelse(all(is.na(Age)), NA, max(Age, na.rm = TRUE)),
    `M Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, round(mean(`Teaching Experience`, na.rm = TRUE), 2)),
    `SD Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, round(sd(`Teaching Experience`, na.rm = TRUE), 2)),
    `Median Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, median(`Teaching Experience`, na.rm = TRUE)),
    `Min Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, min(`Teaching Experience`, na.rm = TRUE)),
    `Max Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, max(`Teaching Experience`, na.rm = TRUE)),
    `M Semester` = ifelse(all(is.na(Semester)), NA, round(mean(Semester, na.rm = TRUE), 2)),
    `SD Semester` = ifelse(all(is.na(Semester)), NA, round(sd(Semester, na.rm = TRUE), 2)),
    `Min Semester` = ifelse(all(is.na(Semester)), NA, min(Semester, na.rm = TRUE)),
    `Max Semester` = ifelse(all(is.na(Semester)), NA, max(Semester, na.rm = TRUE)),
    `M Internship` = ifelse(all(is.na(Internship)), NA, round(mean(Internship, na.rm = TRUE), 2)),
    `SD Internship` = ifelse(all(is.na(Internship)), NA, round(sd(Internship, na.rm = TRUE), 2)),
    `Min Internship` = ifelse(all(is.na(Internship)), NA, min(Internship, na.rm = TRUE)),
    `Max Internship` = ifelse(all(is.na(Internship)), NA, max(Internship, na.rm = TRUE)),
    `Extracurricular Teaching Experience %` = ifelse(all(is.na(`Extracurricular Teaching Experience`)), NA, round(mean(`Extracurricular Teaching Experience`, na.rm = TRUE) * 100, 2)),
    `Secondary Teaching Activities %` = ifelse(all(is.na(`Secondary Teaching Activities`)), NA, round(mean(`Secondary Teaching Activities`, na.rm = TRUE) * 100, 2))
  )

# Separate the tables from the summarized data
demo_table <- summary_data %>%
  select(Group, N, `Women in %`, `M Age`, `SD Age`, `Min Age`, `Max Age`, 
         `M Teaching Exp.`, `SD Teaching Exp.`, `Median Teaching Exp.`, 
         `Min Teaching Exp.`, `Max Teaching Exp.`)

exp_table <- summary_data %>%
  select(Group, N, `M Teaching Exp.`, `SD Teaching Exp.`, `Median Teaching Exp.`, 
         `Min Teaching Exp.`, `Max Teaching Exp.`, `M Semester`, `SD Semester`, 
         `Min Semester`, `Max Semester`, `M Internship`, `SD Internship`, 
         `Min Internship`, `Max Internship`, `Extracurricular Teaching Experience %`, 
         `Secondary Teaching Activities %`)

# Display APA-style HTML tables
demo_table %>%
  kbl(caption = "Demographic Information & Teaching Experience", format = "html") %>%
  kable_styling(full_width = FALSE, position = "center")

exp_table %>%
  kbl(caption = "Experience Details by Group", format = "html") %>%
  kable_styling(full_width = FALSE, position = "center") %>% 
  scroll_box(width = "100%", height = "400px") # Enables scrolling if needed

```

## Measures

### Eye-Tracking Data
```{r eyetracking_read_in_data, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

# Step 1: List all .tsv files in the "data" folder that contain "interval_complete"
file_names <- list.files(path = "data", pattern = "interval_complete.tsv", full.names = TRUE)

# Step 2: Read and bind all .tsv files into a single tibble
df_aoi <- file_names %>%
  map_dfr(~ read_tsv(.x, locale = locale(decimal_mark = ",")) %>%
            select(
              Group,
              Participant,
              TOI,
              Duration_of_interval,
              Start_of_interval,
              starts_with("Total_duration_of_fixations"),
              starts_with("Average_duration_of_fixations"),
              starts_with("Number_of_fixations"),
              starts_with("Time_to_first_fixation")
            ))

# Step 3: Extract only the participant's ID number (3 digits)
df_aoi <- df_aoi %>%
  mutate(Participant = as.numeric(str_extract(Participant, "\\d{3}")))

# Step 4: Exclude invalid participant ID 223
df_aoi <- df_aoi %>%
  filter(Participant != 223)

```
### Check variables

#### Letter search
```{r letter_search, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### LETTER SEARCH #####################
# 1. Filter and Prepare Data for "Letter Search"
df_letter <- df_aoi %>%
  filter(TOI == "Letter_search", !Participant %in% c(111, 205)) %>%  # Exclude participants with IDs 111 and 205
  select(TOI, Group, Participant, Duration_of_interval) %>%
  mutate(Duration_of_interval_sec = round(Duration_of_interval / 1000, 2))

# 2. Descriptive Statistics: N, Mean (M), SD, Min, Max for "Letter Search" Duration
letter_table <- df_letter %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Duration_of_interval_sec), 2),
    SD = round(sd(Duration_of_interval_sec), 2),
    Min = round(min(Duration_of_interval_sec), 2),
    Max = round(max(Duration_of_interval_sec), 2)
  )

# Display Descriptive Statistics Table in APA Style
knitr::kable(letter_table, caption = "N, M, SD, Min & Max Letter Search Duration (in seconds)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Conduct t-Test and Calculate Cohen's d for "Letter Search" Duration
t_test_letter <- t.test(
  Duration_of_interval_sec ~ Group,
  data = df_letter,
  var.equal = TRUE
)

# Calculate Cohen's d for the Effect Size
d_letter <- cohen.d(
  Duration_of_interval_sec ~ Group,
  data = df_letter
)

# 4. Format t-Test and Cohen's d Results for APA Table
t_test_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_letter$statistic, 2),
  "df" = t_test_letter$parameter,
  "p-value" = format.pval(t_test_letter$p.value, digits = 3, eps = .001),
  "Mean Difference" = round(t_test_letter$estimate[1] - t_test_letter$estimate[2], 2),
  "Cohen's d" = round(d_letter$estimate, 2),
  "95% CI (d)" = paste0("[", round(d_letter$conf.int[1], 2), ", ", round(d_letter$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size Results in APA Style
knitr::kable(t_test_result, caption = "t-Test and Effect Size for Letter Search Duration (Experts vs. Novices)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Percentage of fixations on each AOI, relative to the total number of fixations during the entire "Lesson" 
```{r nof_percent, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Step 1: Define AOI categories
aoi_columns <- list(
  "AOI_Students" = c(
    "Number_of_fixations.Anna",
    "Number_of_fixations.Bianca",
    "Number_of_fixations.Carl(a)"
  ),
  "AOI_Disruptive_Person" = c("Number_of_fixations.Disruptive_Person"),
  "AOI_Teacher_Material" = c(
    "Number_of_fixations.Board_Screen",
    "Number_of_fixations.Material_Teacher"
  ),
  "AOI_Student_Desk" = c(
    "Number_of_fixations.Nametag_Anna",
    "Number_of_fixations.Nametag_Bianca",
    "Number_of_fixations.Nametag_Carl(a)",
    "Number_of_fixations.Material_Students"
  ),
  "AOI_Classroom_Others" = c("Number_of_fixations.Classroom_Others")
)

# Step 2: Filter data for "Lesson" and select relevant columns
df_fixations <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(Participant, Group, starts_with("Number_of_fixations"))

# Step 3: Calculate total fixations per participant, excluding "Disruptive Person"
# Extract relevant column names excluding "Disruptive Person"
relevant_columns <- unlist(aoi_columns[c("AOI_Students", "AOI_Teacher_Material", "AOI_Student_Desk", "AOI_Classroom_Others")])

# Calculate row sums for the relevant columns and add as a new column
df_fixations <- df_fixations %>%
  mutate(Total_fixations_excluding_disruptive = rowSums(across(all_of(relevant_columns)), na.rm = TRUE))

# Step 4: Calculate fixation percentages for each AOI based on the adjusted total
aoi_percentages <- df_fixations %>%
  rowwise() %>%
  mutate(
    AOI_Students_pct = sum(c_across(any_of(
      aoi_columns$AOI_Students
    )), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100,
    AOI_Disruptive_Person_pct = sum(c_across(
      any_of(aoi_columns$AOI_Disruptive_Person)
    ), na.rm = TRUE) / (
      Total_fixations_excluding_disruptive + sum(c_across(
        any_of(aoi_columns$AOI_Disruptive_Person)
      ), na.rm = TRUE)
    ) * 100,
    AOI_Teacher_Material_pct = sum(c_across(
      any_of(aoi_columns$AOI_Teacher_Material)
    ), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100,
    AOI_Student_Desk_pct = sum(c_across(any_of(
      aoi_columns$AOI_Student_Desk
    )), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100,
    AOI_Classroom_Others_pct = sum(c_across(
      any_of(aoi_columns$AOI_Classroom_Others)
    ), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100
  ) %>%
  ungroup()

# Step 5: Summarize fixation percentages by Group and arrange in descending order
grouped_aoi_percentages <- aoi_percentages %>%
  group_by(Group) %>%
  summarise(
    Mean_AOI_Students_pct = round(mean(AOI_Students_pct, na.rm = TRUE), 2),
    Mean_AOI_Disruptive_Person_pct = round(mean(AOI_Disruptive_Person_pct, na.rm = TRUE), 2),
    Mean_AOI_Teacher_Material_pct = round(mean(AOI_Teacher_Material_pct, na.rm = TRUE), 2),
    Mean_AOI_Student_Desk_pct = round(mean(AOI_Student_Desk_pct, na.rm = TRUE), 2),
    Mean_AOI_Classroom_Others_pct = round(mean(AOI_Classroom_Others_pct, na.rm = TRUE), 2)
  ) %>%
  pivot_longer(cols = starts_with("Mean_AOI"), names_to = "AOI", values_to = "Fixation_Percentage") %>%
  mutate(AOI = recode(AOI,
                      "Mean_AOI_Students_pct" = "Students",
                      "Mean_AOI_Disruptive_Person_pct" = "Disruptive Person",
                      "Mean_AOI_Teacher_Material_pct" = "Teacher Material",
                      "Mean_AOI_Student_Desk_pct" = "Student Desk",
                      "Mean_AOI_Classroom_Others_pct" = "Classroom/Others")) %>%
  arrange(Group, desc(Fixation_Percentage))

# Step 6: Display the table with APA style
kable(grouped_aoi_percentages, 
      col.names = c("AOI", "Expert (%)", "Novice (%)"),
      caption = "Average Fixation Percentage per AOI by Group (Novice vs. Expert) - Descending Order") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Fixation Percentage" = 2)) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(2:3, width = "3em", border_left = TRUE)

# Step 7: Plot with descending AOI order, y-axis limit set to 100%, and percentage labels inside the bars
ggplot(grouped_aoi_percentages, aes(x = reorder(AOI, -Fixation_Percentage), y = Fixation_Percentage, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = paste0(Fixation_Percentage, "%")), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3, size = 3) +
  labs(
    title = "Average Fixation Percentage by AOI and Group (Descending Order)",
    x = "AOI",
    y = "Fixation Percentage (%)"
  ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(title = "Group")) +
  ylim(0, 100)  # Set y-axis limit to 100%

# Step 8: Perform a t-test between Experts and Novices on the percentage values and calculate Cohen's d
# Define the AOIs you want to analyze with names matching those generated in Step 4
aoi_list <- c("AOI_Students_pct", "AOI_Disruptive_Person_pct", "AOI_Teacher_Material_pct", "AOI_Student_Desk_pct", "AOI_Classroom_Others_pct")

# Initialize an empty data frame to store the results
results <- data.frame()

# Loop over each AOI and calculate t-test and Cohen's d
for (column_name in aoi_list) {

  # Ensure that the column exists in the data
  if (column_name %in% names(aoi_percentages)) {

    # Filter the data for the current AOI and rename the column to Fixation_Percentage for consistency
    aoi_data <- aoi_percentages %>%
      select(Group, !!sym(column_name)) %>%
      rename(Fixation_Percentage = !!sym(column_name))

    # Separate data for each group
    expert_data <- aoi_data$Fixation_Percentage[aoi_data$Group == "Expert"]
    novice_data <- aoi_data$Fixation_Percentage[aoi_data$Group == "Novice"]

    # Check if both groups have at least one observation
    if (length(expert_data) > 0 && length(novice_data) > 0) {

      # Perform a t-test between Experts and Novices on the current AOI's fixation percentage
      t_test_result <- t.test(expert_data, novice_data, var.equal = TRUE)

      # Calculate Cohen's d for the current AOI
      cohen_d_value <- cohen.d(expert_data, novice_data, pooled = TRUE)$estimate

      # Format the results for the current AOI and add to the results data frame
      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),  # Remove "_pct" suffix for readability
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = round(t_test_result$statistic, 2),
          "df" = t_test_result$parameter,
          "p-value" = format.pval(t_test_result$p.value, digits = 3, eps = .001),
          "Mean Difference" = round(t_test_result$estimate[1] - t_test_result$estimate[2], 2),
          "Cohen's d" = round(cohen_d_value, 2),
          "95% CI (d)" = paste0("[", round(t_test_result$conf.int[1], 2), ", ", round(t_test_result$conf.int[2], 2), "]")
        )
      )

    } else {
      # If one group has no data, append N/A values for this AOI
      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),  # Remove "_pct" suffix for readability
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = NA,
          "df" = NA,
          "p-value" = NA,
          "Mean Difference" = NA,
          "Cohen's d" = NA,
          "95% CI (d)" = NA
        )
      )
    }
  }
}

# Clean up AOI names for readability in the final table
aoi_name_mapping <- c(
  "AOI_Students" = "Students",
  "AOI_Disruptive_Person" = "Disruptive Person",
  "AOI_Teacher_Material" = "Teacher Material",
  "AOI_Student_Desk" = "Student Desk",
  "AOI_Classroom_Others" = "Classroom/Others"
)
results$AOI <- aoi_name_mapping[results$AOI]

# Display the combined results table in APA style without row names
knitr::kable(results, caption = "t-Test and Effect Size for Fixation Percentage across AOIs (Experts vs. Novices)", row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Number of fixations per minute (micro-teaching unit)
```{r nof_all, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATION IN MICRO-TEACHING UNIT ########################

# 1. Data Preparation: Filter and Calculate Fixations per Minute in Micro-Teaching Unit
df_aoi_sum <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Group,
    Participant,
    Duration_of_interval,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations"),
    starts_with("Average_duration"),
    -contains("Disruptive_Person")
  ) %>%
  rowwise() %>%
  transmute(
    Group = Group,
    Participant = Participant,
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Sum_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Sum_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Number_fixation_min_mtu = round(Sum_number_fixation / Duration_of_interval_min, 0),
    Group = as_factor(Group)
  ) %>%
  drop_na()

# 2. Descriptive Statistics Table
nof_table <- df_aoi_sum %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Number_fixation_min_mtu), 2),
    SD = round(sd(Number_fixation_min_mtu), 2),
    Min = round(min(Number_fixation_min_mtu), 2),
    Max = round(max(Number_fixation_min_mtu), 2)
  )

# Display the descriptive table in APA style
knitr::kable(nof_table, caption = "N, M, SD, min & max number of fixation per minute (micro-teaching unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Calculate means and standard deviations for Novice and Expert
mean_novice_mtu <- mean(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Novice"], na.rm = TRUE)
sd_novice_mtu <- sd(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Novice"], na.rm = TRUE)
mean_expert_mtu <- mean(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Expert"], na.rm = TRUE)
sd_expert_mtu <- sd(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Expert"], na.rm = TRUE)

# 4. Perform t-test to get the p-value
t_test_mtu <- t.test(
  Number_fixation_min_mtu ~ Group,
  data = df_aoi_sum,
  var.equal = TRUE
)
p_value_mtu <- round(t_test_mtu$p.value, 3)

# 5. Calculate Cohen's d for effect size using formula syntax
library(effsize)
cohen_d_mtu <- round(cohen.d(Number_fixation_min_mtu ~ Group, data = df_aoi_sum)$estimate, 2)

# 6. Plotting the Number of Fixations per Minute (Micro-Teaching Unit)
# Set y-limit based on data to provide space for annotations
y_max_mtu <- max(df_aoi_sum$Number_fixation_min_mtu) * 1.5

plot_number_group <- df_aoi_sum %>%
  ggplot(mapping = aes(x = Group, y = Number_fixation_min_mtu)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_mtu) +
  labs(x = "", y = "Number of fixations per minute") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Number of fixations per minute\n(micro-teaching unit)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  # Add annotations for mean (M) and SD using atop() to place SD below M
  annotate("text", x = 1, y = y_max_mtu * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice_mtu, ", italic(SD) == ", sd_novice_mtu, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max_mtu * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert_mtu, ", italic(SD) == ", sd_expert_mtu, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Add p-value and Cohen's d annotations above the boxplots
  annotate("text", x = 1.5, y = y_max_mtu * 0.95,
           label = paste0("p = ", p_value_mtu, "\nd = ", cohen_d_mtu),
           hjust = 0.5, size = 4, fontface = "italic")

plot_number_group

# 7. Format the t-Test and Cohen's d results for APA Table
t_test_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_mtu$statistic, 2),
  "df" = t_test_mtu$parameter,
  "p-value" = format.pval(t_test_mtu$p.value, digits = 3, eps = .001),
  "Mean Difference" = round(t_test_mtu$estimate[1] - t_test_mtu$estimate[2], 2),
  "Cohen's d" = cohen_d_mtu,
  "95% CI (d)" = paste0("[", round(t_test_mtu$conf.int[1], 2), ", ", round(t_test_mtu$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size results in APA style
knitr::kable(t_test_result, caption = "t-Test and Effect Size for Number of Fixations per Minute (Micro-Teaching Unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Number of fixations per minute (AOI students)
```{r nof_students, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATIONS (AOI STUDENTS) ########################

# 1. Data Preparation: Filter and Calculate Fixations per Minute on AOI Students
df_aoi_stud <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  select(
    Group,
    Duration_of_interval,
    Participant,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations")
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Participant = Participant,
    Stud_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Stud_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Stud_number_fixation_min = round(Stud_number_fixation / Duration_of_interval_min, 2),
    Stud_number_fixation_sec = round(Stud_number_fixation / Duration_of_interval_sec, 2),
    Average_duration_stud = round(Stud_duration_fixation / Stud_number_fixation, 0)
  ) %>%
  drop_na()

# 2. Descriptive Statistics: N, Mean (M), SD, Min, Max for Fixations per Minute (AOI Students)
nof_min_stud_table <- df_aoi_stud %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Stud_number_fixation_min), 2),
    SD = round(sd(Stud_number_fixation_min), 2),
    Min = round(min(Stud_number_fixation_min), 2),
    Max = round(max(Stud_number_fixation_min), 2)
  )

# Display Descriptive Statistics Table in APA Style
knitr::kable(nof_min_stud_table, caption = "N, M, SD, Min & Max Number of Fixations per Minute (AOI Students)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Plotting the Number of Fixations per Minute on AOI Students
# Calculate descriptive statistics
summary_stats <- df_aoi_stud %>%
  group_by(Group) %>%
  summarise(
    M = round(mean(Stud_number_fixation_min), 2),
    SD = round(sd(Stud_number_fixation_min), 2)
  )

# Conduct t-test and calculate effect size
t_test_stud <- t.test(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud,
  var.equal = TRUE
)

# Cohen's d for effect size
d_number_stud <- cohen.d(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud
)

# Extract statistical values
mean_expert <- summary_stats %>% filter(Group == "Expert") %>% pull(M)
sd_expert <- summary_stats %>% filter(Group == "Expert") %>% pull(SD)
mean_novice <- summary_stats %>% filter(Group == "Novice") %>% pull(M)
sd_novice <- summary_stats %>% filter(Group == "Novice") %>% pull(SD)
p_value <- format.pval(t_test_stud$p.value, digits = 3, eps = .001)
cohen_d <- round(d_number_stud$estimate, 2)

# Determine max value for ylim to ensure all text is visible
y_max <- max(df_aoi_stud$Stud_number_fixation_min) * 1.5  # Increase this multiplier if needed

# Plot with adjusted ylim and annotations for readability
plot_number_min_stud <- df_aoi_stud %>%
  ggplot(mapping = aes(x = Group, y = Stud_number_fixation_min)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  labs(x = "", y = "Number of fixations per minute") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Number of fixations per minute\n(AOI students)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  ylim(0, y_max) +  # Set y-axis limit to ensure space for annotations
  # Adjusted annotations for mean (M) and SD using atop() to place SD below M
  annotate("text", x = 1, y = y_max * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice, ", italic(SD) == ", sd_novice, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert, ", italic(SD) == ", sd_expert, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Place p-value and d above the boxplots
  annotate("text", x = 1.5, y = y_max * 0.95,
           label = paste0("p = ", p_value, "\nd = ", cohen_d),
           hjust = 0.5, size = 4, fontface = "italic")

plot_number_min_stud

# 4. t-Test and Effect Size Calculation for Number of Fixations per Minute on AOI Students
t_test_stud <- t.test(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud,
  var.equal = TRUE
)

# Calculate Cohen's d for the Effect Size
d_number_stud <- cohen.d(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud
)

# 5. Format t-Test and Cohen's d Results for APA Table
t_test_result_stud <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_stud$statistic, 2),
  "df" = t_test_stud$parameter,
  "p-value" = format.pval(t_test_stud$p.value, digits = 3, eps = .001),
  "Mean Difference" = round(t_test_stud$estimate[1] - t_test_stud$estimate[2], 2),
  "Cohen's d" = round(d_number_stud$estimate, 2),
  "95% CI (d)" = paste0("[", round(d_number_stud$conf.int[1], 2), ", ", round(d_number_stud$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size Results in APA Style
knitr::kable(t_test_result_stud, caption = "t-Test and Effect Size for Number of Fixations per Minute (AOI Students)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Number of fixations per seconds (AOI disruptive person)
```{r nof_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATION ON AOI DISRUPTIVE PERSON ########################

# Define the list of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")

# Define categories for each type of disruption
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Filter data to include only the specified disruptions
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%
  dplyr::select(Group, Participant, TOI, Duration_of_interval, 
                "Number_of_fixations.Disruptive_Person") %>%
  mutate(Group = as_factor(Group))

# Add a column for Disruption_Category based on TOI values
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) # Keep only rows with valid category

# Convert duration from milliseconds to seconds
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(Duration_of_interval_sec = Duration_of_interval / 1000)

# Summarize total duration in seconds and fixation counts for "Disruptive_Person" by group, participant, and category
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarize(
    total_duration_sec = sum(Duration_of_interval_sec, na.rm = TRUE),
    fixation_count = sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(fixations_per_second = fixation_count / total_duration_sec)

# Step 1: Aggregate disruptions on an individual level (one value per participant)
# Calculate the mean fixation rate per participant
individual_summary <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarize(
    mean_fixations_per_second = mean(fixations_per_second, na.rm = TRUE),
    .groups = "drop"
  )

# First Plot: Overall fixation rate by Group (using aggregated individual data)
plot_overall <- individual_summary %>%
  ggplot(mapping = aes(x = Group, y = mean_fixations_per_second)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  ylim (0, 1) + 
  labs(x = "", y = "Mean Fixations per Second per Participant") +
  ggtitle("Overall Fixation Rate per Participant\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# Second Plot: Fixation rate by Group and Disruption_Category (no aggregation, shows all data points)
plot_category <- disruption_summary %>%
  ggplot(mapping = aes(x = Group, y = fixations_per_second, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) + # Wrap long facet labels
  labs(x = "", y = "Fixations per Second") +
  ggtitle("Fixation Rate by Disruption Category\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10) # Adjust facet text size if necessary
  )

# Display both plots
plot_overall
plot_category

# Function to format t-test and Cohen's d results into a tidy data frame, with checks for non-numeric values
format_results <- function(t_test, cohen_d) {
  data.frame(
    "t-value" = if (is.numeric(t_test$statistic)) round(t_test$statistic, 2) else NA,
    "df" = if (is.numeric(t_test$parameter)) round(t_test$parameter, 2) else NA,
    "p-value" = if (!is.null(t_test$p.value)) format.pval(t_test$p.value, digits = 3, eps = .001) else "NA",
    "Mean Difference" = if (is.numeric(t_test$estimate[1] - t_test$estimate[2])) round(t_test$estimate[1] - t_test$estimate[2], 2) else NA,
    "Cohen's d" = if (is.numeric(cohen_d$estimate)) round(cohen_d$estimate, 2) else NA,
    "95% CI (d)" = if (!is.null(cohen_d$conf.int)) 
                    paste0("[", round(cohen_d$conf.int[1], 2), ", ", round(cohen_d$conf.int[2], 2), "]") 
                   else "NA"
  )
}

# 1. t-Test and Cohen's d for fixations per second between Experts and Novices
t_fixations_group <- t.test(fixations_per_second ~ Group, data = disruption_summary)
d_fixations_group <- cohen.d(fixations_per_second ~ Group, data = disruption_summary)
fixations_group_result <- format_results(t_fixations_group, d_fixations_group)  # Formatting results

# 2. t-Test and Cohen's d for fixations per second by category
# Iterate over each category and calculate t-test and Cohen's d
categories <- unique(disruption_summary$Disruption_Category)
fixations_category_results <- list()

for (cat in categories) {
  cat_data <- disruption_summary %>% filter(Disruption_Category == cat)
  t_test_cat <- t.test(fixations_per_second ~ Group, data = cat_data)
  d_cat <- cohen.d(fixations_per_second ~ Group, data = cat_data)
  fixations_category_results[[cat]] <- format_results(t_test_cat, d_cat)  # Format results for each category
}

# Convert the list of results into a single data frame
fixations_category_df <- do.call(rbind, fixations_category_results)
fixations_category_df <- data.frame(Category = categories, fixations_category_df, row.names = NULL)  # Ensure Category column aligns

# Render tables in HTML format with kableExtra for APA style
knitr::kable(fixations_group_result, caption = "T-test and Effect Size for Fixations per Second (Experts vs. Novices)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

knitr::kable(fixations_category_df, caption = "T-test and Effect Size for Fixations per Second by Category") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

### Duration of disruption
```{r disrup_dur, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Define the list of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")

# Define categories for each type of disruption
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Filter data to include only the specified disruptions
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%
  dplyr::select(Group, Participant, TOI, Duration_of_interval) %>%
  mutate(Group = as_factor(Group))

# Add a column for Disruption_Category based on TOI values
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Disruption_Category)) # Keep only rows with valid category

# Convert duration from milliseconds to seconds and round to 2 decimal places
df_aoi_disrup <- df_aoi_disrup %>%
  mutate(Duration_of_interval_sec = round(Duration_of_interval / 1000, 2))

# First Table: Total duration of disruptions by group (Novice and Expert)
duration_summary_group <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarize(
    total_duration_sec = round(sum(Duration_of_interval_sec, na.rm = TRUE), 2),
    mean_duration_sec = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    sd_duration_sec = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    min_duration_sec = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    max_duration_sec = round(max(Duration_of_interval_sec, na.rm = TRUE), 2)
  )

# Second Table: Total duration of disruptions by group and category
duration_summary_category <- df_aoi_disrup %>%
  group_by(Group, Disruption_Category) %>%
  summarize(
    total_duration_sec = round(sum(Duration_of_interval_sec, na.rm = TRUE), 2),
    mean_duration_sec = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    sd_duration_sec = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    min_duration_sec = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    max_duration_sec = round(max(Duration_of_interval_sec, na.rm = TRUE), 2)
  )

# Display tables
print("Duration Summary by Group:")
print(duration_summary_group)

print("Duration Summary by Group and Category:")
print(duration_summary_category)

# First Plot: Boxplot of disruption duration by group
plot_duration_group <- df_aoi_disrup %>%
  ggplot(mapping = aes(x = Group, y = Duration_of_interval_sec)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  labs(x = "", y = "Disruption Duration (seconds)") +
  ggtitle("Disruption Duration by Group") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# Second Plot: Boxplot of disruption duration by group and category
plot_duration_category <- df_aoi_disrup %>%
  ggplot(mapping = aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) +
  labs(x = "", y = "Disruption Duration (seconds)") +
  ggtitle("Disruption Duration by Group and Category") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# Display the plots
plot_duration_group
plot_duration_category

# Define categories if not already defined
categories <- unique(df_aoi_disrup$Disruption_Category)

# Function to format t-test and Cohen's d results into a tidy data frame with additional debugging
format_results <- function(t_test, cohen_d) {
  if (is.list(cohen_d) && !is.null(cohen_d$estimate) && !is.null(cohen_d$conf.int)) {
    data.frame(
      "t-value" = round(t_test$statistic, 2),
      "df" = round(t_test$parameter, 2),
      "p-value" = format.pval(t_test$p.value, digits = 3, eps = .001),
      "Mean Difference" = round(t_test$estimate[1] - t_test$estimate[2], 2),
      "Cohen's d" = round(cohen_d$estimate, 2),
      "95% CI (d)" = paste0("[", round(cohen_d$conf.int[1], 2), ", ", round(cohen_d$conf.int[2], 2), "]")
    )
  } else {
    print("Cohen's d returned an unexpected structure or NA values. Setting values to NA.")
    print(cohen_d)  # Print the structure of cohen_d for debugging
    data.frame(
      "t-value" = round(t_test$statistic, 2),
      "df" = round(t_test$parameter, 2),
      "p-value" = format.pval(t_test$p.value, digits = 3, eps = .001),
      "Mean Difference" = round(t_test$estimate[1] - t_test$estimate[2], 2),
      "Cohen's d" = NA,
      "95% CI (d)" = NA
    )
  }
}

# 1. Overall t-test and Cohen's d for Duration by Group
t_duration_group <- t.test(Duration_of_interval_sec ~ Group, data = df_aoi_disrup)
d_duration_group <- tryCatch(cohen.d(Duration_of_interval_sec ~ Group, data = df_aoi_disrup), 
                             error = function(e) NA)
duration_group_result <- format_results(t_duration_group, d_duration_group)

# 2. t-test and Cohen's d for Duration by Category with debugging output
duration_category_results <- lapply(categories, function(cat) {
  cat_data <- df_aoi_disrup %>% filter(Disruption_Category == cat)

  # Perform t-test
  t_test_cat <- t.test(Duration_of_interval_sec ~ Group, data = cat_data)

  # Calculate Cohen's d with error handling
  d_cat <- tryCatch(cohen.d(Duration_of_interval_sec ~ Group, data = cat_data), 
                    error = function(e) NA)

  # Debugging: Check structure of d_cat for each category
  print(paste("Category:", cat))
  print("Structure of Cohen's d result:")
  str(d_cat)  # Display the structure of d_cat to debug

  # Format results with error handling
  format_results(t_test_cat, d_cat)
})

# Convert the list of results into a single data frame
duration_category_df <- do.call(rbind, duration_category_results)
duration_category_df <- data.frame(Category = categories, duration_category_df, row.names = NULL)

# Render tables in HTML format with kableExtra for APA style
knitr::kable(duration_group_result, caption = "T-test and Effect Size for Duration (Experts vs. Novices)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

knitr::kable(duration_category_df, caption = "T-test and Effect Size for Duration by Category") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Percentage of fixation durations on each AOI, relative to the total duration of fixations during the entire "Lesson" 
```{r dur_fix_percent, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Step 1: Define AOI categories with correct column names for fixation duration
aoi_columns <- list(
  "AOI_Students" = c(
    "Total_duration_of_fixations.Anna",
    "Total_duration_of_fixations.Bianca",
    "Total_duration_of_fixations.Carl(a)"
  ),
  "AOI_Disruptive_Person" = c("Total_duration_of_fixations.Disruptive_Person"),
  "AOI_Teacher_Material" = c(
    "Total_duration_of_fixations.Board_Screen",
    "Total_duration_of_fixations.Material_Teacher"
  ),
  "AOI_Student_Desk" = c(
    "Total_duration_of_fixations.Nametag_Anna",
    "Total_duration_of_fixations.Nametag_Bianca",
    "Total_duration_of_fixations.Nametag_Carl(a)",
    "Total_duration_of_fixations.Material_Students"
  ),
  "AOI_Classroom_Others" = c("Total_duration_of_fixations.Classroom_Others")
)

# Step 2: Filter data for "Lesson" and select relevant columns for fixation duration
df_durations <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(Participant, Group, starts_with("Total_duration_of_fixations"))

# Step 3: Calculate total fixation duration per participant, excluding "Disruptive Person"
relevant_columns <- unlist(aoi_columns[c("AOI_Students", "AOI_Teacher_Material", "AOI_Student_Desk", "AOI_Classroom_Others")])

df_durations <- df_durations %>%
  mutate(Total_duration_excluding_disruptive = rowSums(across(all_of(relevant_columns)), na.rm = TRUE))

# Step 4: Calculate fixation duration percentages for each AOI based on the adjusted total duration
aoi_durations_percentages <- df_durations %>%
  rowwise() %>%
  mutate(
    AOI_Students_pct = sum(c_across(any_of(aoi_columns$AOI_Students)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100,
    AOI_Disruptive_Person_pct = sum(c_across(any_of(aoi_columns$AOI_Disruptive_Person)), na.rm = TRUE) / (
      Total_duration_excluding_disruptive + sum(c_across(any_of(aoi_columns$AOI_Disruptive_Person)), na.rm = TRUE)
    ) * 100,
    AOI_Teacher_Material_pct = sum(c_across(any_of(aoi_columns$AOI_Teacher_Material)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100,
    AOI_Student_Desk_pct = sum(c_across(any_of(aoi_columns$AOI_Student_Desk)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100,
    AOI_Classroom_Others_pct = sum(c_across(any_of(aoi_columns$AOI_Classroom_Others)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100
  ) %>%
  ungroup()

# Step 5: Summarize fixation duration percentages by Group and arrange in descending order
grouped_aoi_durations_percentages <- aoi_durations_percentages %>%
  group_by(Group) %>%
  summarise(
    Mean_AOI_Students_pct = round(mean(AOI_Students_pct, na.rm = TRUE), 2),
    Mean_AOI_Disruptive_Person_pct = round(mean(AOI_Disruptive_Person_pct, na.rm = TRUE), 2),
    Mean_AOI_Teacher_Material_pct = round(mean(AOI_Teacher_Material_pct, na.rm = TRUE), 2),
    Mean_AOI_Student_Desk_pct = round(mean(AOI_Student_Desk_pct, na.rm = TRUE), 2),
    Mean_AOI_Classroom_Others_pct = round(mean(AOI_Classroom_Others_pct, na.rm = TRUE), 2)
  ) %>%
  pivot_longer(cols = starts_with("Mean_AOI"), names_to = "AOI", values_to = "Duration_Percentage") %>%
  mutate(AOI = recode(AOI,
                      "Mean_AOI_Students_pct" = "Students",
                      "Mean_AOI_Disruptive_Person_pct" = "Disruptive Person",
                      "Mean_AOI_Teacher_Material_pct" = "Teacher Material",
                      "Mean_AOI_Student_Desk_pct" = "Student Desk",
                      "Mean_AOI_Classroom_Others_pct" = "Classroom/Others")) %>%
  arrange(Group, desc(Duration_Percentage))

# Step 6: Plot the fixation duration percentages by AOI and Group
ggplot(grouped_aoi_durations_percentages, aes(x = reorder(AOI, -Duration_Percentage), y = Duration_Percentage, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = paste0(Duration_Percentage, "%")), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3, size = 3) +
  labs(
    title = "Average Fixation Duration Percentage by AOI and Group (Descending Order)",
    x = "AOI",
    y = "Fixation Duration Percentage (%)"
  ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(title = "Group")) +
  ylim(0, 100)  # Set y-axis limit to 100%

# Step 7: Perform a t-test between Experts and Novices on the duration percentages and calculate Cohen's d
aoi_list <- c("AOI_Students_pct", "AOI_Disruptive_Person_pct", "AOI_Teacher_Material_pct", "AOI_Student_Desk_pct", "AOI_Classroom_Others_pct")
results <- data.frame()

for (column_name in aoi_list) {
  if (column_name %in% names(aoi_durations_percentages)) {
    aoi_data <- aoi_durations_percentages %>%
      select(Group, !!sym(column_name)) %>%
      rename(Duration_Percentage = !!sym(column_name))

    expert_data <- aoi_data$Duration_Percentage[aoi_data$Group == "Expert"]
    novice_data <- aoi_data$Duration_Percentage[aoi_data$Group == "Novice"]

    if (length(expert_data) > 0 && length(novice_data) > 0) {
      t_test_result <- t.test(expert_data, novice_data, var.equal = TRUE)
      cohen_d_value <- cohen.d(expert_data, novice_data, pooled = TRUE)$estimate

      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = round(t_test_result$statistic, 2),
          "df" = t_test_result$parameter,
          "p-value" = format.pval(t_test_result$p.value, digits = 3, eps = .001),
          "Mean Difference" = round(t_test_result$estimate[1] - t_test_result$estimate[2], 2),
          "Cohen's d" = round(cohen_d_value, 2),
          "95% CI (d)" = paste0("[", round(t_test_result$conf.int[1], 2), ", ", round(t_test_result$conf.int[2], 2), "]")
        )
      )
    } else {
      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = NA,
          "df" = NA,
          "p-value" = NA,
          "Mean Difference" = NA,
          "Cohen's d" = NA,
          "95% CI (d)" = NA
        )
      )
    }
  }
}

# Map AOI names to clean versions for readability in the final results table
results$AOI <- recode(results$AOI,
                      "AOI_Students" = "Students",
                      "AOI_Disruptive_Person" = "Disruptive Person",
                      "AOI_Teacher_Material" = "Teacher Material",
                      "AOI_Student_Desk" = "Student Desk",
                      "AOI_Classroom_Others" = "Classroom/Others")

# Display the t-Test and Effect Size table in APA style
knitr::kable(results, 
             caption = "t-Test and Effect Size for Fixation Duration Percentage across AOIs (Experts vs. Novices)", 
             row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Average duration of fixations in milliseconds (micro-teaching unit)
```{r dur_all, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### AVERAGE FIXATION DURATION ANALYSIS ########################

# 1. Data Preparation: Filter and Calculate Average Fixation Duration in the Micro-Teaching Unit
df_aoi_sum <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Group,
    Participant,
    Duration_of_interval,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations"),
    starts_with("Average_duration"),
    !"Total_duration_of_fixations.Disruptive_Person",
    !"Number_of_fixations.Disruptive_Person",
    !"Average_duration_of_fixations.Disruptive_Person"
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Participant = Participant,
    Duration_of_interval = Duration_of_interval,
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Sum_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Sum_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Average_duration_fixation = round(Sum_duration_fixation / Sum_number_fixation, 2)
  ) %>%
  drop_na()

# 2. Descriptive Statistics Table
fixation_duration_table <- df_aoi_sum %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Average_duration_fixation, na.rm = TRUE), 2),
    SD = round(sd(Average_duration_fixation, na.rm = TRUE), 2),
    Min = round(min(Average_duration_fixation, na.rm = TRUE), 2),
    Max = round(max(Average_duration_fixation, na.rm = TRUE), 2)
  )

# Display the descriptive table in APA style
knitr::kable(fixation_duration_table, caption = "N, M, SD, min & max Average Fixation Duration (micro-teaching unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Calculate means and standard deviations for Novice and Expert groups
mean_novice_fixation <- round(mean(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Novice"], na.rm = TRUE), 2)
sd_novice_fixation <- round(sd(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Novice"], na.rm = TRUE), 2)
mean_expert_fixation <- round(mean(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Expert"], na.rm = TRUE), 2)
sd_expert_fixation <- round(sd(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Expert"], na.rm = TRUE), 2)

# 4. Perform t-test to get the p-value
t_test_fixation <- t.test(
  Average_duration_fixation ~ Group,
  data = df_aoi_sum,
  var.equal = TRUE
)
p_value_fixation <- round(t_test_fixation$p.value, 3)

# 5. Calculate Cohen's d for effect size using formula syntax
library(effsize)
cohen_d_fixation <- round(cohen.d(Average_duration_fixation ~ Group, data = df_aoi_sum)$estimate, 2)

# 6. Plotting the Average Fixation Duration
y_max_fixation <- max(df_aoi_sum$Average_duration_fixation) * 1.5

plot_avg_fixation_group <- df_aoi_sum %>%
  ggplot(mapping = aes(x = Group, y = Average_duration_fixation)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_fixation) +
  labs(x = "", y = "Average Fixation Duration (ms)") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Average Fixation Duration\n(micro-teaching unit)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  # Add annotations for mean (M) and SD
  annotate("text", x = 1, y = y_max_fixation * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice_fixation, ", italic(SD) == ", sd_novice_fixation, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max_fixation * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert_fixation, ", italic(SD) == ", sd_expert_fixation, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Add p-value and Cohen's d annotations
  annotate("text", x = 1.5, y = y_max_fixation * 0.95,
           label = paste0("p = ", p_value_fixation, "\nd = ", cohen_d_fixation),
           hjust = 0.5, size = 4, fontface = "italic")

plot_avg_fixation_group

# 7. Format the t-Test and Cohen's d results for APA Table without the first column
t_test_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_fixation$statistic, 2),
  "df" = t_test_fixation$parameter,
  "p-value" = round(t_test_fixation$p.value, 3),
  "Mean Difference" = round(t_test_fixation$estimate[1] - t_test_fixation$estimate[2], 2),
  "Cohen's d" = cohen_d_fixation,
  "95% CI (d)" = paste0("[", round(t_test_fixation$conf.int[1], 2), ", ", round(t_test_fixation$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size results in APA style without row names
knitr::kable(t_test_result, caption = "t-Test and Effect Size for Average Fixation Duration (Micro-Teaching Unit)", row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Average duration of fixations (AOI students)
```{r dur_students, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### AVERAGE DURATION ON AOI STUDENTS ########################

# 1. Data Preparation: Filter and Calculate Average Duration of Fixations on AOI Students
df_aoi_stud <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Group,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations")
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Stud_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Stud_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Average_duration_stud = round(Stud_duration_fixation / Stud_number_fixation, 2) # Retained in milliseconds
  ) %>%
  drop_na()

# 2. Descriptive Statistics Table
dur_stud_table <- df_aoi_stud %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M in ms" = round(mean(Average_duration_stud, na.rm = TRUE), 2),
    "SD in ms" = round(sd(Average_duration_stud, na.rm = TRUE), 2),
    "Min in ms" = round(min(Average_duration_stud, na.rm = TRUE), 2),
    "Max in ms" = round(max(Average_duration_stud, na.rm = TRUE), 2)
  )

# Display the descriptive table in APA style
knitr::kable(dur_stud_table,
             caption = "N, M, SD, min & max average duration of fixations in milliseconds (AOI students)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Plotting Average Duration on AOI Students
y_max_stud <- max(df_aoi_stud$Average_duration_stud) * 1.2  # Adjust y-axis limit based on data

plot_average_duration_stud <- df_aoi_stud %>%
  ggplot(mapping = aes(x = Group, y = Average_duration_stud)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_stud) + 
  labs(x = "", y = "Milliseconds") + 
  scale_fill_brewer(palette = "RdBu") + 
  ggtitle("Average duration of fixations\n(AOI students)") +
  theme_cowplot() +
  theme(legend.position = "none")

plot_average_duration_stud

# 4. t-Test and Effect Size Calculation for Average Duration of Fixations on AOI Students
# Conduct t-test
t_test_stud <- t.test(
  Average_duration_stud ~ Group,
  data = df_aoi_stud,
  var.equal = TRUE
)
t_test_stud_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_stud$statistic, 2),
  "df" = t_test_stud$parameter,
  "p-value" = round(t_test_stud$p.value, 3),
  "Mean Difference" = round(t_test_stud$estimate[1] - t_test_stud$estimate[2], 2)
)

# Calculate Cohen's d effect size
library(effsize)
d_aver_stud <- cohen.d(Average_duration_stud ~ Group, data = df_aoi_stud)
t_test_stud_result$`Cohen's d` <- round(d_aver_stud$estimate, 2)
t_test_stud_result$`95% CI (d)` <- paste0("[", round(d_aver_stud$conf.int[1], 2), ", ", round(d_aver_stud$conf.int[2], 2), "]")

# Display t-Test and Effect Size results in APA style
knitr::kable(t_test_stud_result,
             caption = "t-Test and Effect Size for Average Fixation Duration on AOI Students (milliseconds)",
             row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Average duration of fixations (AOI disruptive person)
```{r dur_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### AVERAGE DURATION ON AOI DISRUPTIVE PERSON ########################

# 1. Data Preparation: Filter and Calculate Average Duration of Fixations on AOI Disruptive Person
df_aoi_disrup <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Group,
    "Total_duration_of_fixations.Disruptive_Person",
    "Number_of_fixations.Disruptive_Person"
  ) %>%
  rowwise() %>%
  mutate(
    Group = as_factor(Group),
    Average_duration_disrup = round(Total_duration_of_fixations.Disruptive_Person / Number_of_fixations.Disruptive_Person, 2) # In milliseconds
  ) %>%
  drop_na()

# 2. Descriptive Statistics Table
dur_disrup_table <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M in ms" = round(mean(Average_duration_disrup, na.rm = TRUE), 2),
    "SD in ms" = round(sd(Average_duration_disrup, na.rm = TRUE), 2),
    "Min in ms" = round(min(Average_duration_disrup, na.rm = TRUE), 2),
    "Max in ms" = round(max(Average_duration_disrup, na.rm = TRUE), 2)
  )

# Display the descriptive table in APA style
knitr::kable(dur_disrup_table,
             caption = "N, M, SD, min & max average duration of fixations in milliseconds (AOI disruptive person)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Plotting Average Duration on AOI Disruptive Person
y_max_disrup <- max(df_aoi_disrup$Average_duration_disrup, na.rm = TRUE) * 1.2  # Adjust y-axis limit based on data

plot_average_duration_disrup <- df_aoi_disrup %>%
  ggplot(mapping = aes(x = Group, y = Average_duration_disrup)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_disrup) + 
  labs(x = "", y = "Milliseconds") + 
  scale_fill_brewer(palette = "RdBu") + 
  ggtitle("Average duration of fixations\n(AOI disruptive person)") +
  theme_cowplot() +
  theme(legend.position = "none")

plot_average_duration_disrup

# 4. t-Test and Effect Size Calculation for Average Duration of Fixations on AOI Disruptive Person
# Conduct t-test
t_test_disrup <- t.test(
  Average_duration_disrup ~ Group,
  data = df_aoi_disrup,
  var.equal = TRUE
)
t_test_disrup_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_disrup$statistic, 2),
  "df" = t_test_disrup$parameter,
  "p-value" = round(t_test_disrup$p.value, 3),
  "Mean Difference" = round(t_test_disrup$estimate[1] - t_test_disrup$estimate[2], 2)
)

# Calculate Cohen's d effect size
library(effsize)
d_dur_disrup <- cohen.d(Average_duration_disrup ~ Group, data = df_aoi_disrup)
t_test_disrup_result$`Cohen's d` <- round(d_dur_disrup$estimate, 2)
t_test_disrup_result$`95% CI (d)` <- paste0("[", round(d_dur_disrup$conf.int[1], 2), ", ", round(d_dur_disrup$conf.int[2], 2), "]")

# Display t-Test and Effect Size results in APA style
knitr::kable(t_test_disrup_result,
             caption = "t-Test and Effect Size for Average Fixation Duration on AOI Disruptive Person (milliseconds)",
             row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### t-test & effect size "Average duration of fixations" (AOI disruptive person)
```{r _disrup_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# AVERAGDE DURATION OF FIXATIONS on AOI disruptive person 
# t-test for expertise differences
t.test(x = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Expert"],
       y = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Novice"],
       var.equal = TRUE)

# NUMBER OF FIXATIONS 
# effect size for expertise differences
d_dur_disrup <- CohenD(x = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Expert"],
                        y = df_aoi_disrup$Average_duration_disrup[df_aoi_disrup$Group == "Novice"],
                        na.rm = TRUE)

round(d_dur_disrup, 2)

```

#### Gaze Relational Index (GRI; micro-teaching unit)
```{r gri_all, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### GRI IN MICRO-TEACHING UNIT ########################

# filter group, participants, total duration & number fixation in entire micro-teaching unit
df_aoi_sum %>%
  group_by(Group) %>% 
  mutate(
    GRI = round(Average_duration_mtu / Number_fixation_min_mtu, digits = 2),
    )-> df_gri

# N, M, SD, min & max number of fixation (micro-teaching unit)
gri.table <-
  df_gri %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M" = round(mean(GRI), digits = 2),
    "SD" = round(sd(GRI), digits = 2),
    "Min" = round(min(GRI), digits = 2),
    "Max" = round(max(GRI), digits = 2)
  )

knitr::kable(gri.table,
             caption = "N, M, SD, min & max GRI (micro-teaching unit)")

# plotting number of fixations on ALL AOIS
df_gri %>%
  ggplot(mapping = aes(x = Group, y = GRI)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  ylim(0, 12) +
  labs(x = "", y = "average duration of fixations/\nnumber of fixations per minute") +
  scale_fill_brewer(palette  = "RdBu") +
  ggtitle("GRI\n(micro-teaching unit)") +
  theme_cowplot() +
  theme(legend.position = "none") -> plot_gri_group

plot_gri_group

```

#### t-test & effect size "GRI" (micro-teaching unit)
```{r gri_all_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# GRI (mirco-teaching unit)
# t-test for expertise differences
t.test(x = df_gri$GRI[df_gri$Group == "Expert"],
       y = df_gri$GRI[df_gri$Group == "Novice"],
       var.equal = TRUE)

# GRI (mirco-teaching unit) 
# effect size for expertise differences
d_gri_all <- CohenD(x = df_gri$GRI[df_gri$Group == "Expert"],
                        y = df_gri$GRI[df_gri$Group == "Novice"],
                        na.rm = TRUE)

round(d_gri_all, 2)

```

```{r gri_stud, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# ##################### GRI on AOI students ########################
# 
# # filter group, participants, total duration & number fixation on AOI students
# df_aoi_stud %>%
#   group_by(Group) %>% 
#   mutate(
#     GRI_stud = round(Average_duration_stud / Stud_number_fixation, digits = 2),
#     )-> df_gri_stud
# 
# 
# # N, M, SD, min & max GRI (AOI students)
# gri.stud.table <-
#   df_gri_stud %>%
#   group_by(Group) %>%
#   summarise(
#     N = n(),
#     "M" = round(mean(GRI_stud), digits = 2),
#     "SD" = round(sd(GRI_stud), digits = 2),
#     "Min" = round(min(GRI_stud), digits = 2),
#     "Max" = round(max(GRI_stud), digits = 2)
#   )
# 
# knitr::kable(gri.stud.table,
#              caption = "N, M, SD, min & max GRI (AOI students)")
# 
# 
# # plotting GRI on AOI students
# df_gri_stud %>%
#   ggplot(mapping = aes(x = Group, y = GRI_stud)) +
#   geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
#   geom_point(
#     size = 2,
#     alpha = 0.7,
#     position = position_jitter(
#       seed = 1,
#       width = 0.1,
#       height = 0.1
#     )
#   ) +
#   ylim(0, 2.5) +
#   labs(x = "", y = "mean duration of fixations/\nmean number of fixations") +
#   scale_fill_brewer(palette  = "RdBu") +
#   ggtitle("GRI\n(AOI students)") +
#   theme_cowplot() +
#   theme(legend.position = "none") -> plot_gri_stud
# 
# plot_gri_stud

```

```{r gri_stud_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# # GRI (AOI students)
# # t-test for expertise differences
# t.test(x = df_gri_stud$GRI_stud[df_gri_stud$Group == "Expert"],
#        y = df_gri_stud$GRI_stud[df_gri_stud$Group == "Novice"],
#        var.equal = TRUE)
# 
# # GRI (AOI students) 
# # effect size for expertise differences
# d_gri_stud <- CohenD(x = df_gri_stud$GRI_stud[df_gri_stud$Group == "Expert"],
#                      y = df_gri_stud$GRI_stud[df_gri_stud$Group == "Novice"],
#                      na.rm = TRUE)
# 
# round(d_gri_stud, 2)

```

```{r gri_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# ##################### GRI on AOI students ########################
# 
# # filter
# df_aoi_disrup %>%
#   group_by(Group) %>% 
#   mutate(
#     GRI_disrup = round(Average_duration_disrup / Number_of_fixations.Disruptive_Person, digits = 2),
#     )-> df_gri_disrup
# 
# 
# # N, M, SD, min & max GRI (AOI disruptive person)
# gri.disrup.table <-
#   df_gri_disrup %>%
#   group_by(Group) %>%
#   summarise(
#     N = n(),
#     "M" = round(mean(GRI_disrup), digits = 2),
#     "SD" = round(sd(GRI_disrup), digits = 2),
#     "Min" = round(min(GRI_disrup), digits = 2),
#     "Max" = round(max(GRI_disrup), digits = 2)
#   )
# 
# knitr::kable(gri.disrup.table,
#              caption = "N, M, SD, min & max GRI (AOI disruptive person)")
# 
# 
# # plotting GRI on AOI disruptive person
# df_gri_disrup %>%
#   ggplot(mapping = aes(x = Group, y = GRI_disrup)) +
#   geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
#   geom_point(
#     size = 2,
#     alpha = 0.7,
#     position = position_jitter(
#       seed = 1,
#       width = 0.1,
#       height = 0.1
#     )
#   ) +
#   ylim(0, 25) +
#   labs(x = "", y = "mean duration of fixations/\nmean number of fixations") +
#   scale_fill_brewer(palette  = "RdBu") +
#   ggtitle("GRI\n(AOI disruptive person)") +
#   theme_cowplot() +
#   theme(legend.position = "none") -> plot_gri_disrup
# 
# plot_gri_disrup

```

```{r gri_disrup_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# # GRI (AOI disruptive person)
# # t-test for expertise differences
# t.test(x = df_gri_disrup$GRI_disrup[df_gri_disrup$Group == "Expert"],
#        y = df_gri_disrup$GRI_disrup[df_gri_disrup$Group == "Novice"],
#        var.equal = TRUE)
# 
# # GRI (AOI disruptive person)
# # effect size for expertise differences
# d_gri_disrup <- CohenD(x = df_gri_disrup$GRI_disrup[df_gri_disrup$Group == "Expert"],
#                        y = df_gri_disrup$GRI_disrup[df_gri_disrup$Group == "Novice"],
#                        na.rm = TRUE)
# 
# round(d_gri_disrup, 2)

```

#### Time to first fixation in seconds (AOI disruptive person)
```{r ttff_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

############### TIME TO FIRST FIXATION ####################

df_aoi %>% 
  filter(TOI %in% c("Chatting_with_neighbour",
                    "Clicking_pen",
                    "Drawing",
                    "Drumming_with_hands",
                    "Head_on_table",
                    "Heckling",
                    "Looking_at_phone",
                    "Snipping_with_fingers",
                    "Whispering"
                    )
         ) %>% 
  dplyr::select(Participant,
                TOI,
                Time_to_first_fixation.Disruptive_Person,
                ) %>% 
  rowwise() %>% 
  drop_na() %>% 
  group_by(Participant) %>% 
  filter(!Time_to_first_fixation.Disruptive_Person == 0, # exclude all participants with 0 msec
         !Time_to_first_fixation.Disruptive_Person > 30000) %>% # cut-off 30sec
  summarise(N = n(),
            Time_to_first_fixation.Disruptive_Person_Sum = sum(Time_to_first_fixation.Disruptive_Person/N)
            ) %>% 
  mutate(Group = ifelse(Participant < 200, "Novice", "Expert"),
         Sum_disruptions = sum(N),
         Disrup_time_fixation_sec = round(Time_to_first_fixation.Disruptive_Person_Sum / 1000, 
                                          digits = 2)) -> df_ttff_disrup # changing milliseconds into seconds

############### M, SD, min, max time to first fixation "Disruptive Person"
ttff.disrup.table <-
  df_ttff_disrup %>%
  filter(!Participant %in% c("216", "113", "123")) %>% 
  group_by(Group) %>%
  summarise(
    N = n(),
    "M in sec" = round(mean(Disrup_time_fixation_sec), digits = 2),
    "SD in sec" = round(sd(Disrup_time_fixation_sec), digits = 2),
    "Min in sec" = round(min(Disrup_time_fixation_sec), digits = 2),
    "Max in sec" = round(max(Disrup_time_fixation_sec), digits = 2)
  )

knitr::kable(ttff.disrup.table,
             caption = "N, M, SD, min & max time to first fixation in seconds (AOI disruptive person)")

############### M, SD, min, max of the perceived "Disruptive Person" ############### 
number.perceived.disrup.table <-
 df_ttff_disrup %>%
  filter(!Participant %in% c("216", "113", "123")) %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(N, na.rm = TRUE), digits = 2),
    SD = round(sd(N, na.rm = TRUE), digits = 2), 
    Min = min(N, na.rm = TRUE),
    Max = max(N, na.rm = TRUE)
  )

knitr::kable(number.perceived.disrup.table,
             caption = "N, M, SD, min & max of the perceived ´disruptive person´")

############### PLOTTING ###############

##################### plotting "perceived disruptive person" for groups ##################### 
plot_n_disrup <- 
  df_ttff_disrup %>%
  filter(!Participant %in% c("216", "113", "123")) %>%
  ggplot(mapping = aes(x = Group,
                       y = N)) +
 geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  ylim(0,10) +
  labs(x ="",
       y = "Counts") + 
  scale_fill_brewer(palette  = "RdBu") +
  ggtitle("Counts of the perceived ´disruptive person´\n") +
  theme_cowplot() +
  theme(legend.position="none")

plot_n_disrup

##################### plotting time to first fixation for groups ##################### 
plot_time_disrup <- 
  df_ttff_disrup %>%
  filter(!Participant %in% c("216", "113", "123")) %>%
  ggplot(mapping = aes(x = Group,
                       y = Disrup_time_fixation_sec)) +
 geom_boxplot(mapping = aes(fill = Group),
               outlier.shape = NA) +
  geom_point(size = 2,
             alpha = 0.7,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  ylim(0,10) +
  labs(x ="",
       y = "Seconds") + 
  scale_fill_brewer(palette  = "RdBu") +
  ggtitle("Time to first fixation\n(AOI disruptive person)") +
  theme_cowplot() +
  theme(legend.position="none")

plot_time_disrup

```

#### t-test & effect size "Time to first fixation" (AOI disruptive person)
```{r ttff_disrup_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# TIME TO FIRST FIXATIONS on disruptive person 
# t-test for expertise differences
t.test(x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"],
       y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"],
       var.equal = TRUE)

# TIME TO FIRST FIXATIONS on disruptive person 
# effect size for expertise differences
d_time_disrup <- CohenD(x = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Expert"],
                    y = df_ttff_disrup$Disrup_time_fixation_sec[df_ttff_disrup$Group == "Novice"],
                    na.rm = TRUE)

round(d_time_disrup, 2)

```

### Rating Scales (Disruption Appraisal, Confidence Appraisal, Prevalence Rating)
```{r rating_scales, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# read in data and combine multiple excel sheets into a single table 
sri <-
  excel_sheets("data/Coding_SRI.xlsx") %>%
  map_df(~ read_xlsx("data/Coding_SRI.xlsx", .)) %>%
  filter(
    !ID %in% c(201),
    !disruption_appraisal == -100,
    !confidence_appraisal == -100,
    !disruption_appraisal == -99,
    !confidence_appraisal == -99,
    !disruption_appraisal == -88,
    !confidence_appraisal == -88,
    !prevalence_rating == -100,
    !prevalence_rating == -99,
    !prevalence_rating == -88
    ) %>%
  mutate(ID = ifelse(ID < 200, "Novice", "Expert"))

############## Disruption appraisal #########################

sri.disrup.table <-
  sri %>%
  group_by(ID) %>%
  summarise(
    N = n(),
    "M" = round(mean(disruption_appraisal), digits = 2),
    "SD" = round(sd(disruption_appraisal), digits = 2),
    "Min" = min(disruption_appraisal),
    "Max" = max(disruption_appraisal)
    )

knitr::kable(sri.disrup.table,
             caption = "Disruption Appraisal")

### with event
sri.disrup.event.table <-
  sri %>%
  group_by(ID, event) %>%
  summarise(
    N = n(),
    "M" = round(mean(disruption_appraisal), digits = 2),
    "SD" = round(sd(disruption_appraisal), digits = 2),
    "Min" = min(disruption_appraisal),
    "Max" = max(disruption_appraisal)
  )

# insert a table into HTML
knitr::kable(sri.disrup.event.table,
             caption = "Disruption appraisal with event")

# create a new data frame with appraisal
sri_disrup <- subset.data.frame(sri, select = c(ID, event, disruption_appraisal))

# plotting disruption appraisal for groups
dist_group_plot <- 
  sri_disrup %>% 
  mutate(ID = factor(ID,
                     levels = c("Novice",
                                "Expert"
                                )
                     )
         ) %>% 
  ggplot(mapping = aes(x = ID,
                       y = disruption_appraisal)) +
  geom_boxplot(mapping = aes(fill = ID),
               outlier.shape = NA) +
  geom_point(size = 1,
             alpha = 0.4,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  scale_x_discrete(limits = c("Novice", "Expert")) +
  scale_y_continuous(breaks = seq(from = 0, to = 10, by = 1)) + 
  labs(x = "",
       y = "Disruption Appraisal") + 
  scale_fill_brewer(palette  = "RdBu") +  
  ggtitle("How disruptive was the event for you?") +
  theme_cowplot()

dist_group_plot
```

#### t-Test & effect size "Disruption appraisal"
```{r sri_da_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#################### T-TEST & EFFECT SIZE ############

# Disruption appraisal 
# t-test for expertise differences
t.test(x = sri$disruption_appraisal[sri$ID == "Expert"],
       y = sri$disruption_appraisal[sri$ID == "Novice"],
       var.equal = TRUE)

# Disruption appraisal 
# effect size for expertise differences
d_sri_disrup <- CohenD(x = sri$disruption_appraisal[sri$ID == "Expert"],
                       y = sri$disruption_appraisal[sri$ID == "Novice"],
                       na.rm = TRUE)

round(d_sri_disrup, 2)
```

```{r sri_ca, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

############## Confidence appraisal #########################

sri.confi.table <-
  sri %>%
  group_by(ID) %>%
  summarise(
    "M" = round(mean(confidence_appraisal), digits = 2),
    "SD" = round(sd(confidence_appraisal), digits = 2),
    "Min" = min(confidence_appraisal),
    "Max" = max(confidence_appraisal)
  )

# insert a table into HTML
knitr::kable(sri.confi.table,
             caption = "Confidence appraisal")

### with event
sri.confi.event.table <- 
  sri %>%
  group_by(ID, event) %>%
  summarise(
    N = n(),
    "M" = round(mean(confidence_appraisal), digits = 2),
    "SD" = round(sd(confidence_appraisal), digits = 2),
    "Min" = min(confidence_appraisal),
    "Max" = max(confidence_appraisal)
  )

knitr::kable(sri.confi.event.table,
             caption = "Confidence appraisal with event")

# create a new data frame with appraisal
sri_confi <- subset.data.frame(sri, 
                               select = c(ID, 
                                          event, 
                                          confidence_appraisal)
                               )

# plotting confidence appraisal for groups
confi_group_plot <- 
  sri_confi %>% 
  mutate(ID = factor(ID,
                     levels = c("Novice",
                                "Expert"
                                )
                     )
         ) %>% 
  ggplot(mapping = aes(x = ID,
                       y = confidence_appraisal)) +
  geom_boxplot(mapping = aes(fill = ID),
               outlier.shape = NA) +
  geom_point(size = 1,
             alpha = 0.4,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  scale_x_discrete(limits = c("Novice", "Expert")) +
  scale_y_continuous(breaks = seq(from = 0, to = 10, by = 1)) + 
  labs(x = "",
       y = "Confidence Appraisal") + 
  scale_fill_brewer(palette  = "RdBu") +  
  ggtitle("How confident did you feel dealing with this event?") +
  theme_cowplot()

confi_group_plot
```

#### t-Test & effect size "Confidence appraisal"
```{r sri_ca_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#################### T-TEST & EFFECT SIZE ############

# Confidence Appraisal 
# t-test for expertise differences
t.test(x = sri$confidence_appraisal[sri$ID == "Expert"],
       y = sri$confidence_appraisal[sri$ID == "Novice"],
       var.equal = TRUE)

# Disruption Factor 
# effect size for expertise differences
d_sri_confi <- CohenD(x = sri$confidence_appraisal[sri$ID == "Expert"],
                      y = sri$confidence_appraisal[sri$ID == "Novice"],
                      na.rm = TRUE)

round(d_sri_confi, 2)
```

### Prevalence Rating as manipulation check
```{r sri_pr, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

############## Prevalence Rating #########################

sri.preva.table <-
  sri %>%
  group_by(ID) %>%
  summarise(
    N = n(),
    "M" = round(mean(prevalence_rating), digits = 2),
    "SD" = round(sd(prevalence_rating), digits = 2),
    "Min" = min(prevalence_rating),
    "Max" = max(prevalence_rating)
  )

knitr::kable(sri.preva.table,
             caption = "Prevalence rating")

sri.preva.event.table <-
  sri %>%
  group_by(ID, event) %>%
  summarise(
    N = n(),
    "M" = round(mean(prevalence_rating), digits = 2),
    "SD" = round(sd(prevalence_rating), digits = 2),
    "Min" = min(prevalence_rating),
    "Max" = max(prevalence_rating)
  )

knitr::kable(sri.preva.event.table,
             caption = "Prevalence rating with events")

sri_preva <- subset.data.frame(sri, select = c(ID, event, prevalence_rating))

preva_group_plot <- 
  sri_preva %>% 
  mutate(ID = factor(ID,
                     levels = c("Novice",
                                "Expert"
                                )
                     )
         ) %>% 
  ggplot(mapping = aes(x = ID,
                       y = prevalence_rating)) +
  geom_boxplot(mapping = aes(fill = ID),
               outlier.shape = NA) +
  geom_point(size = 1,
             alpha = 0.4,
             position = position_jitter(seed = 1,
                                        width = 0.1,
                                        height = 0.1)) +
  scale_x_discrete(limits = c("Novice", "Expert")) +
  scale_y_continuous(breaks = seq(from = 0, to = 10, by = 1)) + 
  labs(x = "",
       y = "Prevalence Rating") + 
  scale_fill_brewer(palette  = "RdBu") +  
  ggtitle("How widespread is this event in the classroom?") +
  theme_cowplot()

preva_group_plot

#################### T-TEST & EFFECT SIZE ############

t.test(x = sri$prevalence_rating[sri$ID == "Expert"],
       y = sri$prevalence_rating[sri$ID == "Novice"],
       var.equal = TRUE)

d_sri_preva <- CohenD(x = sri$prevalence_rating[sri$ID == "Expert"],
                      y = sri$prevalence_rating[sri$ID == "Novice"],
                      na.rm = TRUE)

round(d_sri_preva, 2)
```

### Internal consistency (Omega) for disruption and confidence appraisal
```{r sri_omega, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Select the relevant columns (e.g., Disruption Appraisal and Confidence Appraisal)
sri <-
  excel_sheets("data/Coding_SRI.xlsx") %>%
  map_df(~ read_xlsx("data/Coding_SRI.xlsx", .)) %>%
  filter(
    !ID %in% c(201),
    !disruption_appraisal == -100,
    !confidence_appraisal == -100,
    !disruption_appraisal == -99,
    !confidence_appraisal == -99,
    !disruption_appraisal == -88,
    !confidence_appraisal == -88,
    !prevalence_rating == -100,
    !prevalence_rating == -99,
    !prevalence_rating == -88
    )

# change long to wide format
sri_wide <- sri %>%
  pivot_wider(names_from = event, 
              values_from = c(disruption_appraisal, confidence_appraisal, prevalence_rating))

### DISRUPTIONS APPRAISAL

# create new data frame with only disruption items
sri_wide_disrup <- sri_wide %>%
  dplyr::select(starts_with("disruption_appraisal")) %>% 
  drop_na()

# calculate omega with omega() function  
omega_disrup <- omega(sri_wide_disrup)

# display the results 
print(omega_disrup)

### CONFIDENCE APPRAISAL

# create new data frame with only confidence items
sri_wide_confi <- sri_wide %>%
  dplyr::select(starts_with("confidence_appraisal")) %>% 
  drop_na()

# calculate omega with omega() function  
omega_confi <- omega(sri_wide_confi)

# display the results 
print(omega_confi)

```

### Situational Jugdement Test
```{r sjt, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# read in data
df_sjt <- 
  read_excel("./data/SJT.xlsx")

# select relevant columns --> only subject ID and mean 
df_sjt <- df_sjt %>%
  dplyr::select(UI06_05, # ID
         SJT_AL_gek, # Monitoring
         SJT_ST_gek, # Managing momentum
         SJT_R_gek, # Rules and routines
         SJT_KF_gek) %>% # All
  filter(!UI06_05 == 201) # exclude ID with no eye tracking data

# remove NAs
df_sjt <- na.omit(df_sjt)

# rename columns
df_sjt <- rename(df_sjt, c("Group" = "UI06_05",
                           "Monitoring" = "SJT_AL_gek",
                           "Managing momentum" = "SJT_ST_gek",
                           "Rules and routines" = "SJT_R_gek",
                           "All" = "SJT_KF_gek"))

# changing from wide to long format
df_sjt_long <- df_sjt %>% 
  pivot_longer(!Group, names_to = "Facets Classroom Management", values_to = "Mean")

# define expert and novice with ifelse function
df_sjt_long$Group = ifelse(df_sjt_long$Group < 200, "Novice","Expert")

#################### N, MEAN & SD ############

# mean & SD SJT ALL
sjt_mean <- df_sjt_long %>%
  filter(`Facets Classroom Management` == "All") %>%
  group_by(Group) %>%
  summarise(N = n(),
            "M" = round(mean(Mean), 2),
            "SD" = round(sd(Mean), 2), 
            "Min" = round(min(Mean), 2),
            "Max" = round(max(Mean), 2), 
            )

# insert a table into HTML
knitr::kable(sjt_mean,
             caption = "N, M and SD for overall value")

# mean SJT Managing Momentum
sjt_mm <- df_sjt_long %>%
  filter(`Facets Classroom Management` == "Managing momentum") %>%
  group_by(Group) %>%
  summarise(N = n(),
            "M" = round(mean(Mean), 2),
            "SD" = round(sd(Mean), 2), 
            "Min" = round(min(Mean), 2),
            "Max" = round(max(Mean), 2), 
            )

# insert a table into HTML
knitr::kable(sjt_mm,
             caption = "N, M and SD for managing momentum")

# mean SJT Monitoring
sjt_m <- df_sjt_long %>%
  filter(`Facets Classroom Management` == "Monitoring") %>%
  group_by(Group) %>%
  summarise(N = n(),
            "M" = round(mean(Mean), 2),
            "SD" = round(sd(Mean), 2), 
            "Min" = round(min(Mean), 2),
            "Max" = round(max(Mean), 2), 
            )

# insert a table into HTML
knitr::kable(sjt_m,
             caption = "N, M and SD for monitoring")

# mean SJT Rules and routins
sjt_r <- df_sjt_long %>%
  filter(`Facets Classroom Management` == "Rules and routines") %>%
  group_by(Group) %>%
  summarise(N = n(),
            "M" = round(mean(Mean), 2),
            "SD" = round(sd(Mean), 2), 
            "Min" = round(min(Mean), 2),
            "Max" = round(max(Mean), 2), 
            )

# insert a table into HTML
knitr::kable(sjt_r,
             caption = "N, M and SD for rules and routines")

#################### Plots ############

# plotting mean of all aspects 
mean_plot <- 
df_sjt_long %>% 
  mutate(Group = factor(Group,
                        levels = c("Novice", 
                                   "Expert")
                        )
         ) %>% 
  ggplot(mapping = aes(x = Group,
                       y = Mean)) +
  geom_boxplot(mapping = aes(fill = Group), 
               outlier.shape = NA) +
  geom_point(size = 1, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1,
                                        height = 0.1)) +
  labs(x = "") +
  ylim(0,1) + 
  scale_fill_brewer(palette  = "RdBu") +  
  facet_wrap(vars(`Facets Classroom Management`), 
             nrow = 1, strip.position = "bottom") +
  theme_cowplot() +
  ggtitle("SJT") +
    theme(
    plot.title = element_text(size = 25, 
                              face = "bold"),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 15),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text.x = element_text(size = 8),
    axis.title.y = element_text(size = 20),
    axis.title.x = element_text(size = 22)
        )

mean_plot
```

#### t-test & effect size "STJ - All"
```{r sjt_all_ttest, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#################### T-TEST & EFFECT SIZE ##################

# define expert and novice with ifelse function in wide format
df_sjt$Group = ifelse(df_sjt$Group < 200, "Novice","Expert")

#### ALL #### 
# t-test for expertise differences 
t.test(x = df_sjt$All[df_sjt$Group == "Expert"],
       y = df_sjt$All[df_sjt$Group == "Novice"],
       var.equal = TRUE)

# effect size for expertise differences
d_sjt_all <- CohenD(x = df_sjt$All[df_sjt$Group == "Expert"],
                    y = df_sjt$All[df_sjt$Group == "Novice"],
                    na.rm = TRUE
                    )

round(d_sjt_all, 2)

```

#### t-test & effect size "SJT - Managing momentum"
```{r sjt_mm, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#### MANAGING MOMENTUM #### 
# t-test for expertise differences 
t.test(x = df_sjt$`Managing momentum`[df_sjt$Group == "Expert"],
       y = df_sjt$`Managing momentum`[df_sjt$Group == "Novice"],
       var.equal = TRUE
       )

# effect size for expertise differences
d_sjt_mm <- CohenD(x = df_sjt$`Managing momentum`[df_sjt$Group == "Expert"],
                   y = df_sjt$`Managing momentum`[df_sjt$Group == "Novice"],
                   na.rm = TRUE
                  )

round(d_sjt_mm, 2)
```

#### t-test & effect size "SJT - Monitoring"
```{r sjt_m, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#### MONITORING ####
# t-test for expertise differences 
t.test(x = df_sjt$Monitoring[df_sjt$Group == "Expert"],
       y = df_sjt$Monitoring[df_sjt$Group == "Novice"],
       var.equal = TRUE
       )

# effect size for expertise differences
d_sjt_moni <- CohenD(x = df_sjt$Monitoring[df_sjt$Group == "Expert"],
                     y = df_sjt$Monitoring[df_sjt$Group == "Novice"],
                     na.rm = TRUE
                    )

round(d_sjt_moni, 2)
```

#### t-test & effect size "SJT - Rules & routines"
```{r sjt_rr, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#### RULES & ROUTINES #### 
# t-test for expertise differences 
t.test(x = df_sjt$`Rules and routines`[df_sjt$Group == "Expert"],
       y = df_sjt$`Rules and routines`[df_sjt$Group == "Novice"],
       var.equal = TRUE
      )

# effect size for expertise differences
d_sjt_rr <- CohenD(x = df_sjt$`Rules and routines`[df_sjt$Group == "Expert"],
                   y = df_sjt$`Rules and routines`[df_sjt$Group == "Novice"],
                   na.rm = TRUE
                  )

round(d_sjt_rr, 2)

```

### Internal consistency (Omega) for SJT
```{r sjt_omega, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# read in data
df_sjt <- 
  read_excel("./data/SJT.xlsx")

### MANAGING MOMENTUM

# select relevant columns
df_sjt_mm <- df_sjt %>%
    dplyr::select(WT_S1_F1F2_AL:WT_S4_F4F6_AL)

# remove NAs
df_sjt_mm <- na.omit(df_sjt_mm)

# calculate omega with omega() function  
omega_sjt_mm <- omega(df_sjt_mm)

# display the results 
print(omega_sjt_mm)

### MONITORING

# select relevant columns
df_sjt_m <- df_sjt %>%
    dplyr::select(WT_S5_F1F2_ST:WT_S9_F4F5_ST)

# remove NAs
df_sjt_m <- na.omit(df_sjt_m)

# calculate omega with omega() function  
omega_sjt_m <- omega(df_sjt_m)

# display the results 
print(omega_sjt_m)

### RULES & ROUTINES

# select relevant columns
df_sjt_r <- df_sjt %>%
  dplyr::select(WT_S11_F1F5_R:WT_S14_F4F6_R)

# remove NAs
df_sjt_r <- na.omit(df_sjt_r)

# calculate omega with omega() function  
omega_sjt_r <- omega(df_sjt_r)

# display the results 
print(omega_sjt_r)

```

### Classroom Questionnaire
```{r classroom questionnaire, echo=FALSE, results='asis'}

# load data files with modified scales
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LI06_05,
                LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) %>% # non- & paraverbal communication 
  filter(!LI06_05 == 201) %>% # exclude ID with no eye tracking data
  mutate(LI06_05 = ifelse(LI06_05 < 200, "Novice", "Expert")) %>% 
  rename(Group = LI06_05)

# remove NAs
df_quest <- na.omit(df_quest)

# create table with means
df_quest_table <-
  df_quest %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M cm" = round(mean(c_across(starts_with(
      "LM"
    )), na.rm = TRUE), 2),
    "SD cm" = round(sd(c_across(starts_with(
      "LM"
    )), na.rm = TRUE), 2),
    "Min cm" = round(min(c_across(starts_with(
      "LM"
    )), na.rm = TRUE), 2),
    "Max cm" = round(max(c_across(starts_with(
      "LM"
    )), na.rm = TRUE), 2),
    # "N cm" = length(c_across(starts_with("LP")), na.rm = TRUE),
    "M n&pv com" = round(mean(c_across(starts_with(
      "LP"
    )), na.rm = TRUE), 2),
    "SD n&pv com" = round(sd(c_across(starts_with(
      "LP"
    )), na.rm = TRUE), 2),
    "Min n&pv com" = round(min(c_across(starts_with(
      "LP"
    )), na.rm = TRUE), 2),
    "Max n&pv com" = round(max(c_across(starts_with(
      "LP"
    )), na.rm = TRUE), 2)
  )

knitr::kable(df_quest_table,
             caption = "Mean, SD, min, max for classroom managament (cm) and non-/paraverbal communication (n&pv com)")

#################### Plots ############

# create table with means
df_quest_plot <-
  df_quest %>%
  rowwise() %>% 
  transmute(Group = Group,
            "All" = round(mean(c_across(starts_with("L")), na.rm = TRUE), 2),
            "Classroom management" = round(mean(c_across(starts_with("LM")), na.rm = TRUE), 2),
            "Non-/paraverbal communication" = round(mean(c_across(starts_with("LP")), na.rm = TRUE), 2)
            )

# changing from wide to long format
df_quest_long <- df_quest_plot %>% 
  pivot_longer(!Group, names_to = "Scales Classroom Questionnaire", values_to = "Mean")

# plotting mean of all aspects 
quest_plot <- 
df_quest_long %>% 
  mutate(Group = factor(Group,
                        levels = c("Novice", 
                                   "Expert")
                        )
         ) %>% 
  ggplot(mapping = aes(x = Group,
                       y = Mean)) +
  geom_boxplot(mapping = aes(fill = Group), 
               outlier.shape = NA) +
  geom_point(size = 1, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1,
                                        height = 0.1)) +
  labs(x = "") +
  ylim(0,4) + 
  scale_fill_brewer(palette  = "RdBu") +  
  facet_wrap(vars(`Scales Classroom Questionnaire`), 
             nrow = 1, strip.position = "bottom") +
  theme_cowplot() +
  ggtitle("Classroom Questionnaire") +
    theme(
    plot.title = element_text(size = 25, 
                              face = "bold"),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 15),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text.x = element_text(size = 8),
    axis.title.y = element_text(size = 20),
    axis.title.x = element_text(size = 22)
        )

quest_plot
```

#### t-test & effect size "Classroom Questionnaire - All"
```{r classroom questionnaire_all, echo=FALSE, results='asis'}

#################### T-TEST & EFFECT SIZE ##################

#### All #### 
# t-test for expertise differences 
t.test(x = df_quest_plot$All[df_quest_plot$Group == "Expert"],
       y = df_quest_plot$All[df_quest_plot$Group == "Novice"],
       var.equal = TRUE
       )

# effect size for expertise differences
d_quest_all <- CohenD(x = df_quest_plot$All[df_quest_plot$Group == "Expert"],
                      y = df_quest_plot$All[df_quest_plot$Group == "Novice"],
                      na.rm = TRUE
                      )

round(d_quest_all, 2)

```

#### t-test & effect size "Classroom Questionnaire - Classroom Management"
```{r classroom questionnaire_cm, echo=FALSE, results='asis'}

#### Classroom Management #### 
# t-test for expertise differences 
t.test(x = df_quest_plot$`Classroom management`[df_quest_plot$Group == "Expert"],
       y = df_quest_plot$`Classroom management`[df_quest_plot$Group == "Novice"],
       var.equal = TRUE)

# effect size for expertise differences
d_quest_cm <- CohenD(x = df_quest_plot$`Classroom management`[df_quest_plot$Group == "Expert"],
                      y = df_quest_plot$`Classroom management`[df_quest_plot$Group == "Novice"],
                      na.rm = TRUE
                      )

round(d_quest_cm, 2)
```

#### t-test & effect size "Classroom Questionnaire - Non-/paraverbal communication"
```{r classroom questionnaire_npvc, echo=FALSE, results='asis'}

#### Non-/paraverbal communication #### 
# t-test for expertise differences 
t.test(x = df_quest_plot$`Non-/paraverbal communication`[df_quest_plot$Group == "Expert"],
       y = df_quest_plot$`Non-/paraverbal communication`[df_quest_plot$Group == "Novice"],
       var.equal = TRUE)

# effect size for expertise differences
d_quest_npvc <- CohenD(x = df_quest_plot$`Non-/paraverbal communication`[df_quest_plot$Group == "Expert"],
                      y = df_quest_plot$`Non-/paraverbal communication`[df_quest_plot$Group == "Novice"],
                      na.rm = TRUE
                      )

round(d_quest_npvc, 2)

```

### Internal consistency (Omega) for Classroom Questionnaire
```{r cm_omega, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# read in data
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) # non- & paraverbal communication

# calculate omega with omega() function  
df_quest <- omega(df_quest)

# display the results 
print(df_quest)

### CLASSROOM MANAGEMENT
# read in data
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) # non- & paraverbal communication

# select relevant columns
df_quest_cm <- df_quest %>%
    dplyr::select(LM01_01:LM01_08)

# remove NAs
df_quest_cm <- na.omit(df_quest_cm)

# calculate omega with omega() function  
omega_quest_cm <- omega(df_quest_cm)

# display the results 
print(omega_quest_cm)

### NON- & PARAVERBAL COMMUNICATION
# read in data
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LM01_01:LM01_08, # classroom management 
                LP01_01:LP01_08) # non- & paraverbal communication

# select relevant columns
df_quest_npvc <- df_quest %>%
    dplyr::select(LP01_01:LP01_08)

# remove NAs
df_quest_npvc <- na.omit(df_quest_npvc)

# calculate omega with omega() function  
omega_quest_npvc <- omega(df_quest_npvc)

# display the results 
print(omega_quest_npvc)

```

## Correlations
```{r correlations, echo=FALSE, message=FALSE, warning=FALSE}

# demo data 
df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% 
  filter(!LI06_05 == 201,
         !LI06_05 == 223) %>% # exclude participant due to missing eye tracking data
  transmute(Participant = LI06_05,
            `Teaching Experience` = LI04_01,
            )

# average duration, number fixation, GRI (micro-teaching unit)
df_aoi %>%
  filter(TOI == "Lesson") %>% 
  dplyr::select(
    Duration_of_interval,
    Participant,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations"),
    starts_with("Average_duration"),
    !"Total_duration_of_fixations.Disruptive_Person",
    !"Number_of_fixations.Disruptive_Person",
    !"Average_duration_of_fixations.Disruptive_Person"
  ) %>%
  rowwise() %>%
  transmute(
    Participant = Participant,
    Duration_of_interval = Duration_of_interval, 
    Duration_of_interval_min = round(Duration_of_interval/60000, digits = 2),
    Sum_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Sum_average_duration_fixation = sum(c_across(starts_with("Average_duration")), na.rm = TRUE),
    Sum_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Average_duration_mtu = round(Sum_duration_fixation / Sum_number_fixation, digits = 0),
    Number_fixation_min_mtu = round(Sum_number_fixation / Duration_of_interval_min, digits = 0),
    GRI_mtu = round(Average_duration_mtu / Number_fixation_min_mtu, digits = 2)
    ) %>%
  drop_na() %>% 
  dplyr::select(Participant,
         Number_fixation_min_mtu, 
         Average_duration_mtu,
         GRI_mtu) -> df_aoi_numb_dur

# average duration, number fixation, GRI (AOI student)
df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select (
    Participant,
    Duration_of_interval,
    "Total_duration_of_fixations.Anna",
    "Total_duration_of_fixations.Bianca",
    "Total_duration_of_fixations.Carl(a)",
    "Number_of_fixations.Anna",
    "Number_of_fixations.Bianca",
    "Number_of_fixations.Carl(a)"
  ) %>%
  rowwise() %>%
  transmute(
    Participant = Participant,
    Duration_of_interval = Duration_of_interval,
    Duration_of_interval_min = round(Duration_of_interval/60000, digits = 2),
    Stud_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Stud_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Stud_number_fixation_min = round(Duration_of_interval_min / Stud_number_fixation, digits = 2),
    Average_duration_stud = round(Stud_duration_fixation / Stud_number_fixation, digits = 0),
    GRI_stud = round(Average_duration_stud / Stud_number_fixation, digits = 2)
    ) %>%
  dplyr::select(Participant,
                Stud_number_fixation_min,
                Average_duration_stud,
                GRI_stud) -> df_aoi_stud

# average duration, number fixation, GRI (AOI disruptive person)
df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(Participant,
                "Total_duration_of_fixations.Disruptive_Person",
                "Number_of_fixations.Disruptive_Person"
                ) %>%
  rowwise() %>%
  mutate(
    Average_duration_disrup = round(
      Total_duration_of_fixations.Disruptive_Person / Number_of_fixations.Disruptive_Person,
      digits = 0),
    GRI_disrup = round(
      Average_duration_disrup / Number_of_fixations.Disruptive_Person, digits = 2)) %>%
  dplyr::select(Participant,
                Number_of_fixations.Disruptive_Person,
                Average_duration_disrup,
                GRI_disrup
                ) -> df_aoi_disrup

df_aoi %>% 
  filter(TOI %in% c("Chatting_with_neighbour",
                    "Clicking_pen",
                    "Drawing",
                    "Drumming_with_hands",
                    "Head_on_table",
                    "Heckling",
                    "Looking_at_phone",
                    "Snipping_with_fingers",
                    "Whispering"
                    )
         ) %>% 
  dplyr::select(Participant,
                TOI,
                Time_to_first_fixation.Disruptive_Person,
                ) %>% 
  rowwise() %>% 
  drop_na() %>% 
  group_by(Participant) %>% 
  filter(!Time_to_first_fixation.Disruptive_Person == 0, # exclude all participants with 0 msec
         !Time_to_first_fixation.Disruptive_Person > 30000) %>% # cut-off 30sec
  summarise(N = n(),
            Time_to_first_fixation.Disruptive_Person_Sum = sum(Time_to_first_fixation.Disruptive_Person/N)
            ) %>% 
  mutate(Group = ifelse(Participant < 200, "Novice", "Expert"),
         Sum_disruptions = sum(N),
         Disrup_time_fixation_sec = round(Time_to_first_fixation.Disruptive_Person_Sum / 1000, 
                                          digits = 2)) -> df_ttff_disrup # changing milliseconds into seconds

# rating scales
df_sri <-
  excel_sheets("data/Coding_SRI.xlsx") %>%
  map_df(~ read_xlsx("data/Coding_SRI.xlsx", .)) %>%
  filter(
    !ID %in% c(201,
               223),
    !disruption_appraisal == -100,
    !confidence_appraisal == -100,
    !disruption_appraisal == -99,
    !confidence_appraisal == -99,
    !disruption_appraisal == -88,
    !confidence_appraisal == -88,
    !prevalence_rating == -100,
    !prevalence_rating == -99,
    !prevalence_rating == -88
    ) %>% 
  rename(Participant = ID) %>%
  group_by(Participant) %>% 
  summarise(
    Mean_disruption_appraisal = round(mean(disruption_appraisal), 2),
    Mean_confidence_appraisal = round(mean(confidence_appraisal), 2),
    Mean_prevalence_rating = round(mean(prevalence_rating), 2)
    )%>% 
  dplyr::select(Participant,
                Mean_disruption_appraisal,
                Mean_confidence_appraisal)

# sjt
df_sjt <-
  read_excel("./data/SJT.xlsx") %>%
  dplyr::select(UI06_05, # Participant
                SJT_AL_gek, # Monitoring
                SJT_ST_gek, # Managing momentum
                SJT_R_gek, # Rules and routines
                SJT_KF_gek) %>% # All
  filter(!UI06_05 %in% c(201, 223)) %>%  # exclude ID with no eye tracking data
  transmute(
    Participant = UI06_05,
    SJT_Monitoring = round(SJT_AL_gek, 2),
    SJT_All = round(SJT_KF_gek, 2),
  ) %>%
  arrange(Participant)

# classroom questionnaire
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(LI06_05, LM01_01:LM01_08, # classroom management
                LP01_01:LP01_08) %>% # non- & paraverbal communication
  filter(!LI06_05 %in% c(201, 223)) %>% # exclude ID with no eye tracking data
  rename(Participant = LI06_05) %>%
  group_by(Participant) %>%
  transmute(
    Participant = Participant,
    "Quest_All" = round(mean(c_across(starts_with(
      "L"
    )), na.rm = TRUE), 2),
    "Quest_classroom_management" = round(mean(c_across(starts_with(
      "LM"
    )), na.rm = TRUE), 2),
    "Quest_non_paraverbal_communication" = round(mean(c_across(starts_with(
      "LP"
    )), na.rm = TRUE), 2)
  )

# create a big data frame with all measures 
# put all data frames into list
df_list <- list(df_demo,
                df_aoi_numb_dur,
                df_aoi_stud,
                df_aoi_disrup,
                df_ttff_disrup,
                df_sri,
                df_sjt,
                df_quest)

# merge all data frames in list
df_merge <-
  df_list %>% reduce(full_join, by='Participant') %>%
  # filter(`Teaching Experience` > 0) %>% 
  dplyr::select(GRI_mtu,
                SJT_All,
                Quest_All,
                Mean_disruption_appraisal,
                Mean_confidence_appraisal
                )

# create a correlation matrix
cor_tab <- 
  df_merge %>% 
  cor(method = "pearson") %>% 
  round(., digits = 2)

# p-value for GRI x STJ_All
cor.test(df_merge$GRI_mtu, df_merge$SJT_All)

# p-value for GRI x Quest_All
cor.test(df_merge$GRI_mtu, df_merge$Quest_All)

# p-value for GRI x Disruption Rating
cor.test(df_merge$GRI_mtu, df_merge$Mean_disruption_appraisal)

# p-value for GRI x Confidence Rating
cor.test(df_merge$GRI_mtu, df_merge$Mean_confidence_appraisal)

### Teaching experience with only experts
# merge all data frames in list
df_merge_experts <-
  df_list %>% reduce(full_join, by='Participant') %>%
  filter(`Teaching Experience` > 0) %>% 
  dplyr::select(`Teaching Experience`,
                Stud_number_fixation_min,
                Average_duration_stud,
                Number_of_fixations.Disruptive_Person,
                Average_duration_disrup,
                Disrup_time_fixation_sec,
                GRI_mtu,
                SJT_All,
                SJT_Monitoring,
                Quest_All
                )

# create a correlation matrix
cor_tab_experts <- 
  df_merge_experts %>% 
  cor(method = "pearson") %>% 
  round(., digits = 2)

# p-value for GRI x Teaching Experience
cor.test(df_merge_experts$GRI_mtu, df_merge_experts$`Teaching Experience`)

# ### Only novices
# # merge all data frames in list
# df_merge_novices <-
#   df_list %>% reduce(full_join, by='Participant') %>%
#   filter(`Teaching Experience` == 0) %>% 
#   dplyr::select(`Teaching Experience`,
#                 Stud_number_fixation_min,
#                 Average_duration_stud,
#                 Number_of_fixations.Disruptive_Person,
#                 Average_duration_disrup,
#                 Disrup_time_fixation_sec,
#                 GRI_mtu,
#                 SJT_All,
#                 SJT_Monitoring,
#                 Quest_All
#                 )
# 
# # create a correlation matrix
# cor_tab_novices <- 
#   df_merge_novices %>% 
#   cor(method = "pearson") %>% 
#   round(., digits = 2)
# 
# # p-value for GRI x Teaching Experience
# cor.test(cor_tab_novices$GRI_mtu, cor_tab_novices$`Teaching Experience`)

```