### 2024-11-26 02_paper/02_study/02_expertise_paper/expertise_2024_11_18_MK.Rmd
x paper mwu test

---
title: "Through the Eyes of the Teacher - Multimodal Exploration of Expertise Differences in the Perception of Classroom Disruptions in a Laboratory Study"
output: 
  html_document:
    toc: true           # Add a table of contents
    number_sections: true
---

```{r include=FALSE}
# install.packages("needs")

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

library(needs)
needs(
  tidyverse,     # A collection of packages (e.g., ggplot2, dplyr, tidyr) for data manipulation, visualization, and general workflows.
  lubridate,     # For working with dates and times (e.g., parsing, manipulation, arithmetic).
  viridis,       # Provides colorblind-friendly color palettes for plots and visualizations.
  grid,          # A base R package for low-level grid graphics, useful for customizing complex plots.
  gridExtra,     # Extends grid by allowing multiple grid-based plots to be arranged into a single layout.
  cowplot,       # Simplifies creating publication-quality multi-panel plots, often combined with ggplot2.
  readxl,        # For reading Excel files (.xls, .xlsx) into R without requiring external dependencies.
  ARTofR,        # For applying aligned rank transform (ART) to factorial data for nonparametric analysis.
  moments,       # Provides functions to compute statistical moments (e.g., skewness, kurtosis).
  ltm,           # Used for latent trait models, including item response theory (IRT) analysis.
  sjPlot,        # A package to create publication-ready tables and visualizations for statistical models.
  kableExtra,    # Enhances the `knitr::kable()` function for producing polished and customizable tables in R Markdown.
  xtable,        # Converts R objects to LaTeX/HTML tables, especially useful for integration with Sweave/knitr.
  DescTools,     # A toolbox of descriptive statistics, effect size calculations, and data manipulation utilities.
  formattable,   # Allows creation of "formattable" data frames with customized formatting for reporting.
  psych,         # Provides tools for psychological research, including reliability analysis and descriptive statistics.
  apaTables,     # Simplifies the creation of APA-style tables for inclusion in papers or reports.
  rstatix,       # User-friendly pipe-compatible functions for common statistical tests and effect sizes.
  effsize,       # Computes effect sizes (e.g., Cohen's d) for t-tests and other statistical comparisons.
  knitr,         # Enables dynamic reporting by weaving R code and outputs into documents (e.g., R Markdown).
  papaja,        # Helps produce APA-style manuscripts directly from R Markdown.
  afex,          # Performs ANOVAs, including mixed-designs and generalized linear models, with a focus on user-friendliness.
  emmeans        # Computes estimated marginal means (or least-squares means) for post-hoc comparisons in models.
)

```

## Participants
```{r demographicstable, include = TRUE, echo=FALSE}

# Read and process the data
demo_data <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  filter(!LI06_05 %in% c(201, 223)) %>%
  transmute(
    Group = ifelse(LI06_05 < 200, "Novice", "Expert"),
    Gender = as.numeric(ifelse(LI02_01_1 == 2, "1", "0")),
    Age = LI03_01,
    `Teaching Experience` = LI04_01,
    School_type = factor(
      replace(LI18, LI18 == 2, 1),
      levels = c(1, 3, 4, 5, -1),
      labels = c("Primary School", "Secondary School", "Special Education", "Vocational School", "Other School Type")
    ),
    Semester = as.numeric(LI11_01),
    Internship = LI13_01,
    `Extracurricular Teaching Experience` = as.numeric(ifelse(LI14 == 1, "1", "0")),
    `Secondary Teaching Activities` = as.numeric(ifelse(LI17 == 1, "1", "0"))
  )

# Calculate summaries used across both tables, handling cases where all values are NA
summary_data <- demo_data %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    `Women in %` = ifelse(all(is.na(Gender)), NA, round(mean(Gender, na.rm = TRUE) * 100, 2)),
    `M Age` = ifelse(all(is.na(Age)), NA, round(mean(Age, na.rm = TRUE), 2)),
    `SD Age` = ifelse(all(is.na(Age)), NA, round(sd(Age, na.rm = TRUE), 2)),
    `Min Age` = ifelse(all(is.na(Age)), NA, min(Age, na.rm = TRUE)),
    `Max Age` = ifelse(all(is.na(Age)), NA, max(Age, na.rm = TRUE)),
    `M Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, round(mean(`Teaching Experience`, na.rm = TRUE), 2)),
    `SD Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, round(sd(`Teaching Experience`, na.rm = TRUE), 2)),
    `Median Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, median(`Teaching Experience`, na.rm = TRUE)),
    `Min Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, min(`Teaching Experience`, na.rm = TRUE)),
    `Max Teaching Exp.` = ifelse(all(is.na(`Teaching Experience`)), NA, max(`Teaching Experience`, na.rm = TRUE)),
    `M Semester` = ifelse(all(is.na(Semester)), NA, round(mean(Semester, na.rm = TRUE), 2)),
    `SD Semester` = ifelse(all(is.na(Semester)), NA, round(sd(Semester, na.rm = TRUE), 2)),
    `Min Semester` = ifelse(all(is.na(Semester)), NA, min(Semester, na.rm = TRUE)),
    `Max Semester` = ifelse(all(is.na(Semester)), NA, max(Semester, na.rm = TRUE)),
    `M Internship` = ifelse(all(is.na(Internship)), NA, round(mean(Internship, na.rm = TRUE), 2)),
    `SD Internship` = ifelse(all(is.na(Internship)), NA, round(sd(Internship, na.rm = TRUE), 2)),
    `Min Internship` = ifelse(all(is.na(Internship)), NA, min(Internship, na.rm = TRUE)),
    `Max Internship` = ifelse(all(is.na(Internship)), NA, max(Internship, na.rm = TRUE)),
    `Extracurricular Teaching Experience %` = ifelse(all(is.na(`Extracurricular Teaching Experience`)), NA, round(mean(`Extracurricular Teaching Experience`, na.rm = TRUE) * 100, 2)),
    `Secondary Teaching Activities %` = ifelse(all(is.na(`Secondary Teaching Activities`)), NA, round(mean(`Secondary Teaching Activities`, na.rm = TRUE) * 100, 2))
  )

# Separate the tables from the summarized data
demo_table <- summary_data %>%
  select(Group, N, `Women in %`, `M Age`, `SD Age`, `Min Age`, `Max Age`, 
         `M Teaching Exp.`, `SD Teaching Exp.`, `Median Teaching Exp.`, 
         `Min Teaching Exp.`, `Max Teaching Exp.`)

exp_table <- summary_data %>%
  select(Group, N, `M Teaching Exp.`, `SD Teaching Exp.`, `Median Teaching Exp.`, 
         `Min Teaching Exp.`, `Max Teaching Exp.`, `M Semester`, `SD Semester`, 
         `Min Semester`, `Max Semester`, `M Internship`, `SD Internship`, 
         `Min Internship`, `Max Internship`, `Extracurricular Teaching Experience %`, 
         `Secondary Teaching Activities %`)

# Display APA-style HTML tables
demo_table %>%
  kbl(caption = "Demographic Information & Teaching Experience", format = "html") %>%
  kable_styling(full_width = FALSE, position = "center")

exp_table %>%
  kbl(caption = "Experience Details by Group", format = "html") %>%
  kable_styling(full_width = FALSE, position = "center") %>% 
  scroll_box(width = "100%", height = "400px") # Enables scrolling if needed

```

## Measures

### Eye-Tracking Data
```{r eyetracking_read_in_data, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}

# Step 1: List all .tsv files in the "data" folder that contain "interval_complete"
file_names <- list.files(path = "data", pattern = "interval_complete.tsv", full.names = TRUE)

# Step 2: Read and bind all .tsv files into a single tibble
df_aoi <- file_names %>%
  map_dfr(~ read_tsv(.x, locale = locale(decimal_mark = ",")) %>%
            select(
              Group,
              Participant,
              TOI,
              Duration_of_interval,
              Start_of_interval,
              starts_with("Total_duration_of_fixations"),
              starts_with("Average_duration_of_fixations"),
              starts_with("Number_of_fixations"),
              starts_with("Time_to_first_fixation")
            ))

# Step 3: Extract only the participant's ID number (3 digits)
df_aoi <- df_aoi %>%
  mutate(Participant = as.numeric(str_extract(Participant, "\\d{3}")))

# Step 4: Exclude invalid participant ID 223
df_aoi <- df_aoi %>%
  filter(Participant != 223)

```
### Check variables

#### Letter search
```{r letter_search, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### LETTER SEARCH #####################
# 1. Filter and Prepare Data for "Letter Search"
df_letter <- df_aoi %>%
  filter(TOI == "Letter_search", !Participant %in% c(111, 205)) %>%  # Exclude participants with IDs 111 and 205
  select(TOI, Group, Participant, Duration_of_interval) %>%
  mutate(Duration_of_interval_sec = round(Duration_of_interval / 1000, 2))

# 2. Descriptive Statistics: N, Mean (M), SD, Min, Max for "Letter Search" Duration
letter_table <- df_letter %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Duration_of_interval_sec), 2),
    SD = round(sd(Duration_of_interval_sec), 2),
    Min = round(min(Duration_of_interval_sec), 2),
    Max = round(max(Duration_of_interval_sec), 2)
  )

# Display Descriptive Statistics Table in APA Style
knitr::kable(letter_table, caption = "N, M, SD, Min & Max Letter Search Duration (in seconds)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Conduct t-Test and Calculate Cohen's d for "Letter Search" Duration
t_test_letter <- t.test(
  Duration_of_interval_sec ~ Group,
  data = df_letter,
  var.equal = TRUE
)

# Calculate Cohen's d for the Effect Size
d_letter <- cohen.d(
  Duration_of_interval_sec ~ Group,
  data = df_letter
)

# 4. Format t-Test and Cohen's d Results for APA Table
t_test_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_letter$statistic, 2),
  "df" = t_test_letter$parameter,
  "p-value" = format.pval(t_test_letter$p.value, digits = 3, eps = .001),
  "Mean Difference" = round(t_test_letter$estimate[1] - t_test_letter$estimate[2], 2),
  "Cohen's d" = round(d_letter$estimate, 2),
  "95% CI (d)" = paste0("[", round(d_letter$conf.int[1], 2), ", ", round(d_letter$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size Results in APA Style
knitr::kable(t_test_result, caption = "t-Test and Effect Size for Letter Search Duration (Experts vs. Novices)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Percentage of fixations on each AOI, relative to the total number of fixations during the entire "Lesson" 
```{r nof_percent, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Step 1: Define AOI categories
aoi_columns <- list(
  "AOI_Students" = c(
    "Number_of_fixations.Anna",
    "Number_of_fixations.Bianca",
    "Number_of_fixations.Carl(a)"
  ),
  "AOI_Disruptive_Person" = c("Number_of_fixations.Disruptive_Person"),
  "AOI_Teacher_Material" = c(
    "Number_of_fixations.Board_Screen",
    "Number_of_fixations.Material_Teacher"
  ),
  "AOI_Student_Desk" = c(
    "Number_of_fixations.Nametag_Anna",
    "Number_of_fixations.Nametag_Bianca",
    "Number_of_fixations.Nametag_Carl(a)",
    "Number_of_fixations.Material_Students"
  ),
  "AOI_Classroom_Others" = c("Number_of_fixations.Classroom_Others")
)

# Step 2: Filter data for "Lesson" and select relevant columns
df_fixations <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(Participant, Group, starts_with("Number_of_fixations"))

# Step 3: Calculate total fixations per participant, excluding "Disruptive Person"
# Extract relevant column names excluding "Disruptive Person"
relevant_columns <- unlist(aoi_columns[c("AOI_Students", "AOI_Teacher_Material", "AOI_Student_Desk", "AOI_Classroom_Others")])

# Calculate row sums for the relevant columns and add as a new column
df_fixations <- df_fixations %>%
  mutate(Total_fixations_excluding_disruptive = rowSums(across(all_of(relevant_columns)), na.rm = TRUE))

# Step 4: Calculate fixation percentages for each AOI based on the adjusted total
aoi_percentages <- df_fixations %>%
  rowwise() %>%
  mutate(
    AOI_Students_pct = sum(c_across(any_of(
      aoi_columns$AOI_Students
    )), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100,
    AOI_Disruptive_Person_pct = sum(c_across(
      any_of(aoi_columns$AOI_Disruptive_Person)
    ), na.rm = TRUE) / (
      Total_fixations_excluding_disruptive + sum(c_across(
        any_of(aoi_columns$AOI_Disruptive_Person)
      ), na.rm = TRUE)
    ) * 100,
    AOI_Teacher_Material_pct = sum(c_across(
      any_of(aoi_columns$AOI_Teacher_Material)
    ), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100,
    AOI_Student_Desk_pct = sum(c_across(any_of(
      aoi_columns$AOI_Student_Desk
    )), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100,
    AOI_Classroom_Others_pct = sum(c_across(
      any_of(aoi_columns$AOI_Classroom_Others)
    ), na.rm = TRUE) / Total_fixations_excluding_disruptive * 100
  ) %>%
  ungroup()

# Step 5: Summarize fixation percentages by Group and arrange in descending order
grouped_aoi_percentages <- aoi_percentages %>%
  group_by(Group) %>%
  summarise(
    Mean_AOI_Students_pct = round(mean(AOI_Students_pct, na.rm = TRUE), 2),
    Mean_AOI_Disruptive_Person_pct = round(mean(AOI_Disruptive_Person_pct, na.rm = TRUE), 2),
    Mean_AOI_Teacher_Material_pct = round(mean(AOI_Teacher_Material_pct, na.rm = TRUE), 2),
    Mean_AOI_Student_Desk_pct = round(mean(AOI_Student_Desk_pct, na.rm = TRUE), 2),
    Mean_AOI_Classroom_Others_pct = round(mean(AOI_Classroom_Others_pct, na.rm = TRUE), 2)
  ) %>%
  pivot_longer(cols = starts_with("Mean_AOI"), names_to = "AOI", values_to = "Fixation_Percentage") %>%
  mutate(AOI = recode(AOI,
                      "Mean_AOI_Students_pct" = "Students",
                      "Mean_AOI_Disruptive_Person_pct" = "Disruptive Person",
                      "Mean_AOI_Teacher_Material_pct" = "Teacher Material",
                      "Mean_AOI_Student_Desk_pct" = "Student Desk",
                      "Mean_AOI_Classroom_Others_pct" = "Classroom/Others")) %>%
  arrange(Group, desc(Fixation_Percentage))

# # Step 6: Display the table with APA style
# kable(grouped_aoi_percentages, 
#       col.names = c("AOI", "Expert (%)", "Novice (%)"),
#       caption = "Average Fixation Percentage per AOI by Group (Novice vs. Expert) - Descending Order") %>%
#   kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
#   add_header_above(c(" " = 1, "Fixation Percentage" = 2)) %>%
#   row_spec(0, bold = TRUE) %>%
#   column_spec(2:3, width = "3em", border_left = TRUE)

# Step 7: Plot with descending AOI order, y-axis limit set to 100%, and percentage labels inside the bars
ggplot(grouped_aoi_percentages, aes(x = reorder(AOI, -Fixation_Percentage), y = Fixation_Percentage, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = paste0(Fixation_Percentage, "%")), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3, size = 3) +
  labs(
    title = "Average Fixation Percentage by AOI and Group (Descending Order)",
    x = "AOI",
    y = "Fixation Percentage (%)"
  ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(title = "Group")) +
  ylim(0, 100)  # Set y-axis limit to 100%

# Step 8: Perform a t-test between Experts and Novices on the percentage values and calculate Cohen's d
# Define the AOIs you want to analyze with names matching those generated in Step 4
aoi_list <- c("AOI_Students_pct", "AOI_Disruptive_Person_pct", "AOI_Teacher_Material_pct", "AOI_Student_Desk_pct", "AOI_Classroom_Others_pct")

# Initialize an empty data frame to store the results
results <- data.frame()

# Loop over each AOI and calculate t-test and Cohen's d
for (column_name in aoi_list) {

  # Ensure that the column exists in the data
  if (column_name %in% names(aoi_percentages)) {

    # Filter the data for the current AOI and rename the column to Fixation_Percentage for consistency
    aoi_data <- aoi_percentages %>%
      select(Group, !!sym(column_name)) %>%
      rename(Fixation_Percentage = !!sym(column_name))

    # Separate data for each group
    expert_data <- aoi_data$Fixation_Percentage[aoi_data$Group == "Expert"]
    novice_data <- aoi_data$Fixation_Percentage[aoi_data$Group == "Novice"]

    # Check if both groups have at least one observation
    if (length(expert_data) > 0 && length(novice_data) > 0) {

      # Perform a t-test between Experts and Novices on the current AOI's fixation percentage
      t_test_result <- t.test(expert_data, novice_data, var.equal = TRUE)

      # Calculate Cohen's d for the current AOI
      cohen_d_value <- cohen.d(expert_data, novice_data, pooled = TRUE)$estimate

      # Format the results for the current AOI and add to the results data frame
      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),  # Remove "_pct" suffix for readability
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = round(t_test_result$statistic, 2),
          "df" = t_test_result$parameter,
          "p-value" = format.pval(t_test_result$p.value, digits = 3, eps = .001),
          "Mean Difference" = round(t_test_result$estimate[1] - t_test_result$estimate[2], 2),
          "Cohen's d" = round(cohen_d_value, 2),
          "95% CI (d)" = paste0("[", round(t_test_result$conf.int[1], 2), ", ", round(t_test_result$conf.int[2], 2), "]")
        )
      )

    } else {
      # If one group has no data, append N/A values for this AOI
      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),  # Remove "_pct" suffix for readability
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = NA,
          "df" = NA,
          "p-value" = NA,
          "Mean Difference" = NA,
          "Cohen's d" = NA,
          "95% CI (d)" = NA
        )
      )
    }
  }
}

# Clean up AOI names for readability in the final table
aoi_name_mapping <- c(
  "AOI_Students" = "Students",
  "AOI_Disruptive_Person" = "Disruptive Person",
  "AOI_Teacher_Material" = "Teacher Material",
  "AOI_Student_Desk" = "Student Desk",
  "AOI_Classroom_Others" = "Classroom/Others"
)
results$AOI <- aoi_name_mapping[results$AOI]

# Display the combined results table in APA style without row names
knitr::kable(results, caption = "t-Test and Effect Size for Fixation Percentage across AOIs (Experts vs. Novices)", row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Number of fixations per minute (micro-teaching unit)
```{r nof_all, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATIONS PER MINUTE IN MICRO-TEACHING UNIT ########################

# 1. Data Preparation: Filter and Calculate Fixations per Minute in Micro-Teaching Unit
df_aoi_sum <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Group,
    Participant,
    Duration_of_interval,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations"),
    starts_with("Average_duration"),
    -contains("Disruptive_Person")
  ) %>%
  rowwise() %>%
  transmute(
    Group = Group,
    Participant = Participant,
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Sum_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Sum_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Number_fixation_min_mtu = round(Sum_number_fixation / Duration_of_interval_min, 2),
    Group = as_factor(Group)
  ) %>%
  drop_na()

# 2. Descriptive Statistics Table for Number of Fixations per Minute
nof_table <- df_aoi_sum %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Number_fixation_min_mtu), 2),
    SD = round(sd(Number_fixation_min_mtu), 2),
    Min = round(min(Number_fixation_min_mtu), 2),
    Max = round(max(Number_fixation_min_mtu), 2)
  )

# Display the descriptive table in APA style
knitr::kable(nof_table, caption = "N, M, SD, Min & Max Number of Fixations per Minute (Micro-Teaching Unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Calculate means and standard deviations for Novice and Expert
mean_novice_mtu <- round(mean(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Novice"], na.rm = TRUE), 2)
sd_novice_mtu <- round(sd(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Novice"], na.rm = TRUE), 2)
mean_expert_mtu <- round(mean(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Expert"], na.rm = TRUE), 2)
sd_expert_mtu <- round(sd(df_aoi_sum$Number_fixation_min_mtu[df_aoi_sum$Group == "Expert"], na.rm = TRUE), 2)

# 4. Perform t-test to get the p-value
t_test_mtu <- t.test(
  Number_fixation_min_mtu ~ Group,
  data = df_aoi_sum,
  var.equal = TRUE
)
p_value_mtu <- round(t_test_mtu$p.value, 3)

# 5. Calculate Cohen's d for effect size using formula syntax
library(effsize)
cohen_d_mtu <- round(cohen.d(Number_fixation_min_mtu ~ Group, data = df_aoi_sum)$estimate, 2)

# 6. Plotting the Number of Fixations per Minute (Micro-Teaching Unit)
y_max_mtu <- max(df_aoi_sum$Number_fixation_min_mtu) * 1.5

plot_number_group <- df_aoi_sum %>%
  ggplot(mapping = aes(x = Group, y = Number_fixation_min_mtu)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_mtu) +
  labs(x = "", y = "Number of Fixations per Minute") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Number of Fixations per Minute\n(Micro-Teaching Unit)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  # Annotations for mean (M) and SD using atop() to place SD below M
  annotate("text", x = 1, y = y_max_mtu * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice_mtu, ", italic(SD) == ", sd_novice_mtu, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max_mtu * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert_mtu, ", italic(SD) == ", sd_expert_mtu, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Add p-value and Cohen's d annotations above the boxplots
  annotate("text", x = 1.5, y = y_max_mtu * 0.95,
           label = paste0("p = ", p_value_mtu, "\nd = ", cohen_d_mtu),
           hjust = 0.5, size = 4, fontface = "italic")

plot_number_group

# 7. Format the t-Test and Cohen's d results for APA Table
t_test_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_mtu$statistic, 2),
  "df" = t_test_mtu$parameter,
  "p-value" = format.pval(t_test_mtu$p.value, digits = 3, eps = .001),
  "Mean Difference" = round(t_test_mtu$estimate[1] - t_test_mtu$estimate[2], 2),
  "Cohen's d" = cohen_d_mtu,
  "95% CI (d)" = paste0("[", round(t_test_mtu$conf.int[1], 2), ", ", round(t_test_mtu$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size results in APA style
knitr::kable(t_test_result, caption = "t-Test and Effect Size for Number of Fixations per Minute (Micro-Teaching Unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Number of fixations per minute (AOI students)
```{r nof_students, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATIONS (AOI STUDENTS) ########################

# 1. Data Preparation: Filter and Calculate Fixations per Minute on AOI Students
df_aoi_stud <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select (
    Group,
    Duration_of_interval,
    Participant,
    "Total_duration_of_fixations.Anna",
    "Total_duration_of_fixations.Bianca",
    "Total_duration_of_fixations.Carl(a)",
    "Number_of_fixations.Anna",
    "Number_of_fixations.Bianca",
    "Number_of_fixations.Carl(a)"
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Participant = Participant,
    Stud_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Stud_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Stud_number_fixation_min = round(Stud_number_fixation / Duration_of_interval_min, 2),
    Stud_number_fixation_sec = round(Stud_number_fixation / Duration_of_interval_sec, 2),
    Average_duration_stud = round(Stud_duration_fixation / Stud_number_fixation, 0)
  ) %>%
  drop_na()

# 2. Descriptive Statistics: N, Mean (M), SD, Min, Max for Fixations per Minute (AOI Students)
nof_min_stud_table <- df_aoi_stud %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Stud_number_fixation_min), 2),
    SD = round(sd(Stud_number_fixation_min), 2),
    Min = round(min(Stud_number_fixation_min), 2),
    Max = round(max(Stud_number_fixation_min), 2)
  )

# Display Descriptive Statistics Table in APA Style
knitr::kable(nof_min_stud_table, caption = "N, M, SD, Min & Max Number of Fixations per Minute (AOI Students)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Plotting the Number of Fixations per Minute on AOI Students
# Calculate descriptive statistics
summary_stats <- df_aoi_stud %>%
  group_by(Group) %>%
  summarise(
    M = round(mean(Stud_number_fixation_min), 2),
    SD = round(sd(Stud_number_fixation_min), 2)
  )

# Conduct t-test and calculate effect size
t_test_stud <- t.test(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud,
  var.equal = TRUE
)

# Cohen's d for effect size
d_number_stud <- cohen.d(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud
)

# Extract statistical values
mean_expert <- summary_stats %>% filter(Group == "Expert") %>% pull(M)
sd_expert <- summary_stats %>% filter(Group == "Expert") %>% pull(SD)
mean_novice <- summary_stats %>% filter(Group == "Novice") %>% pull(M)
sd_novice <- summary_stats %>% filter(Group == "Novice") %>% pull(SD)
p_value <- format.pval(t_test_stud$p.value, digits = 3, eps = .001)
cohen_d <- round(d_number_stud$estimate, 2)

# Determine the maximum y-value for ylim to ensure all data points and annotations are visible
y_max <- max(df_aoi_stud$Stud_number_fixation_min) * 1.5  # Increase the multiplier if needed

# Create the boxplot with jittered data points for better readability
plot_number_min_stud <- df_aoi_stud %>%
  ggplot(mapping = aes(x = Group, y = Stud_number_fixation_min)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0, alpha = 0.6, color = "black", size = 1.2
  ) +
  labs(x = "", y = "Number of fixations per minute") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Number of fixations per minute\n(AOI students)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  ylim(0, y_max) +  # Adjust y-axis limit to include annotations
  # Annotations for mean (M) and SD using atop() to place SD below M
  annotate("text", x = 1, y = y_max * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice, ", italic(SD) == ", sd_novice, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert, ", italic(SD) == ", sd_expert, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Add p-value and Cohen's d annotations above the boxplots
  annotate("text", x = 1.5, y = y_max * 0.95,
           label = paste0("p = ", p_value, "\nd = ", cohen_d),
           hjust = 0.5, size = 4, fontface = "italic")

# Display the plot
plot_number_min_stud

# 4. t-Test and Effect Size Calculation for Number of Fixations per Minute on AOI Students
t_test_stud <- t.test(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud,
  var.equal = TRUE
)

# Calculate Cohen's d for the Effect Size
d_number_stud <- cohen.d(
  Stud_number_fixation_min ~ Group,
  data = df_aoi_stud
)

# 5. Format t-Test and Cohen's d Results for APA Table
t_test_result_stud <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_stud$statistic, 2),
  "df" = t_test_stud$parameter,
  "p-value" = format.pval(t_test_stud$p.value, digits = 3, eps = .001),
  "Mean Difference" = round(t_test_stud$estimate[1] - t_test_stud$estimate[2], 2),
  "Cohen's d" = round(d_number_stud$estimate, 2),
  "95% CI (d)" = paste0("[", round(d_number_stud$conf.int[1], 2), ", ", round(d_number_stud$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size Results in APA Style
knitr::kable(t_test_result_stud, caption = "t-Test and Effect Size for Number of Fixations per Minute (AOI Students)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Number of fixations per seconds (AOI disruptive person)
```{r nof_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATIONS ON AOI DISRUPTIVE PERSON ########################

# Define the different types of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Step 1: Filter and prepare the data
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>% # Filter rows based on the disruption list
  select(Group, Participant, TOI, Duration_of_interval, "Number_of_fixations.Disruptive_Person") %>%
  mutate(
    Group = as_factor(Group), # Convert Group column to a factor
    Disruption_Category = case_when( # Assign categories based on disruption type
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2) # Convert duration to seconds
  ) %>%
  filter(!is.na(Disruption_Category)) # Keep only rows with valid categories

# Step 2: Summarize data by disruption category
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarise(
    total_duration_sec = round(sum(Duration_of_interval_sec, na.rm = TRUE), 2), # Total time in seconds
    fixation_count = round(sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE), 2), # Total fixations
    fixations_per_second = round(fixation_count / total_duration_sec, 2), # Fixation rate per second
    .groups = "drop"
  )

# Step 3: Summarize by participant
individual_summary <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarise(
    mean_fixations_per_second = round(mean(fixations_per_second, na.rm = TRUE), 2), # Mean fixation rate
    .groups = "drop"
  )

##################### DESCRIPTIVE STATISTICS ########################

# Descriptive statistics for overall fixation rate
overall_stats <- individual_summary %>%
  group_by(Group) %>%
  summarise(
    N = n(), # Number of participants
    Mean = round(mean(mean_fixations_per_second, na.rm = TRUE), 2), # Mean fixation rate
    SD = round(sd(mean_fixations_per_second, na.rm = TRUE), 2), # Standard deviation
    Min = round(min(mean_fixations_per_second, na.rm = TRUE), 2), # Minimum value
    Max = round(max(mean_fixations_per_second, na.rm = TRUE), 2), # Maximum value
    .groups = "drop"
  )

# Display the overall descriptive statistics table
kable(
  overall_stats,
  caption = "Descriptive Statistics (N, Mean, SD, Min, Max) for Overall Fixation Rate by Group",
  col.names = c("Group", "N", "M", "SD", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Descriptive statistics by disruption category
category_stats <- disruption_summary %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n(), # Number of data points
    Mean = round(mean(fixations_per_second, na.rm = TRUE), 2), # Mean fixation rate
    SD = round(sd(fixations_per_second, na.rm = TRUE), 2), # Standard deviation
    Min = round(min(fixations_per_second, na.rm = TRUE), 2), # Minimum value
    Max = round(max(fixations_per_second, na.rm = TRUE), 2), # Maximum value
    .groups = "drop"
  )

# Display the descriptive statistics table by category
kable(
  category_stats,
  caption = "Descriptive Statistics (N, Mean, SD, Min, Max) for Fixation Rate by Group and Category",
  col.names = c("Group", "Category", "N", "M", "SD", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### PLOTS ########################
##################### FIRST PLOT: FIXATION RATE OVERALL ########################

# Calculate values for annotations
mean_novice <- round(mean(individual_summary$mean_fixations_per_second[individual_summary$Group == "Novice"], na.rm = TRUE), 2)
sd_novice <- round(sd(individual_summary$mean_fixations_per_second[individual_summary$Group == "Novice"], na.rm = TRUE), 2)
mean_expert <- round(mean(individual_summary$mean_fixations_per_second[individual_summary$Group == "Expert"], na.rm = TRUE), 2)
sd_expert <- round(sd(individual_summary$mean_fixations_per_second[individual_summary$Group == "Expert"], na.rm = TRUE), 2)

# t-Test and effect size
t_test <- t.test(mean_fixations_per_second ~ Group, data = individual_summary, var.equal = TRUE)
p_value <- round(t_test$p.value, 3)
cohen_d <- round(cohen.d(mean_fixations_per_second ~ Group, data = individual_summary)$estimate, 2)

# Overall fixation rate plot with annotations
y_max_overall <- max(individual_summary$mean_fixations_per_second, na.rm = TRUE) * 1.5
plot_overall <- ggplot(individual_summary, aes(x = Group, y = mean_fixations_per_second, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.7, color = "black", size = 1.5) +
  labs(x = "", y = "Mean Fixations per Second per Participant") +
  ggtitle("Overall Fixation Rate per Participant\n(AOI Disruptive Person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  ylim(0, y_max_overall) +
  theme(legend.position = "none") +
  annotate("text", x = 1, y = y_max_overall * 0.85,
           label = paste0("M = ", mean_novice, "\nSD = ", sd_novice),
           hjust = 0.5, size = 4, fontface = "italic") +
  annotate("text", x = 2, y = y_max_overall * 0.85,
           label = paste0("M = ", mean_expert, "\nSD = ", sd_expert),
           hjust = 0.5, size = 4, fontface = "italic") +
  annotate("text", x = 1.5, y = y_max_overall * 0.95,
           label = paste0("p = ", p_value, "\nd = ", cohen_d),
           hjust = 0.5, size = 4, fontface = "italic")

##################### SECOND PLOT: FIXATION RATE BY CATEGORY ########################

# # Prepare annotation data for labels
# annotation_data <- category_stats_with_tests %>%
#   left_join(y_max_by_category, by = "Disruption_Category") %>%
#   mutate(
#     novice_label = paste0("italic(M) == ", M_Novice, " ~ ", "italic(SD) == ", SD_Novice),
#     expert_label = paste0("italic(M) == ", M_Expert, " ~ ", "italic(SD) == ", SD_Expert),
#     overall_label = paste0("italic(p) == ", p_value, " ~ ", "italic(d) == ", cohen_d)
#   )

# Create the plot
plot_category <- disruption_summary_annotated %>%
  ggplot(mapping = aes(x = Group, y = fixations_per_second, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "fixed", labeller = label_wrap_gen(width = 15)) + # Fixed y-axis across facets
  labs(x = "", y = "Fixations per Second") +
  ggtitle("Fixation Rate by Disruption Category\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10) # Adjust facet text size if necessary
  ) 
# +
#   # Annotate novice statistics
#   geom_text(
#     data = annotation_data,
#     aes(x = 1, y = y_max, label = novice_label),
#     inherit.aes = FALSE,
#     parse = TRUE, size = 3.5, hjust = 0.5, vjust = 0
#   ) +
#   # Annotate expert statistics
#   geom_text(
#     data = annotation_data,
#     aes(x = 2, y = y_max, label = expert_label),
#     inherit.aes = FALSE,
#     parse = TRUE, size = 3.5, hjust = 0.5, vjust = 0
#   ) +
#   # Annotate p-value and Cohen's d
#   geom_text(
#     data = annotation_data,
#     aes(x = 1.5, y = y_max * 1.05, label = overall_label),
#     inherit.aes = FALSE,
#     parse = TRUE, size = 4, fontface = "italic", hjust = 0.5, vjust = 0
#   )

# Display the both plots
plot_overall
plot_category

##################### STATISTICAL TESTS ########################

# Table summarizing overall t-test and Cohen's d
overall_effect <- data.frame(
  Group = c("Overall"),
  `t-value` = round(t_test$statistic, 2),
  `p-value` = p_value,
  `Cohen's d` = cohen_d
)

# Display t-test and effect size table
kable(
  overall_effect,
  caption = "Overall t-Test and Effect Size for Fixation Rate",
  col.names = c("Group", "t-value", "p-value", "Cohen's d"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### REPEATED MEASURES ANOVA ########################

# Perform the repeated-measures ANOVA
anova_afex <- afex::aov_ez(
  id = "Participant", # Random effect
  dv = "fixations_per_second", # Dependent variable
  between = "Group", # Between-subject factor
  within = "Disruption_Category", # Within-subject factor
  data = anova_data
)

# Extract the ANOVA table
anova_results <- as.data.frame(anova_afex$anova_table)

# Recalculate generalized eta squared (η²)
eta_squared_results <- effectsize::eta_squared(anova_afex, partial = FALSE)

# Combine recalculated eta squared with the ANOVA table
anova_table_apa <- anova_results %>%
  rownames_to_column(var = "Effect") %>% # Move row names to a column
  left_join(
    as.data.frame(eta_squared_results), 
    by = c("Effect" = "Parameter") # Match the correct columns
  ) %>%
  mutate(
    `F` = round(`F`, 2), # Round F-values
    `η²` = round(Eta2, 3), # Round generalized eta squared
    `p-value` = ifelse(
      `Pr(>F)` < 0.05, "< .05", round(`Pr(>F)`, 3) # Format p-values
    )
  ) %>%
  select(Effect, `num Df`, `den Df`, `F`, `p-value`, `η²`) # Order columns, place η² last

# Display the table
kable(
  anova_table_apa,
  caption = "Repeated-Measures ANOVA Results (Including Effect Sizes)",
  col.names = c("Effect", "Df1", "Df2", "F", "p", "η²"),
  digits = 2
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Post-hoc comparisons for Disruption_Category
  post_hoc_disruption <- emmeans(anova_afex, pairwise ~ Disruption_Category)

  # Extract pairwise contrasts
  post_hoc_results <- as.data.frame(post_hoc_disruption$contrasts)

  # Clean up category names in the contrast column
  post_hoc_results$contrast <- gsub("_", " ", post_hoc_results$contrast) # Replace underscores with spaces
  post_hoc_results$contrast <- gsub("\\.", " ", post_hoc_results$contrast) # Replace dots with spaces

  # Calculate Cohen's d for each pairwise comparison using pooled SD from emmeans
  post_hoc_results <- post_hoc_results %>%
    mutate(
      pooled_sd = sqrt((SE^2 * 2)), # Approximation for pooled SD from SE
      d = round(estimate / pooled_sd, 2), # Calculate Cohen's d
      p.value = ifelse(
        p.value < 0.05, "< .05", round(p.value, 3)
      ),
      t = round(t.ratio, 2),
      df = round(df, 2),
      Estimate = round(estimate, 2),
      SE = round(SE, 2)
    ) %>%
    select(contrast, Estimate, SE, df, t, p.value, d) # Include Cohen's d

  # Display APA-styled post-hoc table
  kable(
    post_hoc_results,
    caption = "Post-Hoc Pairwise Comparisons for Disruption Categories (Including Effect Sizes)",
    col.names = c("Comparison", "Estimate", "SE", "df", "t", "p", "Cohen's d"),
    digits = 2
  ) %>%
    kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

### Duration of disruption
```{r dur_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### DATA PREPARATION ########################

# Prepare data for analysis: filter, calculate, and categorize disruptions
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions, Duration_of_interval <= 30000) %>%
  select(Group, Participant, TOI, Duration_of_interval) %>%
  mutate(
    Group = as_factor(Group),
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2)
  ) %>%
  filter(!is.na(Disruption_Category)) %>%
  drop_na()

##################### DESCRIPTIVE STATISTICS ########################

# Descriptive Statistics by Group
summary_stats_group <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    SD = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    Median = round(median(Duration_of_interval_sec, na.rm = TRUE), 2),
    IQR = round(IQR(Duration_of_interval_sec, na.rm = TRUE), 2),
    Min = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    Max = round(max(Duration_of_interval_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Descriptive Statistics by Category and Group
summary_stats_category_group <- df_aoi_disrup %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n(),
    Mean = round(mean(Duration_of_interval_sec, na.rm = TRUE), 2),
    SD = round(sd(Duration_of_interval_sec, na.rm = TRUE), 2),
    Median = round(median(Duration_of_interval_sec, na.rm = TRUE), 2),
    IQR = round(IQR(Duration_of_interval_sec, na.rm = TRUE), 2),
    Min = round(min(Duration_of_interval_sec, na.rm = TRUE), 2),
    Max = round(max(Duration_of_interval_sec, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Display Descriptive Tables
kable(
  summary_stats_group,
  caption = "Descriptive Statistics for Disruption Duration by Group",
  col.names = c("Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

kable(
  summary_stats_category_group,
  caption = "Descriptive Statistics for Disruption Duration by Category and Group",
  col.names = c("Category", "Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### PLOTS ########################

# Plot 1: Disruption Duration by Group
plot_duration_group <- ggplot(df_aoi_disrup, aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  labs(x = "Group", y = "Disruption Duration (seconds)") +
  theme_cowplot() +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Duration by Group") +
  theme(legend.position = "none") +
  annotate("text", x = 1, y = max(df_aoi_disrup$Duration_of_interval_sec) + 4,
           label = paste0("Median = ", summary_stats_group$Median[summary_stats_group$Group == "Novice"],
                          "\nIQR = ", summary_stats_group$IQR[summary_stats_group$Group == "Novice"]),
           hjust = 0.5, size = 4) +
  annotate("text", x = 2, y = max(df_aoi_disrup$Duration_of_interval_sec) + 4,
           label = paste0("Median = ", summary_stats_group$Median[summary_stats_group$Group == "Expert"],
                          "\nIQR = ", summary_stats_group$IQR[summary_stats_group$Group == "Expert"]),
           hjust = 0.5, size = 4)

plot_duration_group

# Plot 2: Disruption Duration by Group and Category
plot_duration_category <- ggplot(df_aoi_disrup, aes(x = Group, y = Duration_of_interval_sec, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, alpha = 0.5, color = "black", size = 1.5) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) +
  labs(x = "Group", y = "Disruption Duration (seconds)") +
  theme_cowplot() +
  ylim(0, 30) + 
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Duration by Group and Category") +
  theme(legend.position = "none")

plot_duration_category

##################### STATISTICAL TESTS ########################

# Normality Tests (Shapiro-Wilk)
normality_tests <- df_aoi_disrup %>%
  group_by(Group) %>%
  summarise(
    Shapiro_Wilk_p = round(shapiro.test(Duration_of_interval_sec)$p.value, 3),
    .groups = "drop"
  )

kable(
  normality_tests,
  caption = "Shapiro-Wilk Test for Normality by Group",
  col.names = c("Group", "Shapiro-Wilk p-value"),
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Mann-Whitney U-Test and Effect Size
mann_whitney_test <- wilcox.test(Duration_of_interval_sec ~ Group, data = df_aoi_disrup, exact = FALSE)

# Effect Size Calculation for Mann-Whitney (r)
n1 <- nrow(df_aoi_disrup %>% filter(Group == "Novice"))
n2 <- nrow(df_aoi_disrup %>% filter(Group == "Expert"))
r_effect <- abs(mann_whitney_test$statistic - (n1 * n2 / 2)) / sqrt((n1 * n2 * (n1 + n2 + 1)) / 12)

mann_whitney_results <- data.frame(
  "Test" = "Mann-Whitney U",
  "W" = round(mann_whitney_test$statistic, 2),
  "p-value" = round(mann_whitney_test$p.value, 3),
  "Effect Size (r)" = round(r_effect, 3)
)

kable(
  mann_whitney_results,
  caption = "Mann-Whitney U-Test Results for Disruption Duration",
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Kruskal-Wallis Test and Pairwise Comparisons
kruskal_test <- kruskal.test(Duration_of_interval_sec ~ Disruption_Category, data = df_aoi_disrup)

kruskal_results <- data.frame(
  "Test" = "Kruskal-Wallis",
  "Chi-squared" = round(kruskal_test$statistic, 2),
  "df" = kruskal_test$parameter,
  "p-value" = round(kruskal_test$p.value, 3)
)

kable(
  kruskal_results,
  caption = "Kruskal-Wallis Test Results for Differences Between Disruption Categories",
  row.names = FALSE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Pairwise Wilcoxon test with Bonferroni correction
pairwise_wilcox <- pairwise.wilcox.test(
  x = df_aoi_disrup$Duration_of_interval_sec,
  g = df_aoi_disrup$Disruption_Category,
  p.adjust.method = "bonferroni"
)

# Extract and round the p-values to 3 decimal places
pairwise_results <- as.data.frame(pairwise_wilcox$p.value) %>%
  mutate(across(everything(), ~ round(., 3)))

# Display the results with rounded p-values
kable(
  pairwise_results,
  caption = "Pairwise Wilcoxon Test Results (Bonferroni Corrected)",
  row.names = TRUE
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Percentage of fixation durations on each AOI, relative to the total duration of fixations during the entire "Lesson" 
```{r dur_fix_percent, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Step 1: Define AOI categories with correct column names for fixation duration
aoi_columns <- list(
  "AOI_Students" = c(
    "Total_duration_of_fixations.Anna",
    "Total_duration_of_fixations.Bianca",
    "Total_duration_of_fixations.Carl(a)"
  ),
  "AOI_Disruptive_Person" = c("Total_duration_of_fixations.Disruptive_Person"),
  "AOI_Teacher_Material" = c(
    "Total_duration_of_fixations.Board_Screen",
    "Total_duration_of_fixations.Material_Teacher"
  ),
  "AOI_Student_Desk" = c(
    "Total_duration_of_fixations.Nametag_Anna",
    "Total_duration_of_fixations.Nametag_Bianca",
    "Total_duration_of_fixations.Nametag_Carl(a)",
    "Total_duration_of_fixations.Material_Students"
  ),
  "AOI_Classroom_Others" = c("Total_duration_of_fixations.Classroom_Others")
)

# Step 2: Filter data for "Lesson" and select relevant columns for fixation duration
df_durations <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(Participant, Group, starts_with("Total_duration_of_fixations"))

# Step 3: Calculate total fixation duration per participant, excluding "Disruptive Person"
relevant_columns <- unlist(aoi_columns[c("AOI_Students", "AOI_Teacher_Material", "AOI_Student_Desk", "AOI_Classroom_Others")])

df_durations <- df_durations %>%
  mutate(Total_duration_excluding_disruptive = rowSums(across(all_of(relevant_columns)), na.rm = TRUE))

# Step 4: Calculate fixation duration percentages for each AOI based on the adjusted total duration
aoi_durations_percentages <- df_durations %>%
  rowwise() %>%
  mutate(
    AOI_Students_pct = sum(c_across(any_of(aoi_columns$AOI_Students)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100,
    AOI_Disruptive_Person_pct = sum(c_across(any_of(aoi_columns$AOI_Disruptive_Person)), na.rm = TRUE) / (
      Total_duration_excluding_disruptive + sum(c_across(any_of(aoi_columns$AOI_Disruptive_Person)), na.rm = TRUE)
    ) * 100,
    AOI_Teacher_Material_pct = sum(c_across(any_of(aoi_columns$AOI_Teacher_Material)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100,
    AOI_Student_Desk_pct = sum(c_across(any_of(aoi_columns$AOI_Student_Desk)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100,
    AOI_Classroom_Others_pct = sum(c_across(any_of(aoi_columns$AOI_Classroom_Others)), na.rm = TRUE) / Total_duration_excluding_disruptive * 100
  ) %>%
  ungroup()

# Step 5: Summarize fixation duration percentages by Group and arrange in descending order
grouped_aoi_durations_percentages <- aoi_durations_percentages %>%
  group_by(Group) %>%
  summarise(
    Mean_AOI_Students_pct = round(mean(AOI_Students_pct, na.rm = TRUE), 2),
    Mean_AOI_Disruptive_Person_pct = round(mean(AOI_Disruptive_Person_pct, na.rm = TRUE), 2),
    Mean_AOI_Teacher_Material_pct = round(mean(AOI_Teacher_Material_pct, na.rm = TRUE), 2),
    Mean_AOI_Student_Desk_pct = round(mean(AOI_Student_Desk_pct, na.rm = TRUE), 2),
    Mean_AOI_Classroom_Others_pct = round(mean(AOI_Classroom_Others_pct, na.rm = TRUE), 2)
  ) %>%
  pivot_longer(cols = starts_with("Mean_AOI"), names_to = "AOI", values_to = "Duration_Percentage") %>%
  mutate(AOI = recode(AOI,
                      "Mean_AOI_Students_pct" = "Students",
                      "Mean_AOI_Disruptive_Person_pct" = "Disruptive Person",
                      "Mean_AOI_Teacher_Material_pct" = "Teacher Material",
                      "Mean_AOI_Student_Desk_pct" = "Student Desk",
                      "Mean_AOI_Classroom_Others_pct" = "Classroom/Others")) %>%
  arrange(Group, desc(Duration_Percentage))

# Step 6: Plot the fixation duration percentages by AOI and Group
ggplot(grouped_aoi_durations_percentages, aes(x = reorder(AOI, -Duration_Percentage), y = Duration_Percentage, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = paste0(Duration_Percentage, "%")), 
            position = position_dodge(width = 0.8), 
            vjust = -0.3, size = 3) +
  labs(
    title = "Total Fixation Duration Percentage by AOI and Group (Descending Order)",
    x = "AOI",
    y = "Fixation Duration Percentage (%)"
  ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(title = "Group")) +
  ylim(0, 100)  # Set y-axis limit to 100%

# Step 7: Perform a t-test between Experts and Novices on the duration percentages and calculate Cohen's d
aoi_list <- c("AOI_Students_pct", "AOI_Disruptive_Person_pct", "AOI_Teacher_Material_pct", "AOI_Student_Desk_pct", "AOI_Classroom_Others_pct")
results <- data.frame()

for (column_name in aoi_list) {
  if (column_name %in% names(aoi_durations_percentages)) {
    aoi_data <- aoi_durations_percentages %>%
      select(Group, !!sym(column_name)) %>%
      rename(Duration_Percentage = !!sym(column_name))

    expert_data <- aoi_data$Duration_Percentage[aoi_data$Group == "Expert"]
    novice_data <- aoi_data$Duration_Percentage[aoi_data$Group == "Novice"]

    if (length(expert_data) > 0 && length(novice_data) > 0) {
      t_test_result <- t.test(expert_data, novice_data, var.equal = TRUE)
      cohen_d_value <- cohen.d(expert_data, novice_data, pooled = TRUE)$estimate

      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = round(t_test_result$statistic, 2),
          "df" = t_test_result$parameter,
          "p-value" = format.pval(t_test_result$p.value, digits = 3, eps = .001),
          "Mean Difference" = round(t_test_result$estimate[1] - t_test_result$estimate[2], 2),
          "Cohen's d" = round(cohen_d_value, 2),
          "95% CI (d)" = paste0("[", round(t_test_result$conf.int[1], 2), ", ", round(t_test_result$conf.int[2], 2), "]")
        )
      )
    } else {
      results <- rbind(
        results,
        data.frame(
          "AOI" = gsub("_pct", "", column_name),
          "Group Comparison" = "Experts vs. Novices",
          "t-value" = NA,
          "df" = NA,
          "p-value" = NA,
          "Mean Difference" = NA,
          "Cohen's d" = NA,
          "95% CI (d)" = NA
        )
      )
    }
  }
}

# Map AOI names to clean versions for readability in the final results table
results$AOI <- recode(results$AOI,
                      "AOI_Students" = "Students",
                      "AOI_Disruptive_Person" = "Disruptive Person",
                      "AOI_Teacher_Material" = "Teacher Material",
                      "AOI_Student_Desk" = "Student Desk",
                      "AOI_Classroom_Others" = "Classroom/Others")

# Display the t-Test and Effect Size table in APA style
knitr::kable(results, 
             caption = "t-Test and Effect Size for Fixation Duration Percentage across AOIs (Experts vs. Novices)", 
             row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Average duration of fixations in milliseconds (micro-teaching unit)
```{r dur_all, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### AVERAGE FIXATION DURATION ANALYSIS ########################

# 1. Data Preparation: Filter and Calculate Average Fixation Duration in the Micro-Teaching Unit
df_aoi_sum <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Group,
    Participant,
    Duration_of_interval,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations"),
    starts_with("Average_duration"),
    !"Total_duration_of_fixations.Disruptive_Person",
    !"Number_of_fixations.Disruptive_Person",
    !"Average_duration_of_fixations.Disruptive_Person"
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Participant = Participant,
    Duration_of_interval = Duration_of_interval,
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Sum_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Sum_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Number_fixation_min_mtu = round(Sum_number_fixation / Duration_of_interval_min, 2),
    Average_duration_fixation = round(Sum_duration_fixation / Sum_number_fixation, 2)
  ) %>%
  drop_na()

# 2. Descriptive Statistics Table
fixation_duration_table <- df_aoi_sum %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(Average_duration_fixation, na.rm = TRUE), 2),
    SD = round(sd(Average_duration_fixation, na.rm = TRUE), 2),
    Min = round(min(Average_duration_fixation, na.rm = TRUE), 2),
    Max = round(max(Average_duration_fixation, na.rm = TRUE), 2)
  )

# Display the descriptive table in APA style
knitr::kable(fixation_duration_table, caption = "N, M, SD, min & max Average Fixation Duration (micro-teaching unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Calculate means and standard deviations for Novice and Expert groups
mean_novice_fixation <- round(mean(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Novice"], na.rm = TRUE), 2)
sd_novice_fixation <- round(sd(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Novice"], na.rm = TRUE), 2)
mean_expert_fixation <- round(mean(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Expert"], na.rm = TRUE), 2)
sd_expert_fixation <- round(sd(df_aoi_sum$Average_duration_fixation[df_aoi_sum$Group == "Expert"], na.rm = TRUE), 2)

# 4. Perform t-test to get the p-value
t_test_fixation <- t.test(
  Average_duration_fixation ~ Group,
  data = df_aoi_sum,
  var.equal = TRUE
)
p_value_fixation <- round(t_test_fixation$p.value, 3)

# 5. Calculate Cohen's d for effect size using formula syntax
library(effsize)
cohen_d_fixation <- round(cohen.d(Average_duration_fixation ~ Group, data = df_aoi_sum)$estimate, 2)

# 6. Plotting the Average Fixation Duration
y_max_fixation <- max(df_aoi_sum$Average_duration_fixation) * 1.5

plot_avg_fixation_group <- df_aoi_sum %>%
  ggplot(mapping = aes(x = Group, y = Average_duration_fixation)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_fixation) +
  labs(x = "", y = "Average Fixation Duration (ms)") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Average Fixation Duration\n(micro-teaching unit)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  # Add annotations for mean (M) and SD
  annotate("text", x = 1, y = y_max_fixation * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice_fixation, ", italic(SD) == ", sd_novice_fixation, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max_fixation * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert_fixation, ", italic(SD) == ", sd_expert_fixation, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Add p-value and Cohen's d annotations
  annotate("text", x = 1.5, y = y_max_fixation * 0.95,
           label = paste0("p = ", p_value_fixation, "\nd = ", cohen_d_fixation),
           hjust = 0.5, size = 4, fontface = "italic")

plot_avg_fixation_group

# 7. Format the t-Test and Cohen's d results for APA Table without the first column
t_test_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_fixation$statistic, 2),
  "df" = t_test_fixation$parameter,
  "p-value" = round(t_test_fixation$p.value, 3),
  "Mean Difference" = round(t_test_fixation$estimate[1] - t_test_fixation$estimate[2], 2),
  "Cohen's d" = cohen_d_fixation,
  "95% CI (d)" = paste0("[", round(t_test_fixation$conf.int[1], 2), ", ", round(t_test_fixation$conf.int[2], 2), "]")
)

# Display t-Test and Effect Size results in APA style without row names
knitr::kable(t_test_result, caption = "t-Test and Effect Size for Average Fixation Duration (Micro-Teaching Unit)", row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Average duration of fixations (AOI students)
```{r dur_students, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### AVERAGE DURATION ON AOI STUDENTS ########################

# 1. Data Preparation: Filter and Calculate Average Duration of Fixations on AOI Students
df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select (
    Group,
    Duration_of_interval,
    Participant,
    "Total_duration_of_fixations.Anna",
    "Total_duration_of_fixations.Bianca",
    "Total_duration_of_fixations.Carl(a)",
    "Number_of_fixations.Anna",
    "Number_of_fixations.Bianca",
    "Number_of_fixations.Carl(a)"
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Participant = Participant,
    Duration_of_interval_min = round(Duration_of_interval/60000, digits = 2),
    Duration_of_interval_sec = round(Duration_of_interval/1000, digits = 2),
    Stud_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Stud_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Stud_number_fixation_min = round(Stud_number_fixation / Duration_of_interval_min, digits = 2),
    Stud_number_fixation_sec = round(Stud_number_fixation / Duration_of_interval_sec, digits = 2),
    Average_duration_stud = round(Stud_duration_fixation / Stud_number_fixation, digits = 0)
    ) -> df_aoi_stud

# 2. Descriptive Statistics Table
dur_stud_table <- df_aoi_stud %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M in ms" = round(mean(Average_duration_stud, na.rm = TRUE), 2),
    "SD in ms" = round(sd(Average_duration_stud, na.rm = TRUE), 2),
    "Min in ms" = round(min(Average_duration_stud, na.rm = TRUE), 2),
    "Max in ms" = round(max(Average_duration_stud, na.rm = TRUE), 2)
  )

# Display the descriptive table in APA style
knitr::kable(dur_stud_table,
             caption = "N, M, SD, Min & Max Average Duration of Fixations in Milliseconds (AOI Students)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# 3. Plotting Average Duration on AOI Students
# Calculate values for annotation
mean_novice_stud <- round(mean(df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Novice"], na.rm = TRUE), 2)
sd_novice_stud <- round(sd(df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Novice"], na.rm = TRUE), 2)
mean_expert_stud <- round(mean(df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Expert"], na.rm = TRUE), 2)
sd_expert_stud <- round(sd(df_aoi_stud$Average_duration_stud[df_aoi_stud$Group == "Expert"], na.rm = TRUE), 2)

# Maximum y-axis value for plot
y_max_stud <- max(df_aoi_stud$Average_duration_stud) * 1.2

# Conduct t-test
t_test_stud <- t.test(
  Average_duration_stud ~ Group,
  data = df_aoi_stud,
  var.equal = TRUE
)

# Extract p-value
p_value_stud <- round(t_test_stud$p.value, 3)

# Calculate Cohen's d (full result object)
d_aver_stud_full <- cohen.d(Average_duration_stud ~ Group, data = df_aoi_stud, pooled = TRUE)

# Extract Cohen's d and confidence interval
d_aver_stud <- round(d_aver_stud_full$estimate, 2)
d_conf_int <- paste0("[", round(d_aver_stud_full$conf.int[1], 2), ", ", round(d_aver_stud_full$conf.int[2], 2), "]")

# Annotated Plot
plot_average_duration_stud <- df_aoi_stud %>%
  ggplot(mapping = aes(x = Group, y = Average_duration_stud)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA, width = 0.5) +
  geom_jitter(
    width = 0.15, height = 0.1, alpha = 0.5, color = "black", size = 1.5
  ) +
  ylim(0, y_max_stud) + 
  labs(x = "", y = "Average Fixation Duration (ms)") + 
  scale_fill_brewer(palette = "RdBu") + 
  ggtitle("Average Duration of Fixations\n(AOI Students)") +
  theme_cowplot() +
  theme(legend.position = "none") +
  # Annotate mean (M) and SD
  annotate("text", x = 1, y = y_max_stud * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice_stud, ", italic(SD) == ", sd_novice_stud, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max_stud * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert_stud, ", italic(SD) == ", sd_expert_stud, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Annotate p-value and Cohen's d
  annotate("text", x = 1.5, y = y_max_stud * 0.95,
           label = paste0("p = ", p_value_stud, "\nd = ", d_aver_stud),
           hjust = 0.5, size = 4, fontface = "italic")

plot_average_duration_stud

# 4. t-Test and Effect Size Results Table
t_test_stud_result <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_stud$statistic, 2),
  "df" = t_test_stud$parameter,
  "p-value" = round(t_test_stud$p.value, 3),
  "Mean Difference" = round(t_test_stud$estimate[1] - t_test_stud$estimate[2], 2),
  "Cohen's d" = d_aver_stud,
  "95% CI (d)" = d_conf_int
)

# Display the t-Test and Effect Size results table in APA style
knitr::kable(t_test_stud_result,
             caption = "t-Test and Effect Size for Average Fixation Duration on AOI Students (milliseconds)",
             row.names = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Average duration of fixations (AOI disruptive person)
```{r average_dur_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### AVERAGE DURATION ON AOI DISRUPTIVE PERSON ########################

##################### DATA PREPARATION ########################

# Define the different types of disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Step 1: Filter and prepare the data
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>% # Filter rows based on the disruption list
  select(Group, Participant, TOI, 
         "Total_duration_of_fixations.Disruptive_Person", 
         "Number_of_fixations.Disruptive_Person") %>%
  mutate(
    Group = as_factor(Group), # Convert Group column to a factor
    Disruption_Category = case_when( # Assign categories based on disruption type
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Average_duration_disrup = round(Total_duration_of_fixations.Disruptive_Person / 
                                    Number_of_fixations.Disruptive_Person, 2) # Average duration in milliseconds
  ) %>%
  filter(!is.na(Disruption_Category) & !is.na(Average_duration_disrup)) # Remove NAs

##################### DESCRIPTIVE STATISTICS ########################

# Step 2: Summarize data by disruption category
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarise(
    mean_duration_ms = round(mean(Average_duration_disrup, na.rm = TRUE), 2),
    .groups = "drop"
  )

# Step 3: Summarize by participant
individual_summary <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarise(
    mean_duration = round(mean(mean_duration_ms, na.rm = TRUE), 2), # Mean duration per participant
    .groups = "drop"
  )

# Overall descriptive statistics
overall_stats <- individual_summary %>%
  group_by(Group) %>%
  summarise(
    N = n(), # Number of participants
    Mean = round(mean(mean_duration, na.rm = TRUE), 2), # Mean duration
    SD = round(sd(mean_duration, na.rm = TRUE), 2), # Standard deviation
    Min = round(min(mean_duration, na.rm = TRUE), 2), # Minimum value
    Max = round(max(mean_duration, na.rm = TRUE), 2), # Maximum value
    .groups = "drop"
  )

# Display the overall descriptive statistics table
knitr::kable(
  overall_stats,
  caption = "Descriptive Statistics (N, Mean, SD, Min, Max) for Average Duration by Group",
  col.names = c("Group", "N", "M", "SD", "Min", "Max"),
  row.names = FALSE
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Descriptive statistics by category
category_stats <- disruption_summary %>%
  group_by(Disruption_Category, Group) %>%
  summarise(
    N = n(), # Number of data points
    Mean = round(mean(mean_duration_ms, na.rm = TRUE), 2), # Mean duration
    SD = round(sd(mean_duration_ms, na.rm = TRUE), 2), # Standard deviation
    Min = round(min(mean_duration_ms, na.rm = TRUE), 2), # Minimum value
    Max = round(max(mean_duration_ms, na.rm = TRUE), 2), # Maximum value
    .groups = "drop"
  )

# Display the descriptive statistics table by category
knitr::kable(
  category_stats,
  caption = "Descriptive Statistics (N, Mean, SD, Min, Max) for Average Duration by Group and Category",
  col.names = c("Category", "Group", "N", "M", "SD", "Min", "Max"),
  row.names = FALSE
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### PLOTTING ########################

# Overall plot
mean_novice <- round(mean(individual_summary$mean_duration[individual_summary$Group == "Novice"], na.rm = TRUE), 2)
sd_novice <- round(sd(individual_summary$mean_duration[individual_summary$Group == "Novice"], na.rm = TRUE), 2)
mean_expert <- round(mean(individual_summary$mean_duration[individual_summary$Group == "Expert"], na.rm = TRUE), 2)
sd_expert <- round(sd(individual_summary$mean_duration[individual_summary$Group == "Expert"], na.rm = TRUE), 2)

# t-Test and effect size
t_test <- t.test(mean_duration ~ Group, data = individual_summary, var.equal = TRUE)
p_value <- round(t_test$p.value, 3)
cohen_d <- round(effsize::cohen.d(mean_duration ~ Group, data = individual_summary)$estimate, 2)

# Overall plot
y_max_overall <- max(individual_summary$mean_duration, na.rm = TRUE) * 1.5
plot_overall <- ggplot(individual_summary, aes(x = Group, y = mean_duration, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.7, color = "black", size = 1.5) +
  labs(x = "", y = "Mean Duration (ms) per Participant") +
  ggtitle("Overall Average Duration per Participant\n(AOI Disruptive Person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  ylim(0, y_max_overall) +
  theme(legend.position = "none") +
  annotate("text", x = 1, y = y_max_overall * 0.85,
           label = paste0("M = ", mean_novice, "\nSD = ", sd_novice),
           hjust = 0.5, size = 4, fontface = "italic") +
  annotate("text", x = 2, y = y_max_overall * 0.85,
           label = paste0("M = ", mean_expert, "\nSD = ", sd_expert),
           hjust = 0.5, size = 4, fontface = "italic") +
  annotate("text", x = 1.5, y = y_max_overall * 0.95,
           label = paste0("p = ", p_value, "\nd = ", cohen_d),
           hjust = 0.5, size = 4, fontface = "italic")

# Plot by category
plot_category <- disruption_summary %>%
  ggplot(aes(x = Group, y = mean_duration_ms, fill = Group)) +
  geom_boxplot(outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.1, alpha = 0.7, size = 1.5, color = "black") +
  facet_wrap(~Disruption_Category, scales = "fixed", labeller = label_wrap_gen(width = 20)) +
  ylim(0, 2000) +
  labs(x = "", y = "Mean Duration (ms)") +
  ggtitle("Average Duration of Fixations by Disruption Category and Group") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10)
  )

# Display plots
plot_overall
plot_category

##################### ANOVA AND POST-HOC TESTS ########################

# Perform the repeated-measures ANOVA
anova_afex <- afex::aov_ez(
  id = "Participant", # Random effect
  dv = "mean_duration_ms", # Dependent variable
  between = "Group", # Between-subject factor
  within = "Disruption_Category", # Within-subject factor
  data = disruption_summary
)

# Extract the ANOVA table
anova_results <- as.data.frame(anova_afex$anova_table)

# Recalculate generalized eta squared (η²)
eta_squared_results <- effectsize::eta_squared(anova_afex, partial = FALSE)

# Combine recalculated eta squared with the ANOVA table
anova_table_apa <- anova_results %>%
  rownames_to_column(var = "Effect") %>% # Move row names to a column
  left_join(
    as.data.frame(eta_squared_results), 
    by = c("Effect" = "Parameter") # Match the correct columns
  ) %>%
  mutate(
    `F` = round(`F`, 2), # Round F-values
    `η²` = round(Eta2, 3), # Round generalized eta squared
    `p-value` = ifelse(
      `Pr(>F)` < 0.05, "< .05", round(`Pr(>F)`, 3) # Format p-values
    )
  ) %>%
  select(Effect, `num Df`, `den Df`, `F`, `p-value`, `η²`) # Order columns, place η² last

# Display the ANOVA table
kable(
  anova_table_apa,
  caption = "Repeated-Measures ANOVA Results for Average Duration (Including Effect Sizes)",
  col.names = c("Effect", "Df1", "Df2", "F", "p", "η²"),
  digits = 2
) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### POST-HOC TESTS ########################

# Perform post-hoc pairwise comparisons for Group
post_hoc_group <- emmeans::emmeans(anova_afex, pairwise ~ Group)

# Extract post-hoc pairwise contrasts for Group
post_hoc_results_group <- as.data.frame(post_hoc_group$contrasts) %>%
  mutate(
    Comparison = gsub("_", " ", contrast), # Replace underscores with spaces
    Estimate = round(estimate, 2),
    SE = round(SE, 2),
    t = round(t.ratio, 2),
    p.value = ifelse(p.value < 0.05, "< .05", round(p.value, 3)),
    Cohen_d = round(estimate / sqrt(SE^2 * 2), 2) # Approximate Cohen's d
  ) %>%
  select(Comparison, Estimate, SE, df, t, p.value, Cohen_d) # Keep relevant columns

# Perform post-hoc pairwise comparisons for Disruption_Category
post_hoc_disruption <- emmeans::emmeans(anova_afex, pairwise ~ Disruption_Category)

# Extract post-hoc pairwise contrasts for Disruption_Category
post_hoc_results_disruption <- as.data.frame(post_hoc_disruption$contrasts) %>%
  mutate(
    Comparison = gsub("_", " ", contrast),  # Replace underscores with spaces
    Comparison = gsub("\\.", " ", Comparison), # Replace periods with spaces
    Estimate = round(estimate, 2),
    SE = round(SE, 2),
    t = round(t.ratio, 2),
    p.value = ifelse(p.value < 0.05, "< .05", round(p.value, 3)),
    Cohen_d = round(estimate / sqrt(SE^2 * 2), 2) # Approximate Cohen's d
  ) %>%
  select(Comparison, Estimate, SE, df, t, p.value, Cohen_d) # Keep relevant columns

##################### TABLE FOR GROUP ########################

# Display the post-hoc test results for Group
knitr::kable(
  post_hoc_results_group,
  caption = "Post-Hoc Pairwise Comparisons for Group (Including Effect Sizes)",
  col.names = c("Comparison", "Estimate", "SE", "df", "t", "p", "Cohen's d"),
  digits = 2
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

##################### TABLE FOR DISRUPTION CATEGORY ########################

# Display the post-hoc test results for Disruption_Category
knitr::kable(
  post_hoc_results_disruption,
  caption = "Post-Hoc Pairwise Comparisons for Disruption Categories (Including Effect Sizes)",
  col.names = c("Comparison", "Estimate", "SE", "df", "t", "p", "Cohen's d"),
  digits = 2
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Gaze Relational Index (GRI; micro-teaching unit)
```{r gri_all, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### GRI in MICRO-TEACHING UNIT ########################

# Step 1: Calculate GRI (Gaze Ratio Index) for each participant
df_gri <- df_aoi_sum %>%
  mutate(GRI = round(Average_duration_fixation / Number_fixation_min_mtu, 2))

# Step 2: Summarize GRI statistics (N, Mean, SD, Min, Max) by Group
gri_table <- df_gri %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    "M" = round(mean(GRI, na.rm = TRUE), 2),
    "SD" = round(sd(GRI, na.rm = TRUE), 2),
    "Min" = round(min(GRI, na.rm = TRUE), 2),
    "Max" = round(max(GRI, na.rm = TRUE), 2)
  )

# Display the GRI summary table in APA format
knitr::kable(gri_table,
             caption = "N, M, SD, Min, and Max GRI (Micro-Teaching Unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Step 3: Plot GRI by Group with boxplot and jittered points
# Calculate annotation values
mean_novice_gri <- round(mean(df_gri$GRI[df_gri$Group == "Novice"], na.rm = TRUE), 2)
sd_novice_gri <- round(sd(df_gri$GRI[df_gri$Group == "Novice"], na.rm = TRUE), 2)
mean_expert_gri <- round(mean(df_gri$GRI[df_gri$Group == "Expert"], na.rm = TRUE), 2)
sd_expert_gri <- round(sd(df_gri$GRI[df_gri$Group == "Expert"], na.rm = TRUE), 2)

# Maximum y-axis value for plot
y_max_gri <- max(df_gri$GRI, na.rm = TRUE) * 1.2

# Conduct t-test
t_test_gri <- t.test(
  GRI ~ Group,
  data = df_gri,
  var.equal = TRUE
)

# Extract p-value
p_value_gri <- round(t_test_gri$p.value, 3)

# Calculate Cohen's d (full result object)
d_gri_full <- cohen.d(GRI ~ Group, data = df_gri, pooled = TRUE)

# Extract Cohen's d and confidence interval
d_gri <- round(d_gri_full$estimate, 2)
d_conf_int_gri <- paste0("[", round(d_gri_full$conf.int[1], 2), ", ", round(d_gri_full$conf.int[2], 2), "]")

# Annotated Plot
plot_gri_group <- df_gri %>%
  ggplot(aes(x = Group, y = GRI, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 2, alpha = 0.7, width = 0.1, height = 0.1, seed = 1) +
  ylim(0, y_max_gri) +
  labs(x = "", y = "Average Duration of Fixations /\nNumber of Fixations per Minute") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("GRI\n(Micro-Teaching Unit)") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold")
  ) +
  # Annotate mean (M) and SD
  annotate("text", x = 1, y = y_max_gri * 0.85,
           label = paste0("atop(italic(M) == ", mean_novice_gri, ", italic(SD) == ", sd_novice_gri, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  annotate("text", x = 2, y = y_max_gri * 0.85,
           label = paste0("atop(italic(M) == ", mean_expert_gri, ", italic(SD) == ", sd_expert_gri, ")"),
           hjust = 0.5, size = 3.5, vjust = -0.5, parse = TRUE) +
  # Annotate p-value and Cohen's d
  annotate("text", x = 1.5, y = y_max_gri * 0.95,
           label = paste0("p = ", p_value_gri, "\nd = ", d_gri),
           hjust = 0.5, size = 4, fontface = "italic")

plot_gri_group

##################### t-Test & Effect Size for GRI ########################

# Step 6: Create a table for the t-test and Cohen's d results
gri_t_test_table <- data.frame(
  "Group Comparison" = "Experts vs. Novices",
  "t-value" = round(t_test_gri$statistic, 2),
  "df" = t_test_gri$parameter,
  "p-value" = round(t_test_gri$p.value, 3),
  "Mean Difference" = round(t_test_gri$estimate[1] - t_test_gri$estimate[2], 2),
  "Cohen's d" = d_gri,
  "95% CI (d)" = d_conf_int_gri
)

# Display the t-Test and Effect Size table for GRI in APA format
knitr::kable(gri_t_test_table, 
             caption = "t-Test and Effect Size for GRI (Micro-Teaching Unit)") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### Time to first fixation in seconds (AOI disruptive person)
```{r ttff_disrup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

##################### NUMBER OF FIXATION ON AOI DISRUPTIVE PERSON ########################

# Define disruptions and categories
disruptions <- c("Chatting_with_neighbour", "Clicking_pen", "Drawing", 
                 "Drumming_with_hands", "Head_on_table", "Heckling", 
                 "Looking_at_phone", "Snipping_with_fingers", "Whispering")
verbal_disruptions <- c("Chatting_with_neighbour", "Heckling", "Whispering")
physical_disruptions <- c("Clicking_pen", "Drumming_with_hands", "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone", "Head_on_table", "Drawing")

# Filter and prepare data
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%
  dplyr::select(Group, Participant, TOI, Duration_of_interval, 
                "Number_of_fixations.Disruptive_Person") %>%
  mutate(
    Group = as_factor(Group),
    Disruption_Category = case_when(
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = Duration_of_interval / 1000
  ) %>%
  filter(!is.na(Disruption_Category))

# Summarize data
disruption_summary <- df_aoi_disrup %>%
  group_by(Group, Participant, Disruption_Category) %>%
  summarize(
    total_duration_sec = sum(Duration_of_interval_sec, na.rm = TRUE),
    fixation_count = sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(fixations_per_second = fixation_count / total_duration_sec)

# Individual-level aggregation
individual_summary <- disruption_summary %>%
  group_by(Group, Participant) %>%
  summarize(
    mean_fixations_per_second = mean(fixations_per_second, na.rm = TRUE),
    .groups = "drop"
  )

# 1. APA-Conform Table: Overall Fixation Rate
overall_stats <- individual_summary %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    M = round(mean(mean_fixations_per_second, na.rm = TRUE), 2),
    SD = round(sd(mean_fixations_per_second, na.rm = TRUE), 2),
    Min = round(min(mean_fixations_per_second, na.rm = TRUE), 2),
    Max = round(max(mean_fixations_per_second, na.rm = TRUE), 2)
  )

# Render table for overall fixation rate
overall_table <- knitr::kable(
  overall_stats,
  caption = "Descriptive Statistics for Overall Fixation Rate (AOI Disruptive Person)",
  col.names = c("Group", "N", "M", "SD", "Min", "Max")
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Display overall table
overall_table

# 2. APA-Conform Table: Fixation Rate by Category Clustered by Disruption Category
category_stats <- disruption_summary %>%
  group_by(Disruption_Category, Group) %>%  # Group by category first, then by group
  summarise(
    N = n(),
    M = round(mean(fixations_per_second, na.rm = TRUE), 2),
    SD = round(sd(fixations_per_second, na.rm = TRUE), 2),
    Min = round(min(fixations_per_second, na.rm = TRUE), 2),
    Max = round(max(fixations_per_second, na.rm = TRUE), 2),
    .groups = "drop"  # Ungroup after summarization
  )

# Render the table clustered by Disruption Categories
category_table <- knitr::kable(
  category_stats,
  caption = "Descriptive Statistics for Fixation Rate by Disruption Category (AOI Disruptive Person)",
  col.names = c("Disruption Category", "Group", "N", "M", "SD", "Min", "Max")
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  kableExtra::group_rows(
    index = table(category_stats$Disruption_Category)  # Automatically cluster by category
  )

# Display the clustered category-specific table
category_table

# 3. Plot: Overall Fixation Rate
plot_overall <- individual_summary %>%
  ggplot(mapping = aes(x = Group, y = mean_fixations_per_second)) +
  geom_boxplot(mapping = aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  ylim (0, 1) + 
  labs(x = "", y = "Mean Fixations per Second per Participant") +
  ggtitle("Overall Fixation Rate per Participant\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(legend.position = "none")

# 4. Plot: Fixation Rate by Disruption Category
plot_category <- disruption_summary %>%
  ggplot(mapping = aes(x = Group, y = fixations_per_second, fill = Group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    size = 2,
    alpha = 0.7,
    position = position_jitter(
      seed = 1,
      width = 0.1,
      height = 0.1
    )
  ) +
  facet_wrap(~Disruption_Category, scales = "free", labeller = label_wrap_gen(width = 15)) + # Wrap long facet labels
  labs(x = "", y = "Fixations per Second") +
  ggtitle("Fixation Rate by Disruption Category\n(AOI disruptive person)") +
  scale_fill_brewer(palette = "RdBu") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10) # Adjust facet text size if necessary
  )

# Display plots
plot_overall
plot_category

# Step 5: Statistical Tests by Category
category_results <- list()

for (category in unique(disruption_summary$Disruption_Category)) {
  cat_data <- disruption_summary %>% filter(Disruption_Category == category)
  t_cat <- t.test(fixations_per_second ~ Group, data = cat_data, var.equal = TRUE)
  d_cat <- cohen.d(fixations_per_second ~ Group, data = cat_data)
  category_results[[category]] <- data.frame(
    "Group Comparison" = "Experts vs. Novices",
    "t-value" = round(t_cat$statistic, 2),
    "df" = t_cat$parameter,
    "p-value" = round(t_cat$p.value, 3),
    "Mean Difference" = round(t_cat$estimate[1] - t_cat$estimate[2], 2),
    "Cohen's d" = round(d_cat$estimate, 2),
    "95% CI (d)" = paste0("[", round(d_cat$conf.int[1], 2), ", ", round(d_cat$conf.int[2], 2), "]")
  )
}

# Combine results into a single data frame and include Disruption Category
category_stats <- do.call(rbind, lapply(names(category_results), function(cat) {
  result <- category_results[[cat]]
  result$Disruption_Category <- cat  # Add Disruption Category as a column
  return(result)
}))

# Remove redundant Disruption Category column from the table
category_stats <- category_stats %>%
  relocate(Disruption_Category) %>%
  arrange(Disruption_Category) %>%  # Ensure proper order by category
  select(-Disruption_Category)  # Drop the column after adding grouping below

# Render the table with grouped rows
knitr::kable(
  category_stats,
  caption = "t-Test and Effect Size for Fixation Rate by Disruption Category",
  col.names = c("Group Comparison", "t-value", "df", "p-value", "Mean Difference", "Cohen's d", "95% CI (d)")
) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  kableExtra::group_rows(
    group_label = "Lack of eagerness to learn",
    start_row = 1,
    end_row = 1
  ) %>%
  kableExtra::group_rows(
    group_label = "Physical disruptions",
    start_row = 2,
    end_row = 2
  ) %>%
  kableExtra::group_rows(
    group_label = "Verbal disruptions",
    start_row = 3,
    end_row = 3
  )

```

### Rating Scales (Disruption Appraisal, Confidence Appraisal, Prevalence Rating)
```{r rating_scales, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Data Import and Preprocessing ------------------------------------------
sri <- excel_sheets("data/Coding_SRI.xlsx") %>%
  map_df(~ read_xlsx("data/Coding_SRI.xlsx", .)) %>%
  filter(
    !ID %in% c(201, 223),
    !disruption_appraisal %in% c(-100, -99, -88),
    !confidence_appraisal %in% c(-100, -99, -88),
    !prevalence_rating %in% c(-100, -99, -88)
  ) %>%
  mutate(ID = ifelse(ID < 200, "Novice", "Expert"))

# Descriptive Statistics -------------------------------------------------
## Disruption Appraisal
sri_disrup_table <- sri %>%
  group_by(ID) %>%
  summarise(
    N = n(),
    M = round(mean(disruption_appraisal, na.rm = TRUE), 2),
    SD = round(sd(disruption_appraisal, na.rm = TRUE), 2),
    Min = round(min(disruption_appraisal, na.rm = TRUE), 2),
    Max = round(max(disruption_appraisal, na.rm = TRUE), 2)
  )

## Confidence Appraisal
sri_confi_table <- sri %>%
  group_by(ID) %>%
  summarise(
    N = n(),
    M = round(mean(confidence_appraisal, na.rm = TRUE), 2),
    SD = round(sd(confidence_appraisal, na.rm = TRUE), 2),
    Min = round(min(confidence_appraisal, na.rm = TRUE), 2),
    Max = round(max(confidence_appraisal, na.rm = TRUE), 2)
  )

## Prevalence Rating
sri_preva_table <- sri %>%
  group_by(ID) %>%
  summarise(
    N = n(),
    M = round(mean(prevalence_rating, na.rm = TRUE), 2),
    SD = round(sd(prevalence_rating, na.rm = TRUE), 2),
    Min = round(min(prevalence_rating, na.rm = TRUE), 2),
    Max = round(max(prevalence_rating, na.rm = TRUE), 2)
  )

# APA-Conform Tables -----------------------------------------------------
kable(sri_disrup_table, caption = "Descriptive Statistics for Disruption Appraisal", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(sri_confi_table, caption = "Descriptive Statistics for Confidence Appraisal", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(sri_preva_table, caption = "Descriptive Statistics for Prevalence Rating", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# t-Tests and Effect Sizes -----------------------------------------------
## Disruption Appraisal
disrup_ttest <- t.test(
  disruption_appraisal ~ ID, data = sri, var.equal = TRUE
)
disrup_d <- cohen.d(disruption_appraisal ~ ID, data = sri)

## Confidence Appraisal
confi_ttest <- t.test(
  confidence_appraisal ~ ID, data = sri, var.equal = TRUE
)
confi_d <- cohen.d(confidence_appraisal ~ ID, data = sri)

## Prevalence Rating
preva_ttest <- t.test(
  prevalence_rating ~ ID, data = sri, var.equal = TRUE
)
preva_d <- cohen.d(prevalence_rating ~ ID, data = sri)

# APA-Conform Tables for t-Tests and Effect Sizes ------------------------
disrup_stats <- tibble(
  Statistic = c("t-value", "df", "p-value", "Cohen's d"),
  Value = c(
    round(disrup_ttest$statistic, 2),
    round(disrup_ttest$parameter, 2),
    round(disrup_ttest$p.value, 3),
    round(disrup_d$estimate, 2)
  )
)

confi_stats <- tibble(
  Statistic = c("t-value", "df", "p-value", "Cohen's d"),
  Value = c(
    round(confi_ttest$statistic, 2),
    round(confi_ttest$parameter, 2),
    round(confi_ttest$p.value, 3),
    round(confi_d$estimate, 2)
  )
)

preva_stats <- tibble(
  Statistic = c("t-value", "df", "p-value", "Cohen's d"),
  Value = c(
    round(preva_ttest$statistic, 2),
    round(preva_ttest$parameter, 2),
    round(preva_ttest$p.value, 3),
    round(preva_d$estimate, 2)
  )
)

kable(disrup_stats, caption = "t-Test and Effect Size for Disruption Appraisal", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(confi_stats, caption = "t-Test and Effect Size for Confidence Appraisal", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(preva_stats, caption = "t-Test and Effect Size for Prevalence Rating", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Plotting ---------------------------------------------------------------
## Disruption Appraisal
disrup_plot <- sri %>%
  ggplot(aes(x = ID, y = disruption_appraisal, fill = ID)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 1.5, alpha = 0.6, width = 0.2, height = 0.1) +
  labs(x = "", y = "Disruption Appraisal") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Disruption Appraisal by Group") +
  theme_cowplot() +
  theme(legend.position = "none") +
  annotate(
    "text", x = 1.5, y = max(sri$disruption_appraisal, na.rm = TRUE) * 1.2,
    label = paste0(
      "p = ", round(disrup_ttest$p.value, 3), ", d = ",
      round(disrup_d$estimate, 2)
    ),
    size = 4, hjust = 0.5, fontface = "italic"
  )

print(disrup_plot)

## Confidence Appraisal
confi_plot <- sri %>%
  ggplot(aes(x = ID, y = confidence_appraisal, fill = ID)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 1.5, alpha = 0.6, width = 0.2, height = 0.1) +
  labs(x = "", y = "Confidence Appraisal") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Confidence Appraisal by Group") +
  theme_cowplot() +
  theme(legend.position = "none") +
  annotate(
    "text", x = 1.5, y = max(sri$confidence_appraisal, na.rm = TRUE) * 1.2,
    label = paste0(
      "p = ", round(confi_ttest$p.value, 3), ", d = ",
      round(confi_d$estimate, 2)
    ),
    size = 4, hjust = 0.5, fontface = "italic"
  )

print(confi_plot)

## Prevalence Rating
preva_plot <- sri %>%
  ggplot(aes(x = ID, y = prevalence_rating, fill = ID)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(size = 1.5, alpha = 0.6, width = 0.2, height = 0.1) +
  labs(x = "", y = "Prevalence Rating") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("Prevalence Rating by Group") +
  theme_cowplot() +
  theme(legend.position = "none") +
  annotate(
    "text", x = 1.5, y = max(sri$prevalence_rating, na.rm = TRUE) * 1.2,
    label = paste0(
      "p = ", round(preva_ttest$p.value, 3), ", d = ",
      round(preva_d$estimate, 2)
    ),
    size = 4, hjust = 0.5, fontface = "italic"
  )

print(preva_plot)

# Omega Calculations -----------------------------------------------------
# Data Import
sri <- excel_sheets("data/Coding_SRI.xlsx") %>%
  map_df(~ read_xlsx("data/Coding_SRI.xlsx", .)) %>%
  filter(
    !ID %in% c(201, 223), # Exclude specific IDs
    !disruption_appraisal %in% c(-100, -99, -88),
    !confidence_appraisal %in% c(-100, -99, -88),
    !prevalence_rating %in% c(-100, -99, -88)
  )

# Change Long Format to Wide Format
sri_wide <- sri %>%
  pivot_wider(
    names_from = event,
    values_from = c(disruption_appraisal, confidence_appraisal, prevalence_rating)
  )

# Helper Function for Omega Calculation
calculate_omega_total <- function(data, scale_name) {
  if (nrow(data) == 0 || ncol(data) == 0) {
    warning(paste("No valid data for", scale_name))
    return(NA)
  }

  # Try Omega Calculation
  result <- tryCatch(
    omega(data),
    error = function(e) {
      warning(paste("Omega calculation failed for", scale_name, ":", e$message))
      return(NULL)
    }
  )

  # Check if Result is Valid and Extract omega.tot
  if (is.null(result) || !"omega.tot" %in% names(result)) {
    warning(paste("Omega total not available for", scale_name))
    return(NA)
  }

  return(round(result$omega.tot, 2))
}

# Disruption Appraisal ---------------------------------------------------
sri_wide_disrup <- sri_wide %>%
  dplyr::select(starts_with("disruption_appraisal")) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  drop_na()

omega_total_disrup <- calculate_omega_total(sri_wide_disrup, "Disruption Appraisal")

# Confidence Appraisal ---------------------------------------------------
sri_wide_confi <- sri_wide %>%
  dplyr::select(starts_with("confidence_appraisal")) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  drop_na()

omega_total_confi <- calculate_omega_total(sri_wide_confi, "Confidence Appraisal")

# Prevalence Rating ------------------------------------------------------
sri_wide_preva <- sri_wide %>%
  dplyr::select(starts_with("prevalence_rating")) %>%
  mutate(across(everything(), ~ as.numeric(.))) %>%
  drop_na()

omega_total_preva <- calculate_omega_total(sri_wide_preva, "Prevalence Rating")

# APA-Conform Table ------------------------------------------------------
omega_table <- tibble(
  Measure = c("Disruption Appraisal", "Confidence Appraisal", "Prevalence Rating"),
  Omega_Total = c(omega_total_disrup, omega_total_confi, omega_total_preva)
)

# Display APA-Conform Table ----------------------------------------------
kable(omega_table, caption = "Internal Consistency (Omega Total) for Appraisal Measures", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```

### Situational Jugdement Test
```{r sjt, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# Data Import and Preprocessing ------------------------------------------
df_sjt <- read_excel("./data/SJT.xlsx") %>%
  dplyr::select(
    UI06_05,  # ID
    SJT_AL_gek, # Monitoring
    SJT_ST_gek, # Managing momentum
    SJT_R_gek,  # Rules and routines
    SJT_KF_gek  # All
  ) %>%
  filter(!UI06_05 == 201,
         !UI06_05 == 223) %>% # Exclude specific IDs
  na.omit() %>% # Remove NAs
  rename(
    Group = UI06_05,
    Monitoring = SJT_AL_gek,
    `Managing momentum` = SJT_ST_gek,
    `Rules and routines` = SJT_R_gek,
    All = SJT_KF_gek
  ) %>%
  mutate(Group = ifelse(Group < 200, "Novice", "Expert")) # Define groups

# Reshape Data to Long Format --------------------------------------------
df_sjt_long <- df_sjt %>%
  pivot_longer(
    !Group,
    names_to = "Facets Classroom Management",
    values_to = "Mean"
  )

# Descriptive Statistics Tables ------------------------------------------
## Function to Calculate Descriptive Stats
descriptive_stats <- function(data, facet) {
  data %>%
    filter(`Facets Classroom Management` == facet) %>%
    group_by(Group) %>%
    summarise(
      N = n(),
      M = round(mean(Mean), 2),
      SD = round(sd(Mean), 2),
      Min = round(min(Mean), 2),
      Max = round(max(Mean), 2),
      .groups = "drop"
    )
}

## Generate Tables
sjt_all <- descriptive_stats(df_sjt_long, "All")
sjt_mm <- descriptive_stats(df_sjt_long, "Managing momentum")
sjt_m <- descriptive_stats(df_sjt_long, "Monitoring")
sjt_r <- descriptive_stats(df_sjt_long, "Rules and routines")

## Display APA-Conform Tables
kable(sjt_all, caption = "Descriptive Statistics for Overall SJT Score") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(sjt_mm, caption = "Descriptive Statistics for Managing Momentum") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(sjt_m, caption = "Descriptive Statistics for Monitoring") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

kable(sjt_r, caption = "Descriptive Statistics for Rules and Routines") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Plotting SJT Facets ----------------------------------------------------
mean_plot <- df_sjt_long %>%
  mutate(Group = factor(Group, levels = c("Novice", "Expert"))) %>%
  ggplot(aes(x = Group, y = Mean)) +
  geom_boxplot(aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 1,
    alpha = 0.4,
    position = position_jitter(seed = 1, width = 0.1, height = 0.1)
  ) +
  labs(x = "", y = "Mean SJT Score") +
  scale_fill_brewer(palette = "RdBu") +
  facet_wrap(~`Facets Classroom Management`, nrow = 1, strip.position = "bottom") +
  theme_cowplot() +
  ggtitle("SJT Scores by Group") +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    axis.title.y = element_text(size = 14),
    axis.text.x = element_blank(), # Remove x-axis text
    axis.ticks.x = element_blank(), # Remove x-axis ticks
    strip.text = element_text(size = 10) # Increase facet label font size
  )

print(mean_plot)

# Statistical Tests ------------------------------------------------------
## Function to Perform t-Test and Effect Size
t_test_stats <- function(data, group_col, value_col, facet_name) {
  group_col <- ensym(group_col)  # Ensure proper quoting of group column
  value_col <- ensym(value_col)  # Ensure proper quoting of value column

  # Perform t-Test
  t_test <- t.test(
    formula = as.formula(paste0("`", quo_name(value_col), "` ~ ", quo_name(group_col))),
    data = data,
    var.equal = TRUE
  )

  # Calculate Cohen's d
  cohen_d <- CohenD(
    x = data %>% filter(!!group_col == "Expert") %>% pull(!!value_col),
    y = data %>% filter(!!group_col == "Novice") %>% pull(!!value_col)
  )

  tibble(
    `Facet` = facet_name,
    `t-value` = round(t_test$statistic, 2),
    `p-value` = round(t_test$p.value, 3),
    `Cohen's d` = round(cohen_d, 2)
  )
}

## Generate t-Test Tables
t_test_all <- t_test_stats(df_sjt, "Group", "All", "Overall")
t_test_mm <- t_test_stats(df_sjt, "Group", "Managing momentum", "Managing Momentum")
t_test_m <- t_test_stats(df_sjt, "Group", "Monitoring", "Monitoring")
t_test_r <- t_test_stats(df_sjt, "Group", "Rules and routines", "Rules and Routines")

## Combine Results into a Table
t_test_results <- bind_rows(t_test_all, t_test_mm, t_test_m, t_test_r)

## Display APA-Conform Table
kable(t_test_results, caption = "t-Test and Effect Sizes for SJT Facets") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Internal Consistency (Omega) -------------------------------------------
# Read in Data
df_sjt <- read_excel("./data/SJT.xlsx")  # Replace with the correct path

# Helper Function for Omega Calculation
calculate_omega_total <- function(data, scale_name) {
  if (nrow(data) == 0 || ncol(data) == 0) {
    warning(paste("No valid data for", scale_name))
    return(NA)
  }

  # Try Omega Calculation
  result <- tryCatch(
    omega(data),
    error = function(e) {
      warning(paste("Omega calculation failed for", scale_name, ":", e$message))
      return(NULL)
    }
  )

  # Check if Result is Valid and Extract omega.tot
  if (is.null(result) || !"omega.tot" %in% names(result)) {
    warning(paste("Omega total not available for", scale_name))
    return(NA)
  }

  return(round(result$omega.tot, 2))
}

### MANAGING MOMENTUM ----------------------------------------------------
# Select relevant columns and remove NAs
df_sjt_mm <- df_sjt %>%
  dplyr::select(WT_S1_F1F2_AL:WT_S4_F4F6_AL) %>%
  na.omit()

# Calculate omega
omega_mm <- calculate_omega_total(df_sjt_mm, "Managing Momentum")

### MONITORING -----------------------------------------------------------
# Select relevant columns and remove NAs
df_sjt_m <- df_sjt %>%
  dplyr::select(WT_S5_F1F2_ST:WT_S9_F4F5_ST) %>%
  na.omit()

# Calculate omega
omega_m <- calculate_omega_total(df_sjt_m, "Monitoring")

### RULES & ROUTINES -----------------------------------------------------
# Select relevant columns and remove NAs
df_sjt_r <- df_sjt %>%
  dplyr::select(WT_S11_F1F5_R:WT_S14_F4F6_R) %>%
  na.omit()

# Calculate omega
omega_r <- calculate_omega_total(df_sjt_r, "Rules and Routines")

### OVERALL TEST -----------------------------------------------------
# Select relevant columns and remove NAs
df_overall <- df_sjt %>%
  select(matches("WT_S")) %>%
  mutate(across(everything(), as.numeric)) %>%  # Ensure numeric data
  drop_na()

# Calculate omega
omega_overall <- calculate_omega_total(df_overall, "Overall Test")

### COMBINE RESULTS INTO AN APA-CONFORM TABLE ----------------------------
# APA-Conform Table ------------------------------------------------------
omega_results <- tibble(
  Facet = c("Overall Test", "Managing Momentum", "Monitoring", "Rules and Routines"),
  `Omega Total` = c(omega_overall, omega_mm, omega_m, omega_r)
)

# Display APA-Conform Table ----------------------------------------------
kable(omega_results, caption = "Internal Consistency (Omega) for SJT Facets") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```

### Self-evaluation 
```{r self-eval, echo=FALSE, results='asis'}

# Data Import and Preparation --------------------------------------------
df_quest <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  dplyr::select(
    LI06_05, 
    LM01_01:LM01_08,  # Classroom Management
    LP01_01:LP01_08   # Non-/Paraverbal Communication
  ) %>%
  filter(LI06_05 != 201, LI06_05 != 223) %>%  # Exclude specific IDs
  mutate(Group = ifelse(LI06_05 < 200, "Novice", "Expert")) %>%  # Group Definition
  select(-LI06_05) %>%  # Drop ID column
  na.omit()  # Remove NAs

# Summary Statistics Table -----------------------------------------------
df_quest_table <- df_quest %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    `M cm` = round(mean(c_across(starts_with("LM")), na.rm = TRUE), 2),
    `SD cm` = round(sd(c_across(starts_with("LM")), na.rm = TRUE), 2),
    `Min cm` = round(min(c_across(starts_with("LM")), na.rm = TRUE), 2),
    `Max cm` = round(max(c_across(starts_with("LM")), na.rm = TRUE), 2),
    `M n&pv com` = round(mean(c_across(starts_with("LP")), na.rm = TRUE), 2),
    `SD n&pv com` = round(sd(c_across(starts_with("LP")), na.rm = TRUE), 2),
    `Min n&pv com` = round(min(c_across(starts_with("LP")), na.rm = TRUE), 2),
    `Max n&pv com` = round(max(c_across(starts_with("LP")), na.rm = TRUE), 2)
  )

# Display APA-Conform Table
kable(df_quest_table, 
      caption = "Summary Statistics for Classroom Management and Non-/Paraverbal Communication") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Create Dataset for Plotting --------------------------------------------
df_quest_long <- df_quest %>%
  rowwise() %>%
  transmute(
    Group = Group,
    All = round(mean(c_across(starts_with("L")), na.rm = TRUE), 2),
    `Classroom Management` = round(mean(c_across(starts_with("LM")), na.rm = TRUE), 2),
    `Non-/Paraverbal Communication` = round(mean(c_across(starts_with("LP")), na.rm = TRUE), 2)
  ) %>%
  ungroup() %>%
  pivot_longer(!Group, names_to = "Scales Self-evaluation", values_to = "Mean")

# Plot Mean Scores by Group ----------------------------------------------
quest_plot <- df_quest_long %>%
  mutate(Group = factor(Group, levels = c("Novice", "Expert"))) %>%
  ggplot(aes(x = Group, y = Mean)) +
  geom_boxplot(aes(fill = Group), outlier.shape = NA) +
  geom_point(
    size = 1,
    alpha = 0.4,
    position = position_jitter(seed = 1, width = 0.1, height = 0.1)
  ) +
  labs(x = "", y = "Mean Scores") +
  scale_fill_brewer(palette = "RdBu") +
  facet_wrap(~`Scales Self-evaluation`, nrow = 1, strip.position = "bottom") +
  theme_cowplot() +
  ggtitle("Self-Evaluation Scores") +
  theme(
    plot.title = element_text(size = 25, face = "bold"),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 15),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text.x = element_text(size = 8),
    axis.title.y = element_text(size = 20)
  )

print(quest_plot)

# T-Test and Effect Sizes ------------------------------------------------
# t-Test Function for Long Format ----------------------------------------
t_test_stats_long <- function(data, group_col, scale_filter, scale_name) {
  # Subset Data for the Specific Scale
  filtered_data <- data %>%
    filter(`Scales Self-evaluation` == scale_filter)

  # Perform t-Test
  t_test <- t.test(
    Mean ~ get(group_col),
    data = filtered_data,
    var.equal = TRUE
  )

  # Calculate Cohen's d
  cohen_d <- CohenD(
    x = filtered_data %>% filter(!!rlang::sym(group_col) == "Expert") %>% pull(Mean),
    y = filtered_data %>% filter(!!rlang::sym(group_col) == "Novice") %>% pull(Mean)
  )

  # Return Results
  tibble(
    `Scale` = scale_name,
    `t-value` = round(t_test$statistic, 2),
    `p-value` = round(t_test$p.value, 3),
    `Cohen's d` = round(cohen_d, 2)
  )
}

# Perform t-Tests for Each Scale
t_test_all <- t_test_stats_long(df_quest_long, "Group", "All", "Overall")
t_test_cm <- t_test_stats_long(df_quest_long, "Group", "Classroom Management", "Classroom Management")
t_test_npvc <- t_test_stats_long(df_quest_long, "Group", "Non-/Paraverbal Communication", "Non-/Paraverbal Communication")

# Combine Results into a Single Table
t_test_results <- bind_rows(t_test_all, t_test_cm, t_test_npvc)

# Display Results in APA-Conform Table
kable(t_test_results, caption = "t-Test and Effect Sizes for Self-Evaluation Scales") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Internal Consistency (Omega) -------------------------------------------
# Helper Function for Omega Calculation
calculate_omega_total <- function(data, scale_name) {
  if (nrow(data) == 0 || ncol(data) == 0) {
    warning(paste("No valid data for", scale_name))
    return(NA)
  }

  # Try Omega Calculation
  result <- tryCatch(
    omega(data),
    error = function(e) {
      warning(paste("Omega calculation failed for", scale_name, ":", e$message))
      return(NULL)
    }
  )

  # Check if Result is Valid and Extract omega.tot
  if (is.null(result) || !"omega.tot" %in% names(result)) {
    warning(paste("Omega total not available for", scale_name))
    return(NA)
  }

  return(round(result$omega.tot, 2))
}

### CLASSROOM MANAGEMENT -------------------------------------------------
# Select relevant columns and remove NAs
df_cm <- df_quest %>% select(starts_with("LM"))

# Calculate omega
omega_cm <- calculate_omega_total(df_cm, "Classroom Management")

### NON & PARAVERBAL COMMUNICATION --------------------------------------
# Select relevant columns and remove NAs
df_npvc <- df_quest %>% select(starts_with("LP"))

# Calculate omega
omega_npvc <- calculate_omega_total(df_npvc, "Non-/Paraverbal Communication")

### OVERALL TEST -----------------------------------------------------
# Select relevant columns and remove NAs
df_self_all <- df_quest %>%
  select(matches("L")) %>%
  mutate(across(everything(), as.numeric)) %>%  # Ensure numeric data
  drop_na()

# Calculate omega
omega_self_all <- calculate_omega_total(df_self_all, "Overall Evaluation")

### COMBINE RESULTS INTO AN APA-CONFORM TABLE ----------------------------
# APA-Conform Table ------------------------------------------------------
omega_results <- tibble(
  Facet = c("Overall Evaluation", "Classroom Management", "Non-/Paraverbal Communication"),
  `Omega Total` = c(omega_self_all, omega_cm, omega_npvc)
)

# Display APA-Conform Table ----------------------------------------------
kable(omega_results, caption = "Internal Consistency (Omega) for Self-evaluation") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```
## Correlations

```{r correlation, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# 1. Load and Prepare Data -----------------------------------------------

# Demo Data
df_demo <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  filter(!LI06_05 %in% c(201, 223)) %>%  # Exclude participants without eye-tracking data
  transmute(
    Participant = LI06_05,
    `Teaching Experience` = LI04_01
  ) %>%
  arrange(Participant)

# ET-Measures (Micro-Teaching Unit) - Exclude Disruptive Person
df_aoi_numb_dur <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select(
    Participant,
    starts_with("Total_duration_of_fixations"),
    starts_with("Number_of_fixations"),
    starts_with("Average_duration"),
    Duration_of_interval  # Ensure this variable is included
  ) %>%
  dplyr::select(
    -contains("Disruptive_Person")  # Exclude Disruptive Person
  ) %>%
  rowwise() %>%
  transmute(
    Participant,
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Number_fixation_min_mtu = round(sum(c_across(starts_with("Number_of")), na.rm = TRUE) /
                                      Duration_of_interval_min, 0),
    Average_duration_mtu = round(sum(c_across(starts_with("Total_duration")), na.rm = TRUE) /
                                    sum(c_across(starts_with("Number_of")), na.rm = TRUE), 0),
    GRI_mtu = round(Average_duration_mtu / Number_fixation_min_mtu, 2)
  ) %>%
  arrange(Participant)

# ET-Measures (AOI Students) - Exclude Disruptive Person
df_aoi_stud <- df_aoi %>%
  filter(TOI == "Lesson") %>%
  dplyr::select (
    Group,
    Duration_of_interval,
    Participant,
    "Total_duration_of_fixations.Anna",
    "Total_duration_of_fixations.Bianca",
    "Total_duration_of_fixations.Carl(a)",
    "Number_of_fixations.Anna",
    "Number_of_fixations.Bianca",
    "Number_of_fixations.Carl(a)"
  ) %>%
  rowwise() %>%
  transmute(
    Group = as_factor(Group),
    Duration_of_interval_min = round(Duration_of_interval / 60000, 2),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2),
    Participant = Participant,
    Stud_duration_fixation = sum(c_across(starts_with("Total_duration")), na.rm = TRUE),
    Stud_number_fixation = sum(c_across(starts_with("Number_of")), na.rm = TRUE),
    Stud_number_fixation_min = round(Stud_number_fixation / Duration_of_interval_min, 2),
    Stud_number_fixation_sec = round(Stud_number_fixation / Duration_of_interval_sec, 2),
    Average_duration_stud = round(Stud_duration_fixation / Stud_number_fixation, 0)
  ) %>%
  drop_na()

# ET-Measures (AOI Disruptive Person)
# Number of fixations per second
# Define the different types of disruptions and categories
disruptions <- c(
  "Chatting_with_neighbour",
  "Clicking_pen",
  "Drawing",
  "Drumming_with_hands",
  "Head_on_table",
  "Heckling",
  "Looking_at_phone",
  "Snipping_with_fingers",
  "Whispering"
)
verbal_disruptions <- c("Chatting_with_neighbour",
                        "Heckling",
                        "Whispering")
physical_disruptions <- c("Clicking_pen",
                          "Drumming_with_hands",
                          "Snipping_with_fingers")
lack_of_eagerness <- c("Looking_at_phone",
                       "Head_on_table",
                       "Drawing")

# Step 1: Filter and prepare the data
df_aoi_disrup <- df_aoi %>%
  filter(TOI %in% disruptions) %>%  # Filter rows based on the disruption list
  select(
    Group,
    Participant,
    TOI,
    Duration_of_interval,
    "Number_of_fixations.Disruptive_Person"
  ) %>%
  mutate(
    Group = as_factor(Group),  # Convert Group column to a factor
    Disruption_Category = case_when(
      # Assign categories based on disruption type
      TOI %in% verbal_disruptions ~ "Verbal disruptions",
      TOI %in% physical_disruptions ~ "Physical disruptions",
      TOI %in% lack_of_eagerness ~ "Lack of eagerness to learn",
      TRUE ~ NA_character_
    ),
    Duration_of_interval_sec = round(Duration_of_interval / 1000, 2)  # Convert duration to seconds
  ) %>%
  filter(!is.na(Disruption_Category))  # Keep only rows with valid categories

# Step 2: Aggregate data by Participant and Disruption Category
df_aoi_numb_disrup <- df_aoi_disrup %>%
  group_by(Participant, Disruption_Category) %>%
  summarise(
    Total_Number_of_Fixations = sum(`Number_of_fixations.Disruptive_Person`, na.rm = TRUE),
    Total_Duration_of_Interval_Sec = sum(Duration_of_interval_sec, na.rm = TRUE),
    .groups = "drop"  # Avoid unnecessary grouping
  )

# Step 3: Pivot the data to create a single row for each Participant
df_summary <- df_aggregated %>%
  pivot_wider(
    names_from = Disruption_Category,
    values_from = c(Total_Number_of_Fixations, Total_Duration_of_Interval_Sec),
    values_fill = 0  # Replace missing values with 0
  )

# SRI Rating Data
df_sri <- read_xlsx("data/Coding_SRI.xlsx") %>%
  filter(!ID %in% c(201, 223)) %>%
  rename(Participant = ID) %>%
  group_by(Participant) %>%
  summarise(
    Mean_disruption_appraisal = mean(disruption_appraisal, na.rm = TRUE),
    Mean_confidence_appraisal = mean(confidence_appraisal, na.rm = TRUE),
    Mean_prevalence_rating = mean(prevalence_rating, na.rm = TRUE)
  )

# Situational Judgment Test (SJT)
df_sjt <- read_excel("./data/SJT.xlsx") %>%
  filter(!UI06_05 %in% c(201, 223)) %>%
  transmute(
    Participant = UI06_05,
    SJT_All = SJT_KF_gek,
    SJT_Monitoring = SJT_AL_gek
  )

# Self-Evaluation Data
df_self_eval <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>%
  filter(!LI06_05 %in% c(201, 223)) %>%
  mutate(across(starts_with("L"), as.numeric)) %>%  # Convert all "L"-prefix columns to numeric
  transmute(
    Participant = LI06_05,
    Self_eval_classroom_management = rowMeans(select(., starts_with("LM")), na.rm = TRUE),
    Self_eval_non_paraverbal_communication = rowMeans(select(., starts_with("LP")), na.rm = TRUE),
    Self_eval_total = rowMeans(select(., starts_with("L")), na.rm = TRUE)
  )

# 2. Combine All Data ---------------------------------------------------

# Combine all prepared datasets
df_list <- list(
  df_demo,
  df_aoi_numb_dur,
  df_aoi_stud,
  df_aoi_disrup,
  df_sri,
  df_sjt,
  df_self_eval
)

df_merge <- reduce(df_list, full_join, by = "Participant") %>%
  select(
    GRI_mtu, SJT_All, Self_eval_total, Mean_disruption_appraisal, Mean_confidence_appraisal
  )

# 3. Correlation Analysis -----------------------------------------------

# Compute Correlation Matrix
cor_tab <- cor(df_merge, method = "pearson", use = "complete.obs") %>% round(2)

# Create APA-Style Correlation Table
apa.cor.table(
  cor_tab,
  filename = "correlation_table.docx"
)

# p-values for specific correlations
pval_gri_sjt <- cor.test(df_merge$GRI_mtu, df_merge$SJT_All)$p.value
pval_gri_self_eval <- cor.test(df_merge$GRI_mtu, df_merge$Self_eval_total)$p.value

# Print Correlation Results
print(cor_tab)
cat("P-value (GRI x SJT):", pval_gri_sjt, "\n")
cat("P-value (GRI x Self-evaluation):", pval_gri_self_eval, "\n")

```